### 2021-03-11 visualisations_for_blog.Rmd

---
title: "Visualisations for blog on SCADR website"
author: "Jan Savinc"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r, warning=FALSE, message=FALSE}
library(tidyverse)  # for tidy workflow
library(opendatascot)  # importing data from ScotGov open data website
library(phsmethods)  # methods for working with PHS data
library(readxl)  # for reading excel files
library(lubridate)  # dealing with dates
library(janitor)  # for cleaning column names
library(ISOweek)  # for computing date from ISO week number + year
library(sf)  # for mapping
library(ggrepel)  # for 'mapping'repelling' labels/texst in ggplot
library(patchwork)  # for assembling plots
library(extrafont)  # for working with fonts
library(openxlsx)  # for creating xlsx files
library(scico)  # for continuous grayscale palette
library(kable)  # for printing tables nicely
```

# Load workspace

This was previously saved!

```{r}
load(file = "./workspace.RData")
```

# Set up output directory

Because data are updated on a weekly basis, we'll have a separate directory for every week the data are updated:

```{r}
dir_blog <- paste0("./outputs/scadr_blog_",latest_data_date)
if (!dir.exists(dir_blog)) dir.create(dir_blog)
```

# Tables

# Visualisations

## Common elements

```{r}
## modify theme with relevant font
theme_set(theme_minimal(base_size = 12) +
            theme(panel.grid.minor = element_blank(),
                  axis.title.y = element_text(margin = margin(0, 20, 0, 0)),  # remove gap to the left of y axis title and below the x axis title
                  axis.title.x = element_text(margin = margin(20, 0, 0, 0)),
                  text=element_text(family="Calibri")
            ))
```

## Home deaths by HB: Forth Valley vs Dumfries

```{r}
(home_deaths_by_hb_fv_vs_dag_2week_average <-
  merged_deaths_hb %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%
  filter(ref_area %in% c("Dumfries and Galloway", "Forth Valley")) %>%
  pivot_longer(cols = c(ma2w_deaths_all_causes, ma2w_deaths_non_covid, ma2w_deaths_mean), 
               names_to = "deaths_type", values_to = "ma2w_deaths") %>% 
  mutate(
    deaths_type = factor(x = deaths_type, levels = c("ma2w_deaths_all_causes", "ma2w_deaths_non_covid", "ma2w_deaths_mean"), labels = c("All deaths", "Non Covid-19 deaths", "Historic mean"))
    ) %>%
  mutate(fill_text = "Historic range (min/max)") %>%  # added for ease of using aes(fill) later
  ggplot(aes(x = date_w_c)) +
  annotate(geom = "rect", xmin = compute_start_date_from_week_number(week_number = 1, year_number = 2021), xmax = compute_start_date_from_week_number(week_number = weeks_available_2021, year_number = 2021), ymin = -Inf, ymax = Inf, fill = "grey90", alpha = 0.2) +  # shade area - geom_rect doesn't work here
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "8 weeks", limits = c(earliest_date, most_recent_date_available_2021)) +
  geom_ribbon(aes(ymin = ma2w_deaths_min, ymax = ma2w_deaths_max, fill = fill_text), alpha = 0.5) +
  geom_line(aes(y = ma2w_deaths, colour = deaths_type, linetype = deaths_type)) +
  ylim(0, NA) +  # start scale at 0 for better understanding the scale of deaths
  facet_grid(~ref_area) +
  scale_linetype_manual(values = c("solid","solid","dashed")) +
  scale_fill_manual(values = c("#4477aa")) +  # specify historic range fill
  scale_colour_manual(values = c("red","blue","grey20")) +
  theme(
    legend.position = "top",
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, hjust = 1),
    legend.title = element_blank()
  ) +
  labs(x = "Date w/c",
       y = "Deaths, 2-week average (N)",
       subtitle = date_range_text,
       caption = paste0(c(death_registrations_text, source_nrs_text), collapse="\n")
       # colour = "Cause of death" 
       ) +
  # ## annotate year
  geom_text(aes(x=ymd("2021-01-07"), y=2, label = "2021"), hjust = 0, colour = "grey40") +
  NULL
)
save_output_file(filename = paste0(dir_blog,"/home_deaths_by_hb_fv_vs_dag_2week_average"), extensions = c(".pdf",".svg",".png"), plot = home_deaths_by_hb_fv_vs_dag_2week_average, width = 16, height = 12, dpi = 300, units = "cm")
```

```{r}
total_deaths_hb %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%
  filter(ref_area %in% c("Dumfries and Galloway", "Forth Valley")) %>%
  filter(year == 2020) %>%
  select(year, place_of_death, ref_area, deaths_all_causes, ratio_annual_deaths_to_historical_mean) %>%
  kable(caption = "Total deaths at home in 2020 for health boards.")

merged_deaths_hb %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%
  filter(ref_area %in% c("Dumfries and Galloway", "Forth Valley")) %>%
  group_by(year, place_of_death, ref_area, week_number) %>%
  summarise(
    deaths_all_causes = sum(deaths_all_causes),
    # deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
    deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
  left_join(
    past_deaths_hb %>% filter(year != "average") %>% group_by(week_number, place_of_death, ref_area) %>% summarise(
                                       average_weekly_deaths = sum(number_of_deaths) / 5,
                                       .groups = "drop"
                                     ),
    by = c("week_number", "place_of_death", "ref_area")
  ) %>% 
  group_by(place_of_death, ref_area) %>%
  summarise(
    period = glue::glue("Data from week 1, 2020 to week {weeks_available_2021}, 2021"),
    deaths_all_causes = sum(deaths_all_causes),
    deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
    deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
    mean_annual_deaths_2015_2019 = sum(average_weekly_deaths),
    .groups = "drop"
  ) %>%
  mutate(percent_increase_deaths_over_historic = (deaths_all_causes / mean_annual_deaths_2015_2019) - 1) %>%
  kable(caption = "Total deaths at home in data so far for health boards.")
```