### 2021-04-16 visualisations_for_blog.Rmd
Added age group to blog visualisations

---
title: "Visualisations for blog on SCADR website"
author: "Jan Savinc"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r, warning=FALSE, message=FALSE}
library(tidyverse)  # for tidy workflow
library(opendatascot)  # importing data from ScotGov open data website
library(phsmethods)  # methods for working with PHS data
library(readxl)  # for reading excel files
library(lubridate)  # dealing with dates
library(janitor)  # for cleaning column names
library(ISOweek)  # for computing date from ISO week number + year
library(sf)  # for mapping
library(ggrepel)  # for 'mapping'repelling' labels/texst in ggplot
library(patchwork)  # for assembling plots
library(extrafont)  # for working with fonts
library(openxlsx)  # for creating xlsx files
library(scico)  # for continuous grayscale palette
library(knitr)  # for printing tables nicely
```

# Load workspace

This was previously saved!

```{r}
load(file = "./workspace_with_visualisations.RData")
```

# Set up output directory

Because data are updated on a weekly basis, we'll have a separate directory for every week the data are updated:

```{r}
dir_blog <- paste0("./outputs/scadr_blog_",latest_data_modified_date)
if (!dir.exists(dir_blog)) dir.create(dir_blog)
```

# Tables

# Visualisations

## Common elements

```{r}
## modify theme with relevant font
theme_set(theme_minimal(base_size = 12) +
            theme(panel.grid.minor = element_blank(),
                  axis.title.y = element_text(margin = margin(0, 20, 0, 0)),  # remove gap to the left of y axis title and below the x axis title
                  axis.title.x = element_text(margin = margin(20, 0, 0, 0)),
                  text=element_text(family="Calibri")
            ))
```

## Home deaths by HB: Forth Valley vs Dumfries

Note: decided agaisnt doing that, hard to explain why we're singling HBs out!

```{r}
# (home_deaths_by_hb_fv_vs_dag_2week_average <-
#   merged_deaths_hb %>%
#   filter(place_of_death %in% c("Home & other non-institution")) %>%
#   filter(ref_area %in% c("Dumfries and Galloway", "Forth Valley")) %>%
#   pivot_longer(cols = c(ma2w_deaths_all_causes, ma2w_deaths_non_covid, ma2w_deaths_mean), 
#                names_to = "deaths_type", values_to = "ma2w_deaths") %>% 
#   mutate(
#     deaths_type = factor(x = deaths_type, levels = c("ma2w_deaths_all_causes", "ma2w_deaths_non_covid", "ma2w_deaths_mean"), labels = c("All deaths", "Non Covid-19 deaths", "Historic mean"))
#     ) %>%
#   mutate(fill_text = "Historic range (min/max)") %>%  # added for ease of using aes(fill) later
#   ggplot(aes(x = date_w_c)) +
#   annotate(geom = "rect", xmin = compute_start_date_from_week_number(week_number = 1, year_number = 2021), xmax = compute_start_date_from_week_number(week_number = weeks_available_2021, year_number = 2021), ymin = -Inf, ymax = Inf, fill = "grey90", alpha = 0.2) +  # shade area - geom_rect doesn't work here
#   scale_x_date(date_labels = "%d-%m-%y", date_breaks = "8 weeks", limits = c(earliest_date, most_recent_date_available_2021)) +
#   geom_ribbon(aes(ymin = ma2w_deaths_min, ymax = ma2w_deaths_max, fill = fill_text), alpha = 0.5) +
#   geom_line(aes(y = ma2w_deaths, colour = deaths_type, linetype = deaths_type)) +
#   ylim(0, NA) +  # start scale at 0 for better understanding the scale of deaths
#   facet_grid(~ref_area) +
#   scale_linetype_manual(values = c("solid","solid","dashed")) +
#   scale_fill_manual(values = c("#4477aa")) +  # specify historic range fill
#   scale_colour_manual(values = c("red","blue","grey20")) +
#   theme(
#     legend.position = "top",
#     plot.caption = element_text(size = 10, colour = "gray60"),
#     plot.margin = margin(5, 15, 5, 5),
#     axis.text.x = element_text(angle = 60, hjust = 1),
#     legend.title = element_blank()
#   ) +
#   labs(x = "Date w/c",
#        y = "Deaths, 2-week average (N)",
#        subtitle = date_range_text,
#        caption = paste0(c(death_registrations_text, source_nrs_text), collapse="\n")
#        # colour = "Cause of death" 
#        ) +
#   # ## annotate year
#   geom_text(aes(x=ymd("2021-01-07"), y=2, label = "2021"), hjust = 0, colour = "grey40") +
#   NULL
# )
# save_output_file(filename = paste0(dir_blog,"/home_deaths_by_hb_fv_vs_dag_2week_average"), extensions = c(".pdf",".svg",".png"), plot = home_deaths_by_hb_fv_vs_dag_2week_average, width = 16, height = 12, dpi = 300, units = "cm")
```

```{r}
# total_deaths_hb %>%
#   filter(place_of_death %in% c("Home & other non-institution")) %>%
#   filter(ref_area %in% c("Dumfries and Galloway", "Forth Valley")) %>%
#   filter(year == 2020) %>%
#   select(year, place_of_death, ref_area, deaths_all_causes, ratio_annual_deaths_to_historical_mean) %>%
#   kable(caption = "Total deaths at home in 2020 for health boards.")
# 
# merged_deaths_hb %>%
#   filter(place_of_death %in% c("Home & other non-institution")) %>%
#   filter(ref_area %in% c("Dumfries and Galloway", "Forth Valley")) %>%
#   group_by(year, place_of_death, ref_area, week_number) %>%
#   summarise(
#     deaths_all_causes = sum(deaths_all_causes),
#     # deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
#     deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
#     .groups = "drop"
#   ) %>% 
#   mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
#   left_join(
#     past_deaths_hb %>% filter(year != "average") %>% group_by(week_number, place_of_death, ref_area) %>% summarise(
#                                        average_weekly_deaths = sum(number_of_deaths) / 5,
#                                        .groups = "drop"
#                                      ),
#     by = c("week_number", "place_of_death", "ref_area")
#   ) %>% 
#   group_by(place_of_death, ref_area) %>%
#   summarise(
#     period = glue::glue("Data from week 1, 2020 to week {weeks_available_2021}, 2021"),
#     deaths_all_causes = sum(deaths_all_causes),
#     deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
#     deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
#     mean_annual_deaths_2015_2019 = sum(average_weekly_deaths),
#     .groups = "drop"
#   ) %>%
#   mutate(percent_increase_deaths_over_historic = (deaths_all_causes / mean_annual_deaths_2015_2019) - 1) %>%
#   kable(caption = "Total deaths at home in data so far for health boards.")
```

## Home death increase by HB, 2020 to now

For this figure, I can't take annual average deaths to compare against because only for 2020 do we have the complete year, whereas for 2021 only part of the year is available. Instead, I add up weekly totals for the period in question, and average over the totals; in other words, if looking at the period 2020 w1 to 2021 w15, weeks 1-15 from 2015-2019 get used twice!

```{r, fig.height=12, fig.width=8}
(figure_map_ratio_home_deaths_to_historical_hb <-
  merged_deaths_hb %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%
  group_by(year, place_of_death, ref_area, week_number) %>%
  summarise(
    deaths_all_causes = sum(deaths_all_causes),
    deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
  left_join(
    past_deaths_hb %>% filter(year != "average") %>% group_by(week_number, place_of_death, ref_area) %>% summarise(
                                       average_weekly_deaths = sum(number_of_deaths) / 5,
                                       .groups = "drop"
                                     ),
    by = c("week_number", "place_of_death", "ref_area")
  ) %>% 
  group_by(place_of_death, ref_area) %>%
  summarise(
    period = glue::glue("Data from week 1, 2020 to week {weeks_available_2021}, 2021"),
    deaths_all_causes = sum(deaths_all_causes),
    deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
    deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
    mean_annual_deaths_2015_2019 = sum(average_weekly_deaths),
    .groups = "drop"
  ) %>%
  mutate(percent_increase_deaths_over_historic = (deaths_all_causes / mean_annual_deaths_2015_2019) - 1) %>%
  left_join(shapefile_hb_uk_wide, by=c("ref_area"="name")) %>%
  ggplot(aes(geometry=geometry, fill=percent_increase_deaths_over_historic), colour="grey10") +
  geom_sf(colour="grey10") +
  scale_fill_viridis_b(option = "B", labels = function(x) scales::percent(x, accuracy = 1)) +
  theme(
    axis.text = element_blank(), 
    panel.grid = element_blank(), 
    aspect.ratio = 1.4,  # needed for Scotland HB plot
    plot.margin = margin(5,5,5,5),
    ) +
  labs(x="", y="",
       title = str_wrap("Increased home deaths since 2020, by Health Board", width=40),
       subtitle = str_wrap(date_range_text, width=50),
       caption = paste0(c(source_nrs_text), collapse="\n"),
       fill = "Home death\nincrease"
       ) +
  NULL
)

save_output_file(filename = paste0(dir_blog,"/figure_map_ratio_home_deaths_to_historical_hb"), extensions = c(".pdf",".svg",".png"), plot = figure_map_ratio_home_deaths_to_historical_hb, width = 12, height = 14, dpi = 300, units = "cm")
```

### Table showing increases by area

```{r}
table_increase_home_deaths_by_hb <-
  merged_deaths_hb %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%
  group_by(year, place_of_death, ref_area, week_number) %>%
  summarise(
    deaths_all_causes = sum(deaths_all_causes),
    deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
  left_join(
    past_deaths_hb %>% filter(year != "average") %>% group_by(week_number, place_of_death, ref_area) %>% summarise(
                                       average_weekly_deaths = sum(number_of_deaths) / 5,
                                       .groups = "drop"
                                     ),
    by = c("week_number", "place_of_death", "ref_area")
  ) %>% 
  group_by(place_of_death, ref_area) %>%
  summarise(
    period = glue::glue("Data from week 1, 2020 to week {weeks_available_2021}, 2021"),
    deaths_all_causes = sum(deaths_all_causes),
    deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
    deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
    mean_annual_deaths_2015_2019 = sum(average_weekly_deaths),
    .groups = "drop"
  ) %>%
  mutate(percent_increase_deaths_over_historic = (deaths_all_causes / mean_annual_deaths_2015_2019) - 1) %>%
  arrange(percent_increase_deaths_over_historic) %>%
  left_join(
    weekly_covid_cases_hb %>%
    filter(year == 2021 & week_number == weeks_available_2021) %>%
    select(ref_area, crude_rate_positive_per100k),
    by = c("ref_area")
    ) %>%
  mutate(
    `Percent increase` = scales::percent(percent_increase_deaths_over_historic, accuracy = 0.1),
    `COVID cases per 100,000` = format(round(crude_rate_positive_per100k, digits=0), big.mark = ",")
    ) %>%
  select(`Health Board` = ref_area, `Home deaths` = deaths_all_causes, `Percent increase`, `COVID cases per 100,000`)

write_csv(x = table_increase_home_deaths_by_hb, file = paste0(dir_blog,"/table_increase_home_deaths_by_hb.csv"))
```

## Home deaths by age group

```{r, fig.width=12, fig.height=9, warning=FALSE}
removed_under_15_text <- glue::glue("Excluding individuals aged <15 at death, total N={merged_deaths_age %>% filter(age %in% c('0','1-14') & place_of_death == 'Home & other non-institution') %>% .$deaths_all_causes %>% sum} for weeks 1-{max(merged_deaths_age %>% filter(year==2020) %>% .$week_number)} in 2020 and 1-{max(merged_deaths_age %>% filter(year==2021) %>% .$week_number)} in 2021.")

## to put the arrows explaining historic range etc. only in one panel, need to construct a tibble to pass as data
annotations_figure_home_deaths_by_age <- tribble(
   ~geom, ~place_of_death, ~age, ~label, ~x, ~y, ~xend, ~yend,
  "text", "Home & other non-institution", "45-64",  "Historic range (min & max)", 17, 20, NA, NA,
  "curve", "Home & other non-institution", "45-64", NA, 22, 25, 25, 43,
  "text", "Home & other non-institution", "45-64", "Historic mean", 42, 20, NA, NA,
  "curve", "Home & other non-institution", "45-64", NA, 42, 25, 44, 64
) %>% 
  mutate(across(.cols = c("x","xend"), .fns = ~compute_start_date_from_week_number(week_number = .x, year_number = 2020)))

(figure_deaths_at_home_by_age <-
  merged_deaths_age %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%
  filter(!age %in% c('0','1-14')) %>%
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>% 
  # ggplot(aes(x = week_number_run_over)) +
  ggplot(aes(x = date_w_c)) +
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "8 weeks", limits = c(earliest_date, most_recent_date_available_2021)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = deaths_mean), linetype = "dashed", size = 0.2) +
  # geom_vline(xintercept = 54, linetype = "dotted") +
  annotate(geom = "rect", xmin = compute_start_date_from_week_number(week_number = 1, year_number = 2021), xmax = compute_start_date_from_week_number(week_number = weeks_available_2021, year_number = 2021), ymin = -Inf, ymax = Inf, fill = "grey90", alpha = 0.2) + 
  # facet_wrap(~age, scales = "free_y") +
  facet_wrap(~age) +
  # theme_minimal() +

  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
   theme(
    legend.position = "top",
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = "Week number",
       y = "Deaths (N)",
       subtitle = date_range_text,
       caption = paste0(c(removed_under_15_text, source_nrs_text), collapse="\n"),
       colour = "Cause of death"
       ) +
  ## annotate historic range
  geom_text(data = annotations_figure_home_deaths_by_age %>% filter(geom=="text"), aes(x=x, y=y, label=label),  colour = "grey20", size = 8 / (14/5)) +  # the size is tricky
  geom_curve(data = annotations_figure_home_deaths_by_age %>% filter(geom=="curve"), aes(x=x, y=y, xend=xend, yend=yend), colour = c("#4477aa", "grey20"), alpha = c(0.5, 1), linetype = c("solid","dashed"), curvature = 0.20, arrow = arrow(length = unit(0.2, "cm"))) +
  annotate(geom = "text", x=ymd("2021-01-07"), y=10, label = "2021", hjust = 0, colour = "grey40") +
  NULL
)

save_output_file(filename = paste0(dir_blog,"/figure_deaths_at_home_by_age"), extensions = c(".pdf",".svg",".png"), plot = figure_deaths_at_home_by_age, width = 12, height = 9, dpi = 300, units = "in")
```

### Table showing increases by age

```{r}
table_increase_home_deaths_by_age <-
  merged_deaths_age %>%
  filter(place_of_death %in% c("Home & other non-institution")) %>%
  group_by(year, place_of_death, age, week_number) %>%
  summarise(
    deaths_all_causes = sum(deaths_all_causes),
    deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
  left_join(
    past_deaths_age %>% filter(year != "average") %>% group_by(week_number, place_of_death, age) %>% summarise(
                                       average_weekly_deaths = sum(number_of_deaths) / 5,
                                       .groups = "drop"
                                     ),
    by = c("week_number", "place_of_death", "age")
  ) %>% 
  group_by(place_of_death, age) %>%
  summarise(
    period = glue::glue("Data from week 1, 2020 to week {weeks_available_2021}, 2021"),
    deaths_all_causes = sum(deaths_all_causes),
    deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
    deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
    mean_annual_deaths_2015_2019 = sum(average_weekly_deaths),
    .groups = "drop"
  ) %>%
  mutate(percent_increase_deaths_over_historic = (deaths_all_causes / mean_annual_deaths_2015_2019) - 1) %>%
  arrange(percent_increase_deaths_over_historic) %>%
  mutate(
    `Percent increase` = scales::percent(percent_increase_deaths_over_historic, accuracy = 0.1)
    ) %>%
  filter(!age %in% c("0", "1-14")) %>%  # remove the v.small groups
  select(`Age Group` = age, `Home deaths` = deaths_all_causes, `Percent increase`)

write_csv(x = table_increase_home_deaths_by_age, file = paste0(dir_blog,"/table_increase_home_deaths_by_age.csv"))
```