### 2021-08-04 mocking_up_distance_calculations.Rmd
Started computing distance to hospital computations, using data zones as geographical unit (and computing euclidean distance between OS grid reference coordinates)

---
title: "Mocking up distance/time to travel calculations for hospitals in Scotland"
author: "Jan Savinc"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
  code_folding: hide
toc: true
toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r, warning=FALSE, message=FALSE}
## install packages from github if not yet available!
# remotes::install_github("datasciencescotland/opendatascot", force = TRUE)
# remotes::install_github("Health-SocialCare-Scotland/phsmethods", force = TRUE)

library(tidyverse)  # for tidy workflow
library(opendatascot)  # importing data from ScotGov open data website
library(phsmethods)  # methods for working with PHS data
library(readxl)  # for reading excel files
library(curl)  # for downloading stuff
library(lubridate)  # dealing with dates
library(janitor)  # for cleaning column names
library(ISOweek)  # for computing date from ISO week number + year
library(sf)  # for mapping
library(ggrepel)  # for 'mapping'repelling' labels/texst in ggplot
library(patchwork)  # for assembling plots
library(extrafont)  # for working with fonts
library(openxlsx)  # for creating xlsx files
library(knitr)  # for displaying tables in rmd
library(tictoc)  # for measuring time taken for calculations
```

# Introduction

# Download files

The hospitals data needs to be downloaded, whereas the centroids file was provided by Dr Laurie Berry!

```{r}
url_hospitals <- "https://www.opendata.nhs.scot/dataset/cbd1802e-0e04-4282-88eb-d7bdcfb120f0/resource/c698f450-eeed-41a0-88f7-c1e40a568acc/download/current-hospital_flagged20210506.csv"

if (!file.exists("./distance_to_hospital/current-hospital_flagged20210506.csv")) download.file(url = url_hospitals, destfile = "./distance_to_hospital/current-hospital_flagged20210506.csv")

# 
# url_datazones <- "https://maps.gov.scot/ATOM/shapefiles/SG_DataZoneCent_2011.zip"
# 
# if (!file.exists("./distance_to_hospital/SG_DataZoneCent_2011.zip")) download.file(url = url_datazones, destfile = "./distance_to_hospital/SG_DataZoneCent_2011.zip")
```

# Load data

```{r}
hospitals <- read_csv(file = "./distance_to_hospital/current-hospital_flagged20210506.csv")

datazones <- read_excel("./distance_to_hospital/datazone2011 centroids.xlsx", sheet = 1)

pythagorean_distance <- function(x1, y1, x2, y2) {
  distance <- sqrt((x1 - x2)^2 + (y1 - y2)^2)
}

## check that the XCoordinate and YCoordinate and Easting and Northing are the same thing, by checking the DataZone centroid location against hospital location
hospitals %>%
  select(LocationName, DataZone, XCoordinate, YCoordinate, BasedOnPostcode) %>%
  left_join(datazones, by = c("DataZone")) %>%
  mutate(distance = pythagorean_distance(XCoordinate, YCoordinate, Easting, Northing)) %>%
  arrange(desc(distance))

isds1 <- read_csv("./distance_to_hospital/beds_by_nhs_board_of_treatment_and_specialty.csv", guess=1e5)
```

## Defining acute hospitals

From the [Hospital care website of ISD Scotland](https://www.isdscotland.org/health-topics/hospital-care/):

> This section of our website brings together information on different aspects of acute hospital care, sourced from hospital administrative systems across Scotland. Note that 'Acute' hospital care excludes obstetric, psychiatric and long stay care services that are covered elsewhere on the ISD website. For reference, a listing of all Scottish hospitals (acute and non-acute) can be found by following the Hospitals link on the menu to the left of this page.

Therefore, we can use an ISD(S)1 publication, in this case on Bed Occupancy (https://www.isdscotland.org/Health-Topics/Hospital-Care/Publications/2019-11-26/Acute-Hospital-Publication/data-files/docs/20191126_Beds_by_Health_Board_of_Treatment_and_Specialty.csv), to define Acute hospitals - all hospitals included in this publication are considered acute!

```{r}
acute_hospitals <- hospitals %>% semi_join(isds1, by = "Location")  # keep only hospitals present in the isd(s)1 dataset
```

This reduces the number of hospitals from `r n_distinct(hospitals$LocationName)` to `r n_distinct(acute_hospitals$LocationName)`!

# Compute straight-line distance for each datazone to nearest hospital

For this purpose, I've removed some specialist units that operate within a larger hospital department and are listed at the same geographical grid reference! Otherwise we get two units that are equidistant for some data zones.

```{r}
tic()
smallest_distance_straight_line_datazones <-
  crossing(
    datazones %>% select(DataZone), 
    hospitals %>% select(LocationName) %>% filter(!LocationName %in% c("Orchard View","William Fraser Centre", "Gartnavel General Hospital (Admin Purposes)"))  # remove units located in same place as another hospital
    ) %>%
  left_join(datazones %>% select(DataZone, Easting, Northing), by = "DataZone") %>%
  left_join(hospitals %>% select(LocationName, XCoordinate, YCoordinate), by = "LocationName") %>%
  mutate(distance = pythagorean_distance(x1 = Easting, x2 = XCoordinate, y1 = Northing, y2 = YCoordinate)) %>%
  group_by(DataZone) %>%
  slice_min(order_by = distance, n = 1) %>%
  ungroup
toc(log = TRUE)
```

## Acute hospital distances

```{r}
tic()
smallest_distance_straight_line_datazones_acute_hospitals <-
  crossing(
    datazones %>% select(DataZone), 
    acute_hospitals %>% select(LocationName)
    ) %>%
  left_join(datazones %>% select(DataZone, Easting, Northing), by = "DataZone") %>%
  left_join(hospitals %>% select(LocationName, XCoordinate, YCoordinate), by = "LocationName") %>%
  mutate(distance = pythagorean_distance(x1 = Easting, x2 = XCoordinate, y1 = Northing, y2 = YCoordinate)) %>%
  group_by(DataZone) %>%
  slice_min(order_by = distance, n = 1) %>%
  ungroup
toc(log = TRUE)
```

# Print session info

```{r}
sessionInfo()
```