### 2021-08-04 distance_to_nearest_hospital.Rmd
reorganise the file and wrote up an introduction

---
title: "Distance/time to travel calculations for hospitals in Scotland"
author: "Jan Savinc"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
  code_folding: hide
toc: true
toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r, warning=FALSE, message=FALSE}
## install packages from github if not yet available!
# remotes::install_github("datasciencescotland/opendatascot", force = TRUE)
# remotes::install_github("Health-SocialCare-Scotland/phsmethods", force = TRUE)

library(tidyverse)  # for tidy workflow
library(readxl)  # for reading excel files
library(tictoc)  # for measuring time taken for calculations
```

# Introduction

For investigating increased numbers of deaths at home & other non-institutional settings in Scotland, we'll take into account the distance/travel time to the nearest hospital. For this purpose, we need to compute the distance to nearest hospital.

There are two basic approaches: straight-line distance, and road distance / road travel time:

## Straight-line distance

This can be computed using OS National Grid Reference values (Easting & Northing), using the locations of hospitals, and either:

* data zones
* postcodes

Postcodes are more precise, but also more privacy-invasive when working with administrative records. For this reason, we would ideally provide the computed distances at data linkage rather than after data delivery.

Straight-line distances don't necessarily represent a realistic measure of distance or travel time to hospital, however! The main benefit is its ease of calculation.

## Road distance / travel time

This is more realistic & representative than straight-line distance, but more difficult to obtain, since road distances and road spee (used in calculating travel time) aren't openly available like the OS grid reference data are.

# Download files & load

The following data needs downloading:

* Hospitals with geographic coordinates
* Datazone centroids with geographic coordinates (provided by Dr Laurie Berrie!)
* Postcodes with geographic coordinates

The coordinates are given in Easting & Northing coordinates in the OS National Grid basis. We can compute the straight-line distance between them using the Pythagorean identity function:

```{r}
pythagorean_distance <- function(x1, y1, x2, y2) {
  distance <- sqrt((x1 - x2)^2 + (y1 - y2)^2)
}
```

## Hospitals

```{r}
url_hospitals <- "https://www.opendata.nhs.scot/dataset/cbd1802e-0e04-4282-88eb-d7bdcfb120f0/resource/c698f450-eeed-41a0-88f7-c1e40a568acc/download/current-hospital_flagged20210506.csv"

if (!file.exists("./distance_to_hospital/current-hospital_flagged20210506.csv")) download.file(url = url_hospitals, destfile = "./distance_to_hospital/current-hospital_flagged20210506.csv")

hospitals <- read_csv(file = "./distance_to_hospital/current-hospital_flagged20210506.csv")

## check that the XCoordinate and YCoordinate and Easting and Northing are the same thing, by checking the DataZone centroid location against hospital location
hospitals %>%
  select(LocationName, DataZone, XCoordinate, YCoordinate, BasedOnPostcode) %>%
  left_join(datazones, by = c("DataZone")) %>%
  mutate(distance = pythagorean_distance(XCoordinate, YCoordinate, Easting, Northing)) %>%
  arrange(desc(distance))
```

### Defining acute hospitals

From the [Hospital care website of ISD Scotland](https://www.isdscotland.org/health-topics/hospital-care/):

> This section of our website brings together information on different aspects of acute hospital care, sourced from hospital administrative systems across Scotland. Note that 'Acute' hospital care excludes obstetric, psychiatric and long stay care services that are covered elsewhere on the ISD website. For reference, a listing of all Scottish hospitals (acute and non-acute) can be found by following the Hospitals link on the menu to the left of this page.

Therefore, we can use an ISD(S)1 publication, in this case on Bed Occupancy (https://www.isdscotland.org/Health-Topics/Hospital-Care/Publications/2019-11-26/Acute-Hospital-Publication/data-files/docs/20191126_Beds_by_Health_Board_of_Treatment_and_Specialty.csv), to define Acute hospitals - all hospitals included in this publication are considered acute!

```{r}
isds1 <- read_csv("./distance_to_hospital/beds_by_nhs_board_of_treatment_and_specialty.csv", guess=1e5)

acute_hospitals <- hospitals %>% semi_join(isds1, by = "Location")  # keep only hospitals present in the isd(s)1 dataset
```

This reduces the number of hospitals from `r n_distinct(hospitals$LocationName)` to `r n_distinct(acute_hospitals$LocationName)`!

## DataZones

```{r}
# url_datazones <- "https://maps.gov.scot/ATOM/shapefiles/SG_DataZoneCent_2011.zip"
# 
# if (!file.exists("./distance_to_hospital/SG_DataZoneCent_2011.zip")) download.file(url = url_datazones, destfile = "./distance_to_hospital/SG_DataZoneCent_2011.zip")
datazones <- read_excel("./distance_to_hospital/datazone2011 centroids.xlsx", sheet = 1)
```

## Postcode directory

~~I'm using the August 2020 postcode directory [available here](https://geoportal.statistics.gov.uk/datasets/ons::ons-postcode-directory-august-2020/about)

I've opted for the `.csv` version and limited the columns to read because it is over 1GB in size!~~

The 2021-1 version of the Scottish postcode directory is [available here](https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/scottish-postcode-directory/2021-1)

There are a few postcodes with duplicate entries for the same postcode where the geographical location is slightly different between them. I've opted to keep the last entry in a series of duplicates, as it probably reflects the most recent information available.

```{r}
# postcodes <- read_csv(file = "./distance_to_hospital/ONSPD_AUG_2020_UK.csv", col_names = c("postcode_PCDS","easting","northing"), col_types = "__c________ii_____________________________________", skip = 1)  # this format ensures we only read the postcode and the easting & northing data

# paste0(rep("_", times = 50), collapse="")  # used for precisely specifying number of skipped columns!

postcodes <- read_csv(file = "./distance_to_hospital/SPD_PostcodeIndex_Cut_21_1_CSV/SmallUser.csv", col_names = c("postcode","easting","northing"), col_types = "c____ii__________________________________________________", skip = 1)  # this format ensures we only read the postcode and the easting & northing data

postcodes <-  # keep last entry where there is more than one record per postcode
  postcodes %>%
  group_by(postcode) %>%
  slice(n()) %>%
  ungroup
```

# DataZone straight-line distance to nearest hospital

For this purpose, I've removed some duplicate entries in the list of hospitals - I think they are specialist units that operate within a larger hospital and are listed at the same geographical grid reference! Otherwise we get two units that are equidistant for some data zones.

```{r}
tic()
smallest_distance_straight_line_datazones <-
  crossing(
    datazones %>% select(DataZone), 
    hospitals %>% select(LocationName) %>% filter(!LocationName %in% c("Orchard View","William Fraser Centre", "Gartnavel General Hospital (Admin Purposes)"))  # remove units located in same place as another hospital
    ) %>%
  left_join(datazones %>% select(DataZone, Easting, Northing), by = "DataZone") %>%
  left_join(hospitals %>% select(LocationName, XCoordinate, YCoordinate), by = "LocationName") %>%
  mutate(distance = pythagorean_distance(x1 = Easting, x2 = XCoordinate, y1 = Northing, y2 = YCoordinate)) %>%
  group_by(DataZone) %>%
  slice_min(order_by = distance, n = 1) %>%
  ungroup %>%
  select(DataZone, LocationName, distance)
toc(log = TRUE)
```

## Acute hospital distances

```{r}
tic()
smallest_distance_straight_line_datazones_acute_hospitals <-
  crossing(
    datazones %>% select(DataZone), 
    acute_hospitals %>% select(LocationName)
    ) %>%
  left_join(datazones %>% select(DataZone, Easting, Northing), by = "DataZone") %>%
  left_join(hospitals %>% select(LocationName, XCoordinate, YCoordinate), by = "LocationName") %>%
  mutate(distance = pythagorean_distance(x1 = Easting, x2 = XCoordinate, y1 = Northing, y2 = YCoordinate)) %>%
  group_by(DataZone) %>%
  slice_min(order_by = distance, n = 1) %>%
  ungroup %>%
  select(DataZone, LocationName, distance)
toc(log = TRUE)
```

# Postcode straight-line distance to acute hospitals

```{r}
tic()
smallest_distance_straight_line_postcodes_acute_hospitals <-
  crossing(
    postcodes %>% select(postcode), 
    acute_hospitals %>% select(LocationName)
    ) %>%
  left_join(postcodes %>% select(postcode, easting, northing), by = "postcode") %>%
  left_join(hospitals %>% select(LocationName, XCoordinate, YCoordinate), by = "LocationName") %>%
  mutate(distance = pythagorean_distance(x1 = easting, x2 = XCoordinate, y1 = northing, y2 = YCoordinate)) %>%
  group_by(postcode) %>%
  slice_min(order_by = distance, n = 1) %>%
  ungroup %>%
  select(postcode, LocationName, distance) %>%
  distinct
toc(log = TRUE)

smallest_distance_straight_line_postcodes_acute_hospitals %>%
  write_csv(file = "./distance_to_hospital/postcode_distance_to_nearest_acute_hospital_straightline.csv")
```

# Print session info

```{r}
sessionInfo()
```