### 2021-09-30 highlands_data_and_visualisations.Rmd
cleaned up script a bit and added a table summarising number of deaths in Highland; deprivation and cause of death aren't available on HB/LA level so they can't be included.

---
title: "Deaths at home in NHS Highland / Highland LA"
author: "Jan Savinc"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r, warning=FALSE, message=FALSE}
## install packages from github if not yet available!
# remotes::install_github("datasciencescotland/opendatascot", force = TRUE)
# remotes::install_github("Health-SocialCare-Scotland/phsmethods", force = TRUE)

library(tidyverse)  # for tidy workflow
library(opendatascot)  # importing data from ScotGov open data website
library(phsmethods)  # methods for working with PHS data
library(readxl)  # for reading excel files
# library(SPARQL)  # taken care of by opendatascot
library(lubridate)  # dealing with dates
library(janitor)  # for cleaning column names
library(ISOweek)  # for computing date from ISO week number + year
library(sf)  # for mapping
library(ggrepel)  # for 'mapping'repelling' labels/texst in ggplot
library(patchwork)  # for assembling plots
library(extrafont)  # for working with fonts
library(openxlsx)  # for creating xlsx files
library(knitr)  # for displaying tables in rmd
library(lemon)  # for showing axis & axis labels in every facet in ggplot
library(patchwork)  # for assembling multiple plots
```

# Copyright attribution

All data used for this study were obtained from the National Records Scotland (NRS) and are © Crown copyright, 2020; the details of the licence can be viewed on the [Open Government Licence website](http://www.nationalarchives.gov.uk/doc/open-government-licence/open-government-licence.htm)

# Load most recent data

```{r}
load(file = "./workspace_with_visualisations.RData")
```

# Visualisations

Note: visualisations have been moved to a separate .Rmd file

## Common elements

```{r}
loadfonts()  # load imported fonts
# Sys.setenv(R_GSCMD = "C:/Program Files/gs/gs9.27/bin/gswin64c.exe")  # set the location of the GS file  # this not needed, just use device=device_cairo for pdfs in ggsave

theme_set(theme_minimal(base_size = 12) +
            theme(panel.grid.minor = element_blank(),
                  axis.title.y = element_text(margin = margin(0, 20, 0, 0)),  # remove gap to the left of y axis title and below the x axis title
                  axis.title.x = element_text(margin = margin(20, 0, 0, 0)),
                  text=element_text(family="Calibri")
            ))

date_range_text <- glue::glue("Data from w/c {format(earliest_date, '%d %B %Y')} to w/c {format(most_recent_date_available_2021, '%d %B %Y')} (inclusive).")

dashed_line_historical_range_text <- "Range of deaths 2015-19 shown as shaded area; mean shown as dashed line."

only_2020_text <- glue::glue("Data from 2020 included, starting with w/c {format(earliest_date)}.")
only_2021_text <- glue::glue("Data for 2021 only include weeks 1-{format(weeks_available_2021)} (w/c {format(most_recent_date_available_2021)}).")

source_nrs_text <- "Source: National Records of Scotland"
source_phs_text <- "Source: Public Health Scotland"

death_registrations_text <- "Note: all figures are based on date of registration."

vertical_line_text <- "Vertical dotted line at week number 54 represents start of 2021."

```

## Function to save files to multiple formats

Saving to multiple formats is easy enough - the trouble is with saving PDF files with custom fonts, since they need to be embedded in the .pdf file. This requires:

* telling R where to load the font from, using `extrafont::font_import()` - this was done in a previous step
* registering fonts with R using `extrafont::loadfonts()` - note: this isn't strictly necessary, since once imported, `extrafont` keeps track of fonts to load when the `extrafont` package is loaded
* ~~telling R where the `gs` executable is, i.e. the installation of ghostscript on the current machine~~
* ~~embedding the fonts in the `.pdf` files once they've been written, using `extrafont::embed_fonts()`~~
* instead of manually embedding fonts (which resulted in garbled text on my setup!), just use `device=device_cairo` as an argument to `ggsave()` when saving `.pdf` files and it just works

```{r}
save_output_file <- function(filename, extensions, plot, width, height, units, dpi) {

  walk(
    .x = extensions,
    .f = function(ext) {
    # if (tolower(ext)==".pdf") device <- cairo_pdf
    ggsave(filename = paste0(filename, ext), plot = plot, width = width, height = height, units = units, dpi = dpi, device = if (tolower(ext)==".pdf") cairo_pdf else NULL)
    }
  )
}
```

# NHS Highlands / Highland LA data

On 28 September, Michael Loynd of NHS Highland asked for a data excerpt for NHS Highland and North Highland (sic) local authority.

## Create directory

```{r}
dir_outputs_highlands <- gsub(x = dir_outputs, pattern = "outputs", replacement = "outputs/highlands")
if (!dir.exists(dir_outputs_highlands)) dir.create(dir_outputs_highlands, recursive = TRUE)
```

## Table of deaths

```{r}
table_current_deaths_compared_to_past_highland <-
  bind_rows(
    ## all available data
    merged_deaths_hb %>%
      mutate(geography = "HB") %>%
      filter(ref_area == "Highland") %>%
      group_by(year, place_of_death, week_number, geography) %>%
      summarise(
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
      mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
      left_join(
        past_deaths_hb %>% filter(year != "average" & ref_area == "Highland") %>% group_by(week_number, place_of_death) %>% summarise(
                                           average_weekly_deaths = sum(number_of_deaths) / 5,
                                           .groups = "drop"
                                         ),
        by = c("week_number", "place_of_death")
      ) %>% 
      group_by(place_of_death, geography) %>%
      summarise(
        period = glue::glue("Data from week 1, 2020 to week {weeks_available_2021}, 2021"),
        deaths_all_causes = sum(deaths_all_causes),
        deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        mean_annual_deaths_2015_2019 = sum(average_weekly_deaths),
        .groups = "drop"
      )
    ,
    ## 2020 only
    merged_deaths_hb %>%
      mutate(geography = "HB") %>%
      filter(ref_area == "Highland" & year == 2020) %>%
      group_by(place_of_death, geography) %>%
      summarise(
        period = "2020",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
      mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      historical_annual_average_hb %>% 
        filter(ref_area == "Highland") %>%
        select(place_of_death, mean_annual_deaths_2015_2019 = annual_deaths_mean),
      by = "place_of_death"
    )
    ,
    ## 2020 from week 12 onwards (including week 12) - pandemic part of 2020
    merged_deaths_hb %>%
      mutate(geography = "HB") %>%
      filter(ref_area == "Highland" &
          year == 2020 & week_number >= 12
      ) %>%
      group_by(place_of_death, geography) %>%
      summarise(
        period = "2020, week 12 - week 53",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
    mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      past_deaths_hb %>% 
        filter(ref_area == "Highland", week_number >= 12 & year != "average") %>%
        group_by(place_of_death) %>%
        summarise(mean_annual_deaths_2015_2019 = sum(number_of_deaths) / 5),
      by = "place_of_death"
    ),
    ## all available data
    merged_deaths_la %>%
      mutate(geography = "LA") %>%
      filter(ref_area == "Highland") %>%
      group_by(year, place_of_death, week_number, geography) %>%
      summarise(
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
      mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
      left_join(
        past_deaths_la %>% filter(year != "average" & ref_area == "Highland") %>% group_by(week_number, place_of_death) %>% summarise(
                                           average_weekly_deaths = sum(number_of_deaths) / 5,
                                           .groups = "drop"
                                         ),
        by = c("week_number", "place_of_death")
      ) %>% 
      group_by(place_of_death, geography) %>%
      summarise(
        period = glue::glue("Data from week 1, 2020 to week {weeks_available_2021}, 2021"),
        deaths_all_causes = sum(deaths_all_causes),
        deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        mean_annual_deaths_2015_2019 = sum(average_weekly_deaths),
        .groups = "drop"
      )
    ,
    ## 2020 only
    merged_deaths_la %>%
      mutate(geography = "LA") %>%
      filter(ref_area == "Highland" & year == 2020) %>%
      group_by(place_of_death, geography) %>%
      summarise(
        period = "2020",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
      mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      historical_annual_average_la %>% 
        filter(ref_area == "Highland") %>%
        select(place_of_death, mean_annual_deaths_2015_2019 = annual_deaths_mean),
      by = "place_of_death"
    )
    ,
    ## 2020 from week 12 onwards (including week 12) - pandemic part of 2020
    merged_deaths_la %>%
      mutate(geography = "LA") %>%
      filter(ref_area == "Highland" &
          year == 2020 & week_number >= 12
      ) %>%
      group_by(place_of_death, geography) %>%
      summarise(
        period = "2020, week 12 - week 53",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
    mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      past_deaths_la %>% 
        filter(ref_area == "Highland", week_number >= 12 & year != "average") %>%
        group_by(place_of_death) %>%
        summarise(mean_annual_deaths_2015_2019 = sum(number_of_deaths) / 5),
      by = "place_of_death"
    )
  ) %>%
  mutate(
    ratio_deaths_all_causes_to_historical = deaths_all_causes / mean_annual_deaths_2015_2019,
    ratio_non_covid_deaths_to_historical = deaths_non_covid / mean_annual_deaths_2015_2019,
    proportion_covid_deaths = deaths_covid_related / deaths_all_causes
  ) %>% relocate(geography, place_of_death)

write_csv(x = table_current_deaths_compared_to_past_highland, file = file.path(dir_outputs_highlands, "annual_death_rates_comparison_with_historical_deaths_highland.csv"))
```

## Home deaths

```{r}
annotations_deaths_at_home_highland <- tribble(
  ~geography, ~place_of_death, ~x, ~y, ~label,
  "NHS Highland", "Home & other non-institution", 15, 36, 2020,
  "NHS Highland", "Home & other non-institution", 8, 43, 2021,
  "Highland (Local authority)", "Home & other non-institution", 21, 31, 2020,
  "Highland (Local authority)", "Home & other non-institution", 10, 28, 2021,
) %>% mutate(alpha = if_else(label==2021, 1, 0.4)) %>%   # change alpha for 2020
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home")))  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)

arrows_deaths_at_home_highland <- tribble(
  ~geography, ~geom, ~place_of_death, ~label, ~x, ~y, ~xend, ~yend,
  "NHS Highland", "text", "Home & other non-institution", "Historic range (min & max)", 11, 5, NA, NA,
  "NHS Highland", "curve", "Home & other non-institution", NA, 11, 7, 15, 11,
  "NHS Highland", "text", "Home & other non-institution", "Historic mean", 20, 1, NA, NA,
  "NHS Highland", "curve", "Home & other non-institution", NA, 23, 3, 29, 18
) %>% 
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home")))  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)

(figure_deaths_at_home_2020_and_2021_highland <-
  bind_rows(  # combine the two geographical types
    merged_deaths_hb %>%
      filter(ref_area == "Highland") %>%
      filter(place_of_death %in% c("Home & other non-institution")) %>%
      mutate(geography = "NHS Highland"),
    merged_deaths_la %>%
      filter(ref_area == "Highland") %>%
      filter(place_of_death %in% c("Home & other non-institution")) %>%
      mutate(geography = "Highland (Local authority)")
  ) %>%
  mutate(year = factor(year)) %>%
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid),
               names_to = "deaths_type", values_to = "deaths") %>%
  ggplot(aes(x = week_number)) +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  # geom_line(data = . %>% filter(year==2020), aes(y = deaths, colour = deaths_type, alpha = week_number, group = deaths_type), show.legend = FALSE) +
  geom_line(data = . %>% filter(year==2020), aes(y = deaths, colour = deaths_type, group = deaths_type), show.legend = FALSE, alpha = 0.3) +  # use a fixed alpha value, otherwise the lines are chopped up into rectangular segments and don't look good zoomed in
  geom_line(data = . %>% filter(year==2021), aes(y = deaths, colour = deaths_type)) +
  scale_x_continuous(breaks = seq(0,60,by=12)) +
  ylim(0, 50) +  # start scale at 0 for better understanding the scale of deaths
  facet_grid(~geography) +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths",
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  scale_alpha_continuous(range = c(0.2, 0.6), guide = NULL) +
  theme(legend.position = "top", plot.caption = element_text(size = 10, colour = "gray60")) +
  labs(x = "Week number",
       y = "Weekly deaths (N)",
       title = "Highland: Health Board & Local Authority",
       subtitle = date_range_text,
       caption = paste0(c(death_registrations_text, source_nrs_text), collapse="\n"),
       colour = "Cause of death"
       ) +
  ## annotate historic range
  geom_text(data = annotations_deaths_at_home_highland %>% filter(label==2021), aes(x=x, y=y, label=label), fontface = "bold") +
  geom_text(data = annotations_deaths_at_home_highland %>% filter(label==2020), aes(x=x, y=y, label=label), alpha = 0.6) +
  geom_text(data = arrows_deaths_at_home_highland %>% filter(geom=="text"), aes(x=x,y=y,label=label)) +
  geom_curve(data = arrows_deaths_at_home_highland %>% filter(geom=="curve") %>% slice(1), aes(x=x,y=y,xend=xend,yend=yend), curvature = 0.20, arrow = arrow(length = unit(0.2, "cm")), alpha = 0.5) +
  geom_curve(data = arrows_deaths_at_home_highland %>% filter(geom=="curve") %>% slice(2), aes(x=x,y=y,xend=xend,yend=yend), curvature = 0.20, arrow = arrow(length = unit(0.2, "cm")), linetype = "dashed") +
  NULL
)

save_output_file(filename = paste0(dir_outputs_highlands,"/deaths_at_home_2020_and_2021_highland"), extensions = c(".pdf",".png"), plot = figure_deaths_at_home_2020_and_2021_highland, width = 24, height = 14, dpi = 300, units = "cm")
```

## Proportion of deaths by place by year

```{r}
(figure_proportion_deaths_by_place_by_year_highland <-
  bind_rows(
    merged_proportions_of_deaths_by_place %>% mutate(geography = "scotland"),
    proportions_of_deaths_by_place_by_geography
    ) %>%
  mutate(place_of_death = factor(x = place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4)])) %>%  # move home to 2nd position to match order in plot
  mutate(geography = case_when(
    geography == "hb" ~ "NHS Highland",
    geography == "la" ~ "Highland (Local Authority)",
    tolower(geography) == "scotland" ~ "Scotland",
    TRUE ~ NA_character_
  )) %>%
  filter(
    geography == "Scotland" | (!is.na(ref_area) & ref_area == "Highland")
  ) %>%
  ggplot(., aes(x=year, y=proportion_of_total, colour=place_of_death)) +
  geom_point() +
  geom_line(size = 1) +
  geom_text(data = ~filter(.x, year==2021), aes(x=2021.2, y=proportion_of_total, label="*"), inherit.aes = FALSE) +
  scale_x_continuous(breaks = 2015:2021) +
  scale_y_continuous(label = function(x) scales::percent(x = x, accuracy = 1)) +
  scale_colour_viridis_d(option="C") +
  facet_wrap(~geography) +
  theme(legend.position = "top", plot.caption = element_text(size = 10, colour = "gray60")) +
  labs(
    x = "Year", y= "Proportion of yearly deaths (%)",
    colour = "",
    title = "Highland Health Board & Local Authority",
    subtitle = paste0("* ",only_2021_text),
    caption = paste0(c(source_nrs_text),collapse="\n")
    ) +
  guides(colour = guide_legend(nrow=2)) +  # spread colour legend over two rows
  NULL
)

save_output_file(filename = paste0(dir_outputs_highlands,"/figure_proportion_home_deaths_over_time_highland"), extensions = c(".pdf",".png"), plot = figure_proportion_deaths_by_place_by_year_highland, width = 22, height = 10, units = "cm", dpi = 300)
```

## Cause of death & deprivation

Unfortunately, as of 30 September 2021 these aren't available in combination with HB/LA, so they can't be reported for Highland individually.

# Print session info

```{r}
sessionInfo()
```