### 2021-11-08 extract_elixhauser_icd_10_codes_quan_paper_from_icd_package.Rmd
Wrote up extraction of Elixhauser comorbidity codes

---
title: "Extract Elixhauser ICD-10 comorbidity codes from `icd` package"
author: "Jan Savinc"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
bibliography: references_extract_elixhauser.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

To use Elixhauser comorbidity codes [@elixhauser] in our study, we needed to provide the criteria for each comorbidity to eDRIS - these will be used to derive binary variables for each person based on their historical hospitalisation records. The original codes were provided for ICD-9-CM, but have since been translated to ICD-10 [@quan]. Unfortunately, the paper is from 2005 and only available as a `.pdf` file, and the codes are listed in a table broken over two pages, so they are not very easy to extract from the paper. The OCR done on the paper isn't great, either, misreading codes, for example, `I12` was read as `112` instead.

Fortunately, the `icd` package [@icd] contains a list of Elixhauser code from the Quan paper, so they can be extracted from the package instead of having to be input by hand.

# Install & load `icd` package

As of November 2021, the `icd` package isn't available on CRAN, so I've used the github version instead!

```{r, warning=FALSE, message=FALSE}
## uncomment if installation needed - note it will take a while because it needs compilation
# remotes::install_github("jackwasey/icd")

library(tidyverse)  # for tidy workflow
library(openxlsx)  # for writing excel files
```

# Get the codes out of `icd`

The object `icd::icd10_map_quan_elix` contains the Quan et al. (2005) provided codes, but its default printing method only shows the first 7 comorbidities and the first 7 codes associated with each. We need to `unclass()` it to get the underlying `list` object!

Note that the convention here is that ICD-10 stands for ICD-10-CM (Clinical Modification, the version of ICD-10 used in the USA) - some of these codes don't exist in the UK (WHO) version of ICD-10. It is also clear that rather than having been specified generally, e.g. F18 for drug abuse, all valid ICD-10-CM codes within F18 were enumerated... for our purposes, this is too detailed and some of the very detailed codes don't exist in the UK code list.

For this reason, I actually saved the codes from the `icd` package as a `.csv` and went through them to remove most codes specified at 5th character detail or more, to produce a more compact code that relates to ICD-10 codes rather than ICD-10-CM.

```{r}
list_of_elixhauser_codes_icd10cm <- unclass(icd::icd10_map_quan_elix)

tbl_of_elixhauser_codes_icd10cm <- 
  enframe(list_of_elixhauser_codes_icd10cm, name = "comorbidity", value = "code") %>%
  unnest(code) %>%
  mutate(
    code = as.character(code),
    code_decimal = paste0(str_sub(code, start = 1, end = 3), ".", str_sub(code, start = 4))
    )

write.csv(x = tbl_of_elixhauser_codes_icd10cm, file = "./elixhauser_comorbidities_quan_2005_icd10cm.csv")
```

After manual review, these codes can be loaded and cleaned up...

```{r, warning=FALSE, message=FALSE}
tbl_of_elixhauser_codes_icd10 <- 
  read_csv(file = "./elixhauser_comorbidities_quan_2005_icd10_manual.csv") %>%
  rename(icd10_prefix = code) %>%
  mutate(icd10_prefix_decimal = if_else(
    condition = nchar(icd10_prefix) > 3,
    true = paste0(str_sub(icd10_prefix, start = 1, end = 3), ".", str_sub(icd10_prefix, start = 4)),
    false = icd10_prefix
  ))

# datapasta::tribble_paste(tbl_of_elixhauser_codes_icd10 %>% select(comorbidity) %>% distinct %>% mutate(comorbidity_long = " "))
long_comorbidity_names <- tibble::tribble(
                              ~comorbidity, ~comorbidity_long,
                                     "CHF",               "Congestive heart failure",
                              "Arrhythmia",               "cardiac arrhythmias",
                                "Valvular",               "Valvular disease",
                                    "PHTN",               "Pulmonary circulation disorders",
                                     "PVD",               "Peripheral vascular disorders",
                                     "HTN",               "Hypertension, uncomplicated",
                                   "HTNcx",               "Hypertension, complicated",
                               "Paralysis",               "Paralysis",
                              "NeuroOther",               "Other neurological disorders",
                               "Pulmonary",               "Chronic pulmonary disease",
                                      "DM",               "Diabetes, uncomplicated",
                                    "DMcx",               "Diabetes, complicated",
                             "Hypothyroid",               "Hypothyroidism",
                                   "Renal",               "Renal failure",
                                   "Liver",               "Liver disease",
                                     "PUD",               "Peptic ulcer disease",
                                     "HIV",               "AIDS/HIV",
                                "Lymphoma",               "Lymphoma",
                                    "Mets",               "Metastatic cancer",
                                   "Tumor",               "Solid tumour without metastasis",
                               "Rheumatic",               "Rheumatoid arthritis/collagen vascular diseases",
                            "Coagulopathy",               "Coagulopathy",
                                 "Obesity",               "Obesity",
                              "WeightLoss",               "Weight loss",
                             "FluidsLytes",               "Fluid and electrolyte disorders",
                               "BloodLoss",               "Blood loss anaemia",
                                  "Anemia",               "Deficiency anaemia",
                                 "Alcohol",               "Alcohol abuse",
                                   "Drugs",               "Drug abuse",
                               "Psychoses",               "Psychoses",
                              "Depression",               "Depression"
                            )

tbl_of_elixhauser_codes_icd10 <-
  tbl_of_elixhauser_codes_icd10 %>%
  left_join(long_comorbidity_names) %>%
  relocate(comorbidity, comorbidity_long)

## condensed version of the table
tbl_of_elixhauser_codes_icd10_condensed <-
  tbl_of_elixhauser_codes_icd10 %>%
  group_by(comorbidity, comorbidity_long) %>%
  summarise(across(.cols = matches("icd10"), .fns = ~paste0(.x, collapse = ", ")), .groups = "drop")
```

# Save tables for use/publishing

```{r}
if (!dir.exists("./comorbidity/")) dir.create("./comorbidity/")

## write the table & condensed table separately as .csv files
tbl_of_elixhauser_codes_icd10 %>% write_csv(file = "./comorbidity/elixhauser_comorbidities_quan_2005_icd10.csv")
tbl_of_elixhauser_codes_icd10_condensed %>% write_csv(file = "./comorbidity/elixhauser_comorbidities_quan_2005_icd10_condensed.csv")

descriptive_note_for_excel <- tibble(
  Notes = c("ICD-10 codes are taken from the Quan et al. 2005 paper 'Coding Algorithms for Defining Comorbidities in ICD-9-CM and ICD-10 Administrative Data', https://www.jstor.org/stable/3768193, and are based on the Elixhauser et al. 1998 paper 'Comorbidity measures for use with administrative data'.", "The codes should be treated as prefixes - any ICD-10 code matching the first 3 or 4 characters provided in this list should be included as a comorbidity of the appropriate type even if the actual diagnosis code is longer; depending on the data format, this can be implemented using the regex caret operator (^) to match the start of a string, e.g. ^E52 to capture codes starting with E52. Codes are provided both with and without a decimal comma.", "The 'icd' package for R (https://github.com/jackwasey/icd) was also used in preparing the files; the package provides a list of ICD-10-CM codes which were truncated to ICD-10 codes.")
  )

write.xlsx(x = list(
  "Description" = descriptive_note_for_excel,
  "Prefixes (long)" = tbl_of_elixhauser_codes_icd10,
  "Prefixes (condensed)" = tbl_of_elixhauser_codes_icd10_condensed
), file = "./comorbidity/elixhauser_comorbidities_quan_2005_icd10.xlsx")
```

# REFERENCES