### 2022-05-18 compile_avoidable_death_codes_2019.Rmd
Fixed issue with some of the diseases that were treatable not being labelled appropriately because I used an asterisk rather than • character

---
title: "Compiling avoidable deaths codes (as coded by NRS)"
author: "Jan Savinc"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

NRS publish an annual report on the number of avoidable/treatable deaths. The latest report is from December 2020 and reports on data from 2019, available on [the NRS website](https://www.nrscotland.gov.uk/statistics-and-data/statistics/statistics-by-theme/vital-events/deaths/avoidable-mortality).

The data used in the report can be downloaded in .csv and .xlsx format - I'll extract the code list for avoidable deaths from the excel file (.xlsx): [https://www.nrscotland.gov.uk/files//statistics/avoidable-mortality/2019/avoid-mortality-19-all-tabs.xlsx](https://www.nrscotland.gov.uk/files//statistics/avoidable-mortality/2019/avoid-mortality-19-all-tabs.xlsx)

# Load packages

```{r}
library(tidyverse)
library(readxl)  # for loading excel files
library(janitor)  # for cleaning data
library(knitr)  # for pretty tables in the report
```

# Download file

```{r}
url_avoidable_deaths_2019 <- "https://www.nrscotland.gov.uk/files//statistics/avoidable-mortality/2019/avoid-mortality-19-all-tabs.xlsx"
path_avoidable_deaths_2019 <- "./downloaded_data/avoid-mortality-19-all-tabs.xlsx"

if (!file.exists(path_avoidable_deaths_2019)) download.file(url = url_avoidable_deaths_2019, destfile = path_avoidable_deaths_2019, mode = "wb")
```

# Load definitions

The excel sheet containing definitions is helpfully named 'Definitions':

```{r}
avoidable_deaths_raw <- read_excel(path = path_avoidable_deaths_2019, sheet = "Definitions", range = "A4:E225") %>%
  clean_names
```

## Inclusions and exclusions

The column `ICD-10 codes` lists codes in the following formats:

* most codes are listed as comma-separated values (e.g. `A35, A36, A80`) - these can be easily 'split' into several rows with one value per row
* some codes are listed as ranges (using dash/hyphen, e.g. `A00-A09`) - these need to be parsed. This can be done in a 'dumb' manner without reference to the specifics of ICD-10 where not all codes in such a range exist, because invalid codes simply won't match actual data
* some codes are listed with exclusions, which are always listed in parentheses (e.g. `A40 (excl. A40.3), A41 (excl. A41.3)`). This could be somewhat complex because we might need to track which exclusion applies to which code. However, it looks like there are only 2 entries with excepetions, and there is only ever one exception per listed code or code range, which means the exclusion can simply go in a new column.
    - Note that it is possible that the same death record may contain both codes that are included under avoidable deaths and excluded. This is something to check in the actual data!

```{r}
avoidable_deaths_raw %>% filter(str_detect(string = icd_10_codes, pattern = regex("excl", ignore_case = TRUE))) %>% kable(caption = "Entries with exclusions listed")

avoidable_deaths_raw %>% filter(str_detect(string = icd_10_codes, pattern = regex("\\-", ignore_case = TRUE))) %>% kable(caption = "Entries specifying code ranges")

parse_icd_10_range <- function(code_range) {
  ## helper function for generating code ranges - based on the icd package, function .icd10cm_get_majors_possible()
  if (is.na(code_range) | !str_detect(string = code_range, pattern="\\-")) return(code_range)

  letter_1 <- str_replace(string = code_range, pattern = "^([A-Za-z]).*$", replacement = "\\1")
  letter_2 <- str_replace(string = code_range, pattern = "^.+\\-([A-Za-z]).*$", replacement = "\\1")
  letters <- LETTERS[which(LETTERS==letter_1) : which(LETTERS==letter_2)]

  decimal <- str_detect(string = code_range, pattern = "\\.")
  number_of_digits <- 2
  if (decimal) number_of_digits <- 3

  number_1 <- str_replace(string = str_remove_all(code_range, pattern="\\."), pattern = "^[A-Za-z](\\d{1,3})\\-.*$", replacement = "\\1")
  number_2 <- str_replace(string = str_remove_all(code_range, pattern="\\."), pattern = "^.+\\-[A-Za-z](\\d{1,3})$", replacement = "\\1")

  code_1 <- paste0(letter_1,number_1)
  code_2 <- paste0(letter_2,number_2)

  if (decimal) {
    candidates <- sort(apply(
        expand.grid(letters, 0:9, 0:9, 0:9),
        MARGIN = 1,
        FUN = paste0,
        collapse = ""
      ))
  } else {
    candidates <- sort(apply(
        expand.grid(letters, 0:9, 0:9),
        MARGIN = 1,
        FUN = paste0,
        collapse = ""
      ))
  }

  parsed_codes <- candidates[seq.int(
    from = which(candidates == code_1),
    to = which(candidates == code_2)
  )]
  if (decimal) parsed_codes <- paste0(str_sub(parsed_codes, 1, 3), ".", str_sub(parsed_codes, 4, 4))
  parsed_codes <- paste0(parsed_codes, collapse=",")
  return(parsed_codes)
}
```

```{r}
avoidable_deaths <-
  avoidable_deaths_raw %>%
  drop_na(condition_group_and_cause) %>%  # remove blank rows used for blank space in the excel file
  mutate(
    group = if_else(condition = is.na(icd_10_codes), true = condition_group_and_cause, false = NA_character_),
    original_codes = icd_10_codes
    ) %>%
  fill(group, .direction = "down") %>%  # forward-fill group
  drop_na(icd_10_codes) %>%
  separate_rows(icd_10_codes, sep=",\\s*", ) %>%  # separate comma-separated values into rows so we have one entry per row
  extract(col = icd_10_codes, into = "exclusions", regex = "\\(excl\\.\\s*(.+)\\)", remove = FALSE) %>%  # extract the exclusions and put them in a separate column
  mutate(
    icd_10_codes = str_remove_all(string = icd_10_codes, pattern = "\\s*\\(excl.+\\)"),  # remove the exclusions from the code ranges
    icd_10_codes = map_chr(.x = icd_10_codes, .f = ~parse_icd_10_range(.x))
  ) %>%
  separate_rows(icd_10_codes, sep=",\\s*") %>%  # separate comma-separated values again now that we've converted them from the ranges
  separate(col = age, into = c("age_min", "age_max") ,sep = "\\-", remove = TRUE) %>%
  mutate(
    treatable = case_when(  # make more explicit the treatable/preventable distinction
      str_detect(treatable, pattern = "50") ~ 0.5,
      str_detect(treatable, pattern = "\\•") ~ 1,
      TRUE ~ 0
    ),
    preventable = case_when(
      str_detect(preventable, pattern = "50") ~ 0.5,
      str_detect(preventable, pattern = "\\•") ~ 1,
      TRUE ~ 0
    ),
  ) %>%
  relocate(group, condition_group_and_cause, original_codes)
```

## Checking parsing of ranges

```{r}
avoidable_deaths_raw %>%
  drop_na(condition_group_and_cause) %>%  # remove blank rows used for blank space in the excel file
  separate_rows(icd_10_codes, sep=",\\s*") %>%  # separate comma-separated values into rows so we have one entry per row
  extract(col = icd_10_codes, into = "exclusions", regex = "\\(excl\\.\\s*(.+)\\)", remove = FALSE) %>%  # extract the exclusions and put them in a separate column
  filter(str_detect(string = icd_10_codes, pattern = "\\-")) %>%
  mutate(
    icd_10_codes = str_remove_all(string = icd_10_codes, pattern = "\\s*\\(excl.+\\)"),  # remove the exclusions from the code ranges
    icd_10_codes_parsed = map_chr(.x = icd_10_codes, .f = ~parse_icd_10_range(.x))
  ) %>%
  select(icd_10_codes, icd_10_codes_parsed) %>% 
  kable(caption = "Parsing of codes defined as ranges.")
```

# Notes on usage

When using this with real-world data, make sure to check whether the same record contains both an inclusion and an exclusion code according to this list! What to do in this case will differ from case to case.

# Save as `.csv`

```{r}
if (!dir.exists("./avoidable_deaths")) dir.create("./avoidable_deaths")
write_csv(avoidable_deaths, file = "./avoidable_deaths/avoidable_deaths_2019_inclusions.csv")
```