### 2022-07-22 covid_nrs_place_of_death_data_2022.Rmd
modified script to account for change in data - from mid-July 2022 the report no longer includes cause of death figures for 2021

---
title: "Place of death in NRS COVID-19 deaths: data processing for analysis, 2022 data"
author: "Jan Savinc"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
  code_folding: hide
toc: true
toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r, warning=FALSE, message=FALSE}
## install packages from github if not yet available!
# remotes::install_github("datasciencescotland/opendatascot", force = TRUE)
# remotes::install_github("Health-SocialCare-Scotland/phsmethods", force = TRUE)

library(tidyverse)  # for tidy workflow
library(opendatascot)  # importing data from ScotGov open data website
library(phsmethods)  # methods for working with PHS data
library(readxl)  # for reading excel files
# library(SPARQL)  # taken care of by opendatascot
library(lubridate)  # dealing with dates
library(janitor)  # for cleaning column names
library(ISOweek)  # for computing date from ISO week number + year
library(sf)  # for mapping
library(ggrepel)  # for 'mapping'repelling' labels/texst in ggplot
library(patchwork)  # for assembling plots
library(extrafont)  # for working with fonts
library(openxlsx)  # for creating xlsx files
library(knitr)  # for displaying tables in rmd

## load functions
source("./functions.R")
```

# Copyright attribution

All data used for this study were obtained from the National Records Scotland (NRS) and are © Crown copyright, 2020; the details of the licence can be viewed on the [Open Government Licence website](http://www.nationalarchives.gov.uk/doc/open-government-licence/open-government-licence.htm)

# Dates represent event registration, not occurrence!

From NRS website:

> All routine vital events information we publish is based on the date of registration, not the date on which the event occurred. Inevitably, of course, there are delays between the occurrence of an event and its registration.

# Load info about latest data

A separate script checks for updated data and downloads it, also updating these files with the locations of the downloaded data:

```{r}
file_data_sources <- "./table_of_data_sources.csv"
file_historical_data_sources <- "./table_of_historical_data_sources.csv"
file_case_trends_data_sources <- "./table_of_case_trends_data_sources.csv"

table_of_data_sources <- read_csv(file = file_data_sources)
table_of_historical_data_sources <- read_csv(file = file_historical_data_sources)
table_of_case_trends_data_sources <- read_csv(file = file_case_trends_data_sources)

table_of_data_sources %>%
  knitr::kable(caption = "Latest data used in this report.")

files_weekly_deaths <- set_names(x = as.list(table_of_data_sources$path_latest), nm = table_of_data_sources$short_name)
files_historical_deaths <- set_names(x = as.list(table_of_historical_data_sources$path_latest), nm = table_of_historical_data_sources$short_name)
files_case_counts <- set_names(x = as.list(table_of_case_trends_data_sources$path_latest), nm = table_of_case_trends_data_sources$short_name)
```

## Set up output directory

Because data are updated on a weekly basis, we'll have a separate directory for every week the data are updated:

```{r}
latest_data_modified_date <- strftime(max(table_of_data_sources$last_modified), format="%F")
latest_year <- year(latest_data_modified_date)

dir_outputs <- file.path("./outputs", latest_year, latest_data_modified_date)
if (!dir.exists(dir_outputs)) dir.create(dir_outputs, recursive = TRUE)
```

# Import & wrangle data

There are several files that the data are imported from.

## Overall data

```{r}
## a separate raw file makes it easier to debug, and also reduces the number of requests I make while debugging!
ods_raw <- ods_dataset("deaths-involving-coronavirus-covid-19", geography = "sc")

weekly_deaths_overall <-
  ods_raw %>%
  make_human_readable_ref_area %>%
  clean_names() %>%
  ## below series of rename operations are a stupid way of dealing with the possibility of different variable name conventions in the ods_dataset data: sometimes they are in snake_case, and sometimes in camelCase, seemingly unpredictably
  rename_at(.vars = vars(matches("location", ignore.case = TRUE)), ~"place_of_death") %>%
  rename_at(.vars = vars(matches("period", ignore.case = TRUE)), ~"ref_period") %>%
  rename_at(.vars = vars(matches("area", ignore.case = TRUE)), ~"ref_area") %>%
  rename_at(.vars = vars(matches("measure", ignore.case = TRUE)), ~"measure_type") %>%
  rename_at(.vars = vars(matches("cause", ignore.case = TRUE)), ~"cause_of_death") %>%
  rename_at(.vars = vars(matches("sex", ignore.case = TRUE)), ~"sex") %>%
  rename_at(.vars = vars(matches("age", ignore.case = TRUE)), ~"age") %>%
  rename_at(.vars = vars(matches("value", ignore.case = TRUE)), ~"number_of_deaths") %>%  # rename so it's consistent with other data
  filter(!ref_period %in% as.character(2020:2022)) %>% # remove totals for 2020, 2021, 2022, we only want the weekly figures
  mutate(
    year = parse_integer(str_extract(ref_period, pattern="^(202[0-3])")),
    week_number = parse_integer(str_remove(ref_period, pattern = "^(202[0-3])\\-")),
    date_w_c = compute_start_date_from_week_number(week_number = week_number, year = year),
    number_of_deaths = as.numeric(number_of_deaths),  # convert to numeric!
    week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020)  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
  ) %>%
  select(-c(measure_type)) %>%  #  measure_type is count for all data
  recode_place_of_death() %>%
  recode_cause_of_death() %>%
  recalculate_number_of_deaths_after_recoding_cause_of_death %>%  # not needed here probably, but no harm in including it
  pivot_wider(names_from = cause_of_death, values_from = number_of_deaths) %>%  # this dataset is well-formatted and has no implicit missing values!
  mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related)  # compute non-covid deaths from the other two figures - will be NA where either is missing
```

## Sex & age

```{r}
weekly_deaths_sex_age <-  # structure here is a bit more complicated - headers in row 3, but sex & age headers in row 4
  readxl::read_excel(files_weekly_deaths$sex_age, sheet = 2, skip = 4, col_names = c("week_number","place_of_death","sex","age","cause_of_death","number_of_deaths"), guess_max = 1e5) %>%
  drop_na(number_of_deaths) %>%    # remove empty rows & one row where it's just the copyright notice
  parse_year_and_week_number_from_ad_hoc_data %>%
  recode_cause_of_death() %>%
  recalculate_number_of_deaths_after_recoding_cause_of_death %>%  # recompute the sum of deaths
  tidyr::complete(  # complete implicit missing entries and make them 0
    nesting(week_number,week_number_run_over,date_w_c,year),
    age,sex,place_of_death,cause_of_death,
    fill = list(number_of_deaths = 0)
  ) %>%
  recode_place_of_death() %>%  # this needs to be done after recoding cause of death, otherwise we also include the "all" figure which isn't included in this dataset
  pivot_wider(names_from = cause_of_death, values_from = number_of_deaths) %>%   # make separate columns for causes of death figures
  mutate(deaths_all_causes = deaths_non_covid + deaths_covid_related) %>%  # compute all cause deaths from the other two figures reported
  set_covid_and_non_covid_deaths_to_zero_before_reporting_started()
```

### Sex & age separate splits

```{r}
## compute weekly data for sex and age separately
## sum over sex or age, respectively
weekly_deaths_sex <-
  weekly_deaths_sex_age %>%
  filter(age!="all") %>%
  group_by(week_number, week_number_run_over, year, date_w_c, sex, place_of_death) %>%
  summarise(across(starts_with("deaths_"), ~sum(.x)), .groups = "drop")

weekly_deaths_age <-
  weekly_deaths_sex_age %>%
  filter(sex!="all") %>%
  group_by(week_number, week_number_run_over, year, date_w_c, age, place_of_death) %>%
  summarise(across(starts_with("deaths_"), ~sum(.x)), .groups = "drop")
```

## LA & HB

```{r}
weekly_deaths_la <- 
  read_excel(files_weekly_deaths$la, sheet = 2, skip = 2) %>% 
  clean_names %>%
  rename(week_number = week_of_occurrence, ref_area = council, place_of_death = location_of_death, number_of_deaths = deaths) %>%
  filter(!is.na(number_of_deaths)) %>%  # remove blank rows & copyright entry
  parse_year_and_week_number_from_ad_hoc_data %>%
  recode_cause_of_death() %>%
  recalculate_number_of_deaths_after_recoding_cause_of_death %>%  # recompute the sum of deaths
  tidyr::complete(
      nesting(year, week_number, week_number_run_over, date_w_c),  # these are 'identifying' variables that shouldn't be combinatorily exploded
      ref_area, cause_of_death, place_of_death,  # these are the variables we want to have all combinations of, and missing combinations are set to 0 deaths
      fill = list(number_of_deaths = 0)
      ) %>%
  recode_place_of_death() %>%  # this needs to be done before recoding cause of death, otherwise we also include the "all" figure which isn't included in this dataset
  pivot_wider(names_from = cause_of_death, values_from = number_of_deaths) %>%   # make separate columns for causes of death figures
  mutate(deaths_all_causes = deaths_non_covid + deaths_covid_related) %>%  # compute all cause deaths from the other two figures reported
  set_covid_and_non_covid_deaths_to_zero_before_reporting_started()

weekly_deaths_hb <- 
  read_excel(files_weekly_deaths$hb, sheet = 2, skip = 2) %>% 
  clean_names %>%
  rename(week_number = week_of_occurrence, ref_area = health_board, place_of_death = location_of_death, number_of_deaths = deaths) %>%
  filter(!is.na(number_of_deaths)) %>%  # remove blank rows & copyright entry
  parse_year_and_week_number_from_ad_hoc_data %>%
  recode_cause_of_death() %>%
  recalculate_number_of_deaths_after_recoding_cause_of_death %>%  # recompute the sum of deaths
  tidyr::complete(
      nesting(year, week_number, week_number_run_over, date_w_c),  # these are 'identifying' variables that shouldn't be combinatorily exploded
      ref_area, cause_of_death, place_of_death,  # these are the variables we want to have all combinations of, and missing combinations are set to 0 deaths
      fill = list(number_of_deaths = 0)
      ) %>%
  recode_place_of_death() %>%  # this needs to be done before recoding cause of death, otherwise we also include the "all" figure which isn't included in this dataset
  pivot_wider(names_from = cause_of_death, values_from = number_of_deaths) %>%  # make separate columns for causes of death figures
  mutate(deaths_all_causes = deaths_non_covid + deaths_covid_related) %>%  # compute all cause deaths from the other two figures reported
  set_covid_and_non_covid_deaths_to_zero_before_reporting_started()
```

## Weekly Covid as contributory vs underlying cause of death

```{r}
weekly_covid_contributory_vs_underlying_cause <-
  readxl::read_excel(files_weekly_deaths$sex_age, sheet = 2, skip = 4, col_names = c("week_number","place_of_death","sex","age","cause_of_death","number_of_deaths"), guess_max = 1e5) %>%
  drop_na(number_of_deaths) %>%    # remove empty rows & one row where it's just the copyright notice
  parse_year_and_week_number_from_ad_hoc_data %>%
  filter(
    cause_of_death %in% c("COVID-19 contributory factor", "COVID-19 underlying cause")
  ) %>%
  tidyr::complete(  # complete implicit missing entries and make them 0
    nesting(week_number,week_number_run_over,date_w_c,year),
    age,sex,place_of_death,cause_of_death,
    fill = list(number_of_deaths = 0)
  ) %>%
  recode_place_of_death() %>%  # this needs to be done after recoding cause of death, otherwise we also include the "all" figure which isn't included in this dataset
  pivot_wider(names_from = cause_of_death, values_from = number_of_deaths) %>%   # make separate columns for causes of death figures
  clean_names() %>%
  filter(age!="all") %>%
  filter(sex!="all") %>%
  filter(place_of_death != "All") %>%
  group_by(week_number, week_number_run_over, year, date_w_c, place_of_death) %>%  # sum over age & sex
  summarise(across(starts_with("covid_"), ~sum(.x)), .groups = "drop") %>%
  ({function(x)
    bind_rows(
      x,
      x %>% 
        group_by(week_number, week_number_run_over, year, date_w_c) %>%  # sum across place of death
        summarise(across(starts_with("covid_"), ~sum(.x)), .groups = "drop") %>%
        mutate(place_of_death = "All")
    )
  })

## check that the figures actually make sense! i.e. the two figures add up to the total covid-related figure
stopifnot(
  (weekly_deaths_overall %>%
  filter(sex=="all" & age=="all") %>%
  inner_join(weekly_covid_contributory_vs_underlying_cause) %>%
  filter(deaths_covid_related != covid_19_contributory_factor + covid_19_underlying_cause) %>%
  nrow) == 0
)

annual_covid_contributory_vs_underlying_cause <-
  weekly_covid_contributory_vs_underlying_cause %>%
  group_by(year, place_of_death) %>%
  summarise(across(starts_with("covid_"), ~sum(.x)), .groups = "drop") %>%
  ({function(x)
    bind_rows(
      x %>% mutate(year = as.character(year)),
      x %>% 
        group_by(place_of_death) %>%  # sum across eyars
        summarise(across(starts_with("covid_"), ~sum(.x)), .groups = "drop") %>%
        mutate(year = "2020-2021")
    )
  }) %>%
  mutate(deaths_covid_related = covid_19_contributory_factor + covid_19_underlying_cause) %>%
  mutate(
    proportion_contributory = covid_19_contributory_factor / deaths_covid_related,
    proportion_underlying = covid_19_underlying_cause / deaths_covid_related
    )
```

## 'Main report' file

The format of the main report file has changed from what was used in 2021.

However, the main report file is still only being used for cause of death data, which simplifies parsing the file!

Causes of death are reported in `sheet 8` of the report, ~~and include figure from 2021 and 2022. The worksheet contains data split by location of death, which are contained within different tables arranged vertically, so their boundaries need to be written out; the tables have been pre-filled for all of 2021 and 2022, so some proportion of entries will be blank for 2022.~~

Note: the above no longer holds as of 20 July 2022 - only 2022 data re included in `sheet 8` now, which means the 2021 data need to be imported from an older file.

Note that excess deaths and 5-year average deaths are computed differently between 2020-2021 and 2022:

> For data registered in 2021, any comparison to a five year average is being made with the 2015-2019 average. For 2022, the comparison is being made to the average of 2016, 2017, 2018, 2019, 2021.

> Excess deaths are calculated by comparing the current year to the five year average from previous years. NRS, along with ONS and NISRA, agreed to leave 2020 out of five year average calculations so the five-year average which is used to compare 2022 figures against covers the years 2016, 2017, 2018, 2019 and 2021. More information about this can be found in a paper published on our website (see link on the right). Moveable public holidays, when registration offices are closed, affect the number of registrations made in the published weeks and in the corresponding weeks in previous years.

### Weekly causes of death

```{r}
sheet_cause_of_death_data <- "8"

# TODO: delete below, it's for the previous releases that included 2021 data also
# cell_ranges_for_causes_of_death_data <- list(
#   "All" = "A6:V110",
#   "Care home" = "A113:V217",
#   "Home & other non-institution" = "A220:V324",
#   "Hospital" = "A327:V431",
#   "Other" = "A434:V538"
# )

cell_ranges_for_causes_of_death_data <- list(
  "All" = "A6:V58",
  "Care home" = "A61:V113",
  "Home & other non-institution" = "A116:V168",
  "Hospital" = "A171:V223",
  "Other" = "A226:V278"
)

weekly_deaths_cause_of_death_2022_onwards <-
  map_dfr(
    .x = cell_ranges_for_causes_of_death_data,
    .f = ~readxl::read_excel(files_weekly_deaths$main_report, sheet = sheet_cause_of_death_data, range = .x), 
    .id = "place_of_death"
  ) %>%
  janitor::clean_names() %>%
  rename(year = registration_year, date_w_c = week_beginning) %>%
  select(-matches("_excess$|_average$")) %>%  # we can compute excess & averages later
  pivot_longer(cols = 5:ncol(.), names_to = "name", values_to = "number_of_deaths") %>%
  filter(!str_detect(name, "all_causes")) %>%  # we can calculate that from the others
  mutate(
    year = as.integer(year),
    cause_of_death = str_remove_all(name, "_five_year_average|_deaths"),
    cause_of_death = return_pretty_cause_of_death(cause_of_death),
    date_w_c = as.Date(date_w_c, origin = "1899-12-30"),
    week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020)  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
  ) %>%
  select(-name) %>%
  filter(!is.na(number_of_deaths))  # remove blank entries - these are for the current year but for dates in the future
```

### Data from 2020 & 2021

The reports available from NRS in 2022 include data from 2021-2022 up until week 25, and partly in weeks 26 and 27 where there is data from 2021 for specific places of death but not for all places of death. I also include data from 2020 for context. For this I'll load the last 'main report' file from 2021.

The weekly causes of death data for 2020 is included in the sheet `Table 3 (2020)`.

The weekly causes of death data for 2021 is included in the week 25 'main report' file, table `8`.

#### 2020

```{r}
# this uses the combined 2020-2021 file
# nrs_weekly_data_sheet_2020 <- 
#   excel_sheets(files_historical_deaths$`2020-2021`)[which(str_detect(tolower(excel_sheets(files_historical_deaths$`2020-2021`)), pattern = "table 3.*2020"))]

nrs_weekly_data_sheet_2020 <- 
  excel_sheets(files_historical_deaths$`2020`)[which(str_detect(tolower(excel_sheets(files_historical_deaths$`2020`)), pattern = "table 3.*2020"))]

## construct column names for the data
headers_in_nrs_data_2020 <-
  c(
    "place_of_death",
    "cause_of_death",
    paste0("week_", 1:53),
    "blank_remove",  # this and the rest are irrelevant entries that we'll remove
    "remove_1",
    "remove_2"
  )

## extract data... this is dense but I've tried to annotate it clearly!
weekly_deaths_cause_of_death_2020 <-
  readxl::read_excel(files_historical_deaths$`2020`, sheet = nrs_weekly_data_sheet_2020, col_names = headers_in_nrs_data_2020, range = cell_rows(6:125)) %>%  # the data are inrows 6:125
  select(-matches("remove")) %>%  # remove irrelevant summary columns at the end
  mutate(across(.cols = matches("week"), .fns = as.numeric)) %>%  # remove erroneous "count1, count2, ..." entries due to superscript in previous column
  mutate(dataset = if_else(condition = str_detect(string = cause_of_death, pattern = regex(pattern = "^registered|^difference", ignore_case = TRUE)), true = cause_of_death, false = NA_character_)) %>%  # extract the Registered deaths: historical, 2020/2021, and Difference sections
  tidyr::fill(place_of_death, .direction = "down") %>%  # complete the entries for place of death from the "last observed"
  tidyr::fill(dataset, .direction = "down") %>%  # same as above but for "dataset" column
  drop_na(week_1) %>%  # remove blank "header" rows with no data recorded
  pivot_longer(cols = matches("week"), names_to = "week_number", values_to = "number_of_deaths", values_drop_na = TRUE) %>%  # convert to long data format
  mutate(
    week_number = as.integer(parse_number(week_number)),  # extract week number
    place_of_death =  # find best match in the reference names used in other datasets
      order_of_place_of_death_levels[stringdist::amatch(x = str_sub(place_of_death,1,4), table = str_sub(order_of_place_of_death_levels,1,4), method = "lv", maxDist = 5)],  # the reference is in the variable order_of_place_of_death_levels
    year = 2020,
    dataset = str_remove_all(dataset , "Registered deaths\\: "),
    date_w_c = compute_start_date_from_week_number(week_number = week_number, year = year),
    week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020)  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
    ) %>%
  filter(dataset!="Difference") %>%
  pivot_wider(names_from = dataset, values_from = number_of_deaths) %>%
  rename(deaths_mean = `five year average`, number_of_deaths = `2020`)
```

#### 2021

```{r}
cell_ranges_for_causes_of_death_data_2021 <- list(
  "All" = "A6:V58",
  "Care home" = "A113:V165",
  "Home & other non-institution" = "A220:V272",
  "Hospital" = "A327:V379",
  "Other" = "A434:V486"
)

weekly_deaths_cause_of_death_2021 <-
  map_dfr(
    .x = cell_ranges_for_causes_of_death_data_2021,
    .f = ~readxl::read_excel(files_historical_deaths$`2021`, sheet = sheet_cause_of_death_data, range = .x), 
    .id = "place_of_death"
  ) %>%
  janitor::clean_names() %>%
  rename(year = registration_year, date_w_c = week_beginning) %>%
  select(-matches("_excess$|_average$")) %>%  # we can compute excess & averages later
  pivot_longer(cols = 5:ncol(.), names_to = "name", values_to = "number_of_deaths") %>%
  filter(!str_detect(name, "all_causes")) %>%  # we can calculate that from the others
  mutate(
    year = as.integer(year),
    cause_of_death = str_remove_all(name, "_five_year_average|_deaths"),
    cause_of_death = return_pretty_cause_of_death(cause_of_death),
    date_w_c = as.Date(date_w_c, origin = "1899-12-30"),  # origin to account for Excel's integer date format
    week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020)  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
  ) %>%
  select(-name) %>%
  filter(!is.na(number_of_deaths))  # remove blank entries - these are for the current year but for dates in the future
```

#### Historical causes of death

Historical here refers to the 2015-2019 period for 2020 & 2021, and 2016-2021 (skipping 2020) for 2022, useful for a comparison with the historical context of the pre-pandemic years. (There is another 'historical' dataset, which includes cause of death figures for 2020 and for 2021.)

In the `2015-2019` period there were no Covid-19 deaths, so the historical average should be `NA`, which can possibly be interpreted as 0 for the purposes of comparing a year to the historical average.
For the `2016-2019, 2021` period the average number of covid deaths should be equivalent to the deaths in 2021 - we wouldn't want to make an average between 4 years of 0 deaths and one year with any deaths! Similarly if the period were `2017-2019, 2021-2022`, we would compute the average for 2021-2022, and so on. The year 2020 is considered exceptional and so is not used for computing the historical averages in 2021 and onwards.

Because the data from July 2022 onwards are split up so that I load the 2020 and 2021 from different sources, I'll compute the

```{r}
## this is an ugly hack - we set average weekly deaths due to C19 to NA for 2021
## and we set the average covid deaths for 2015-2019, 2021 to the same as in 2021
## however, for the latter, we set the dates to 2022
# average_weekly_covid_deaths_2021_onwards <-
average_weekly_covid_deaths_2021_onwards <-
  bind_rows(
    ## set 2021 averages to NA
    weekly_deaths_cause_of_death_2021 %>%
    filter(cause_of_death == "COVID-19") %>%
    mutate(number_of_deaths = NA_real_) %>%
    rename(deaths_mean = number_of_deaths),
    ## use 2021 weekly deaths as 2022 historic average
    weekly_deaths_cause_of_death_2021 %>%
    filter(cause_of_death == "COVID-19") %>%
    mutate(year = 2022) %>%
    select(-c(week_number_run_over, date_w_c)) %>%
    rename(deaths_mean = number_of_deaths) %>%
    inner_join(weekly_deaths_cause_of_death_2022_onwards %>% 
                select(year, week_number, week_number_run_over, date_w_c, place_of_death, cause_of_death)
              )
  ) %>%
  mutate(historic_period = historic_period_depending_on_year(year))

average_weekly_noncovid_deaths_2021_onwards <-
    ## extract 5-year averages from most recent file for 2021 onwards
    ## this doesn't include covid, which we've manually extracted above!
  bind_rows(
    ## 2022+ data
    map_dfr(
      .x = cell_ranges_for_causes_of_death_data,
      .f = ~readxl::read_excel(files_weekly_deaths$main_report, sheet = sheet_cause_of_death_data, range = .x), 
      .id = "place_of_death"
    ),
    ## 2021 data
    map_dfr(
      .x = cell_ranges_for_causes_of_death_data_2021,
      .f = ~readxl::read_excel(files_historical_deaths$`2021`, sheet = sheet_cause_of_death_data, range = .x), 
      .id = "place_of_death"
    )
  ) %>%
    janitor::clean_names() %>%
    rename(year = registration_year, date_w_c = week_beginning) %>%
    select(place_of_death, year, week_number, date_w_c, matches("five_year_average")) %>%
    pivot_longer(cols = 5:ncol(.), names_to = "name", values_to = "deaths_mean") %>%
    mutate(
      year = as.integer(year),
      cause_of_death = str_remove_all(name, "_five_year_average|_deaths"),
      cause_of_death = return_pretty_cause_of_death(cause_of_death),
      cause_of_death = if_else(cause_of_death == "All causes", "All", cause_of_death),
      date_w_c = as.Date(date_w_c, origin = "1899-12-30"),
      week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020),  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
      historic_period = historic_period_depending_on_year(year)
    ) %>%
    select(-name) %>%
    drop_na(deaths_mean)  # remove blank entries - these are for the current year but for dates in the future

historical_average_cause_of_death_2021_onwards <-
  bind_rows(
    ## the hacked averages for 2021 and 2022 above
    average_weekly_covid_deaths_2021_onwards,
    ## the non-covid deaths in the 'main report' files
    average_weekly_noncovid_deaths_2021_onwards
  )
```

#### Merge 2020 & 2021 data with current data

Note: the 5-year average for 2022 (spanning 2016-2019 & 2021) is only published along with the running weekly counts of deaths by cause of death; it's just that NRS has decided to only publish them at the same time as new data comes out! This is a bit of a bummer - it means we can't populate the whole year of 'historical context' data for 2022.

They could be calculated by hand if the 2015-2019 weekly data were available broken down by cause of death - as it stands, only the average 2015-2019 figure is available.

```{r}
weekly_deaths_cause_of_death <-
  bind_rows(
    weekly_deaths_cause_of_death_2022_onwards,
    weekly_deaths_cause_of_death_2021,
    weekly_deaths_cause_of_death_2020 %>% 
      select(-deaths_mean) %>%
      filter(cause_of_death != "All")  # don't need all cause data in causes breadkown!
  )

historical_average_cause_of_death <-
  bind_rows(
    historical_average_cause_of_death_2021_onwards,
    weekly_deaths_cause_of_death_2020 %>% 
      select(-number_of_deaths) %>% 
      mutate(historic_period = historic_period_depending_on_year(year))
  )
```

## Historical weekly deaths

The format of the .xlsx files is that sheet 1 is a table of contents, sheet 2 contains the data, and sheet 3 is a pivot table for subsetting data. These are downloaded from the ad-hoc section also.

```{r, warning=FALSE}
weekly_deaths_2015_2019_overall <- 
  read_csv(files_historical_deaths$overall, skip = 2) %>%
  rename(year = 1) %>%  # rename first column
  mutate(
    place_of_death = if_else(is.na(`1`), true = year, NA_character_)
  ) %>%
  fill(place_of_death, .direction="down") %>%  # LOCF on place_of_death
  drop_na(`1`) %>%
  pivot_longer(cols = all_of(as.character(1:53)), values_to = "number_of_deaths", names_to = "week_number") %>%
  mutate(week_number=as.integer(week_number)) %>%
  recode_place_of_death()

weekly_deaths_2015_2019_hb <- 
  readxl::read_excel(files_historical_deaths$hb, sheet = 2, skip = 2, col_names = TRUE) %>%  # first two rows are empty, row 3 has headers
  clean_names() %>%
  drop_na(number_of_deaths) %>%  # remove empty rows & one row where it's just the copyright notice
  rename(  # same format as current data
    week_number = matches("week"),
    place_of_death = matches("location"),
    ref_area = health_board
  ) %>%
  mutate(
    week_number = as.integer(week_number)
  ) %>%
  tidyr::complete(
    nesting(week_number,year),  # unique combinations that we don't want to combinatorily explode
    ref_area, place_of_death,  # fill in missing combinations of hb & pod with 0 deaths
    fill = list(number_of_deaths = 0)
  ) %>% 
  recode_place_of_death()  # recode place of death after completing missing combinations, otherwise the factor level "All" gets exploded also

weekly_deaths_2015_2019_la <-
  readxl::read_excel(files_historical_deaths$la, sheet = 2, skip = 2, col_names = TRUE) %>%  # first two rows are empty, row 3 has headers
  clean_names() %>%
  drop_na(number_of_deaths) %>%  # remove empty rows & one row where it's just the copyright notice
  rename(  # same format as current data
    week_number = matches("week"),
    place_of_death = matches("location"),
    ref_area = council_area
  ) %>%
  mutate(
    week_number = as.integer(week_number)
  ) %>%
  tidyr::complete(
    nesting(week_number,year),  # unique combinations that we don't want to combinatorily explode
    ref_area, place_of_death,  # fill in missing combinations of la & pod with 0 deaths
    fill = list(number_of_deaths = 0)
  ) %>% 
  recode_place_of_death()  # recode place of death after completing missing combinations, otherwise the factor level "All" gets exploded also

weekly_deaths_2015_2019_sex_age <-  # structure here is a bit more complicated - headers in row 3, but sex & age headers in row 4
  readxl::read_excel(files_historical_deaths$sex_age, sheet = 2, skip = 4, col_names = c("year","week_number","place_of_death","sex","age","number_of_deaths"), guess_max = 1e5) %>%
  drop_na(number_of_deaths) %>%    # remove empty rows & one row where it's just the copyright notice
  mutate(
    year = 2000L + as.integer(year),
    week_number = as.integer(week_number)
  ) %>%
  tidyr::complete(
    nesting(week_number,year),  # unique combinations that we don't want to combinatorily explode
    sex, age, place_of_death,  # fill in missing combinations of age, sex, pod, with 0 deaths
    fill = list(number_of_deaths = 0)
  ) %>% 
  recode_place_of_death()

## compute weekly data for sex and age separately
## sum over sex or age, respectively
weekly_deaths_2015_2019_sex <-
  weekly_deaths_2015_2019_sex_age %>%
  group_by_all %>%
  ungroup(age, number_of_deaths) %>%  # this is done by grouping over everything except the number of deaths & sex
  summarise(number_of_deaths = sum(number_of_deaths), .groups = "drop")

weekly_deaths_2015_2019_age <-
  weekly_deaths_2015_2019_sex_age %>%
  group_by_all %>%
  ungroup(sex, number_of_deaths) %>%  # this is done by grouping over everything except the number of deaths & sex
  summarise(number_of_deaths = sum(number_of_deaths), .groups = "drop")
```

## Map data

I'm using `sf` to read these and to include them invisualisations.

* Local Authority Districts (LAD) for 2020 (BUC-level resolution): https://geoportal.statistics.gov.uk/datasets/local-authority-districts-december-2020-uk-buc/
* Health Boards https://data.gov.uk/dataset/27d0fe5f-79bb-4116-aec9-a8e565ff756a/nhs-health-boards
  - Note: this is a fairly high resolution, don't need this much detail!
  - I found a discussion of this on [a github covid-related project](https://github.com/tomwhite/covid-19-uk-data/issues/18#issuecomment-622353841), and user **robchallen** produced a zip file containing [UK-wide health board boundaries](https://github.com/tomwhite/covid-19-uk-data/files/4563933/UK_covid_reporting_regions.zip)

```{r}
shapefile_hb_uk_wide <- st_read(dsn = "./map_data/hb_uk_wide/UK_covid_reporting_regions")
shapefile_la <- st_read(dsn = "./map_data/la")
```

## COVID case counts by LA/HB

```{r}
cases_hb <- read_csv(
  files_case_counts$hb, 
  col_types = 
    cols(
    .default = col_double(),
    `_id` = col_skip(),
    DailyDeaths = col_skip(),  # not needed, we get this elsewhere!
    Date = col_date(format = "%Y%m%d"),
    HB = col_character(),
    HBName = col_character(),
    HospitalAdmissionsQF = col_character(),
    ICUAdmissionsQF = col_character()
    ),
  ) %>% clean_names() %>%
  rename(ref_area = hb_name) %>%
  mutate(
    year = parse_integer(str_sub(ISOweek(date), start = 1, end = 4)),  # parse ISO year, otherwise the first few days of 2021 will belong to 2021 erroneously!
    week_number = parse_integer(str_sub(ISOweek(date), start = 7, end = 8)),
    ref_area = phsmethods::match_area(hb)  # convert to standard HB names
    )

cases_la <- read_csv(
  file = files_case_counts$la, 
  col_types = 
    cols(
    .default = col_double(),
    `_id` = col_skip(),
    DailyDeaths = col_skip(),  # not needed, we get this elsewhere!
    Date = col_date(format = "%Y%m%d"),
    CA = col_character(),
    CAName = col_character()
  )
) %>% clean_names() %>%
  rename(ref_area = ca_name) %>%
  mutate(
    year = parse_integer(str_sub(ISOweek(date), start = 1, end = 4)),  # parse ISO year, otherwise the first few days of 2021 will belong to 2021 erroneously!
    week_number = parse_integer(str_sub(ISOweek(date), start = 7, end = 8)),
    ref_area = phsmethods::match_area(ca)  # convert to standard LA names, jsut in case they weren't already!
    )
```

# Historical averages, minima & maxima

Due to the counting of weeks, most years have 52 weeks with some years (2015 & 2020) having 53. Therefore we need to decide how to deal with week 53 in 2020 since only 2015 has a corresponding weekly count. The solution used here is to calculate a weekly rate for week 53 from week 53 in 2015, and week 52 in 2016-2019 - this represents the "last week of the year", which includes December 31st.

Note also that since 2022, the 5-year historical period has been defined as 2016-2021, with 2020 excluded, so we need to calculate two weekly averages!

```{r, warning=FALSE}
historical_weekly_average_overall <-
  bind_rows(
    ## 2015-2019 period
    weekly_deaths_2015_2019_overall %>%
    filter(year!="average") %>%  # remove averages
    group_by(year, week_number) %>%  # remove week 53 if it was 0 for all places  - this means there was no week 53!
    filter(!all(number_of_deaths==0 & week_number==53)) %>%
    ungroup %>%
    create_week_53_data_for_years_with_52_weeks %>%
    group_by(week_number,place_of_death) %>% 
    summarise(
      across(number_of_deaths, .fns = list(mean=mean,min=min,max=max)),
      .groups = "drop"
    ) %>% 
    rename_all(~str_replace_all(.x, pattern = "number_of_deaths", replacement = "deaths")) %>%
    mutate(historic_period = historic_period_depending_on_year(2020))
    ,
    ## 2016-2021, excluding 2020
    bind_rows(
      ## 2016-2019
      weekly_deaths_2015_2019_overall %>%
        mutate(year = as.integer(year)) %>%
        filter(year %in% 2016:2019),
      ## 2021
      weekly_deaths_overall %>% 
        rename(number_of_deaths = deaths_all_causes) %>%
        filter(year == 2021 & age == "all" & sex == "all")
      ) %>%
    group_by(year, week_number) %>%  # remove week 53 if it was 0 for all places  - this means there was no week 53!
    filter(!all(number_of_deaths==0 & week_number==53)) %>%
    ungroup %>%
    create_week_53_data_for_years_with_52_weeks %>%
    group_by(week_number,place_of_death) %>% 
    summarise(
      across(number_of_deaths, .fns = list(mean=mean,min=min,max=max)),
      .groups = "drop"
    ) %>% 
    rename_all(~str_replace_all(.x, pattern = "number_of_deaths", replacement = "deaths")) %>%
    mutate(historic_period = historic_period_depending_on_year(2022))
  )

week_53_text <- "Due to the way ISO 8601 counts weeks, only 2015 in the 2015-2019 period had 53 weeks; averages for week 53 were computed from figures for week 53 in 2015 and week 52 for 2016-2019."

# TODO: weekly averages may be too detailed for this level od subcategorising!

deaths_mean_min_max <- function(data_tbl) {
  data_tbl %>%
    summarise(
      deaths_mean = mean(number_of_deaths),
      deaths_min = min(number_of_deaths),
      deaths_max = max(number_of_deaths),
      .groups = "drop"
    )
}

make_2016_to_2019_and_2021_data <- function(historical_data, current_data) {
  bind_rows(
      historical_data %>% filter(year %in% 2016:2019),
      current_data %>% filter(year == 2021) %>% rename(number_of_deaths = deaths_all_causes)
    )
}

historical_weekly_average_hb <-
  bind_rows(
    ## 2015-2019
    weekly_deaths_2015_2019_hb %>%
    create_week_53_data_for_years_with_52_weeks() %>%
    group_by(
      week_number, place_of_death, ref_area
    ) %>%
    deaths_mean_min_max %>%
    mutate(historic_period = historic_period_depending_on_year(2020))
    ,
    ## 2016-2019, 2021
    make_2016_to_2019_and_2021_data(
      historical_data = weekly_deaths_2015_2019_hb, 
      current_data = weekly_deaths_hb
      ) %>%
    create_week_53_data_for_years_with_52_weeks() %>%
    group_by(
      week_number, place_of_death, ref_area
    ) %>%
    deaths_mean_min_max %>%
    mutate(historic_period = historic_period_depending_on_year(2022))
  )

historical_weekly_average_la <-
  bind_rows(
    ## 2015-2019
    weekly_deaths_2015_2019_la %>%
    create_week_53_data_for_years_with_52_weeks() %>%
    group_by(
      week_number, place_of_death, ref_area
    ) %>%
    deaths_mean_min_max %>%
    mutate(historic_period = historic_period_depending_on_year(2020))
    ,
    ## 2016-2019, 2021
    make_2016_to_2019_and_2021_data(
      historical_data = weekly_deaths_2015_2019_la, 
      current_data = weekly_deaths_la
      ) %>%
    create_week_53_data_for_years_with_52_weeks() %>%
    group_by(
      week_number, place_of_death, ref_area
    ) %>%
    deaths_mean_min_max %>%
    mutate(historic_period = historic_period_depending_on_year(2022))
  )

historical_weekly_average_sex_age <-
  bind_rows(
    ## 2015-2019
    weekly_deaths_2015_2019_sex_age %>%
    create_week_53_data_for_years_with_52_weeks() %>%
    group_by(
      week_number, place_of_death, age, sex
    ) %>%
    deaths_mean_min_max %>%
    mutate(historic_period = historic_period_depending_on_year(2020))
    ,
    ## 2016-2019, 2021
    make_2016_to_2019_and_2021_data(
      historical_data = weekly_deaths_2015_2019_sex_age, 
      current_data = weekly_deaths_sex_age
      ) %>%
    create_week_53_data_for_years_with_52_weeks() %>%
    group_by(
      week_number, place_of_death, age, sex
    ) %>%
    deaths_mean_min_max %>%
    mutate(historic_period = historic_period_depending_on_year(2022))
  )

historical_weekly_average_age <-
  bind_rows(
    ## 2015-2019
    weekly_deaths_2015_2019_age %>%
    create_week_53_data_for_years_with_52_weeks() %>%
    group_by(
      week_number, place_of_death, age
    ) %>%
    deaths_mean_min_max %>%
    mutate(historic_period = historic_period_depending_on_year(2020))
    ,
    ## 2016-2019, 2021
    make_2016_to_2019_and_2021_data(
      historical_data = weekly_deaths_2015_2019_age, 
      current_data = weekly_deaths_age
      ) %>%
    create_week_53_data_for_years_with_52_weeks() %>%
    group_by(
      week_number, place_of_death, age
    ) %>%
    deaths_mean_min_max %>%
    mutate(historic_period = historic_period_depending_on_year(2022))
  )

historical_weekly_average_sex <-
  bind_rows(
    ## 2015-2019
    weekly_deaths_2015_2019_sex %>%
    create_week_53_data_for_years_with_52_weeks() %>%
    group_by(
      week_number, place_of_death, sex
    ) %>%
    deaths_mean_min_max %>%
    mutate(historic_period = historic_period_depending_on_year(2020))
    ,
    ## 2016-2019, 2021
    make_2016_to_2019_and_2021_data(
      historical_data = weekly_deaths_2015_2019_sex, 
      current_data = weekly_deaths_sex
      ) %>%
    create_week_53_data_for_years_with_52_weeks() %>%
    group_by(
      week_number, place_of_death, sex
    ) %>%
    deaths_mean_min_max %>%
    mutate(historic_period = historic_period_depending_on_year(2022))
  )

# TODO: include the split between historical periods here also
compute_historical_annual_average_deaths <- function(data_tbl, ...) {
  grouping_vars <- enquos(...)
  data_tbl %>%
    group_by(year, place_of_death, historic_period, !!!grouping_vars) %>%
    summarise(  # calculate annual totals first...
    number_of_deaths = sum(number_of_deaths), 
    .groups = "keep"
  ) %>%
  ungroup(year) %>%  # remove year as grouping to average over year now
  summarise(
    annual_deaths_mean = mean(number_of_deaths),
    annual_deaths_min = min(number_of_deaths),
    annual_deaths_max = max(number_of_deaths),
    .groups = "drop"
  )
}

historical_annual_average_overall <-
  bind_rows(
      ## 2016-2019
      weekly_deaths_2015_2019_overall %>%
        mutate(year = as.integer(year)) %>%
        filter(year %in% 2016:2019) %>%
        mutate(historic_period = historic_period_depending_on_year(2020)),
      ## 2021
      weekly_deaths_overall %>% 
        rename(number_of_deaths = deaths_all_causes) %>%
        filter(year == 2021 & age == "all" & sex == "all") %>%
        mutate(historic_period = historic_period_depending_on_year(2022))
      ) %>%
  compute_historical_annual_average_deaths()

make_2016_to_2019_and_2021_data_with_historic_periods <- function (historical_data, current_data) {
  bind_rows(
    historical_data %>% 
      filter(year %in% 2016:2019) %>%
      mutate(historic_period = historic_period_depending_on_year(2020)),
    current_data %>% 
      filter(year == 2021) %>% 
      rename(number_of_deaths = deaths_all_causes) %>%
      mutate(historic_period = historic_period_depending_on_year(2022))
  )
}

historical_annual_average_hb <-
  make_2016_to_2019_and_2021_data_with_historic_periods(
    historical_data = weekly_deaths_2015_2019_hb, 
    current_data = weekly_deaths_hb
      ) %>%
  compute_historical_annual_average_deaths(ref_area)

historical_annual_average_la <-
  make_2016_to_2019_and_2021_data_with_historic_periods(
    historical_data = weekly_deaths_2015_2019_la, 
    current_data = weekly_deaths_la
      ) %>%
  compute_historical_annual_average_deaths(ref_area)

historical_annual_average_sex_age <-
  make_2016_to_2019_and_2021_data_with_historic_periods(
    historical_data = weekly_deaths_2015_2019_sex_age, 
    current_data = weekly_deaths_sex_age
      ) %>%
  compute_historical_annual_average_deaths(sex, age)

historical_annual_average_sex <-
  make_2016_to_2019_and_2021_data_with_historic_periods(
    historical_data = weekly_deaths_2015_2019_sex, 
    current_data = weekly_deaths_sex
      ) %>%
  compute_historical_annual_average_deaths(sex)

historical_annual_average_age <-
  make_2016_to_2019_and_2021_data_with_historic_periods(
    historical_data = weekly_deaths_2015_2019_age, 
    current_data = weekly_deaths_age
      ) %>%
  compute_historical_annual_average_deaths(age)
```

## Historical annual average by cause

I can approximate this by taking the sum of the average weekly deaths for `2015-2019`; for the `2016-2019 + 2021` period this isn't readily available because NRS only publish the weekly averages alongside the weekly updated data, but not for the entire year. Therefore I'll just calculate the `2015-2019` period average for now, and a compromise `2015-2019 + 2021` average:

$$\frac{5}{6}average\_deaths_{2015-2019} + \frac{1}{6}deaths_{2021}$$

```{r}
# TODO: incorporate the 2016-2019 + 2021 period
historical_annual_average_cause_of_death_2015_2019 <-
  historical_average_cause_of_death %>%
  filter(historic_period == "2015-2019" & year == 2021) %>%  # this is slightly strange but some of the averages disagree, probably because the 2015-2019 data were updated in 2020, and the 2021 data were published with modified figures. NRS deaths data are truly a living document!
  group_by(place_of_death, cause_of_death, historic_period) %>%
  summarise(
    annual_deaths_mean = sum(deaths_mean),
    .groups = "drop"
  )

deaths_by_cause_2021 <-
  weekly_deaths_cause_of_death_2021 %>% 
  group_by(place_of_death, cause_of_death) %>% 
  summarise(deaths_2021 = sum(number_of_deaths), .groups = "drop")

historical_annual_average_cause_of_death <-
  bind_rows(
    historical_annual_average_cause_of_death_2015_2019
    ,
    ## compromise figures for 2015-2019 + 2021
    historical_annual_average_cause_of_death_2015_2019 %>%
    filter(cause_of_death != "All") %>%
    left_join(
      deaths_by_cause_2021, by = c("place_of_death", "cause_of_death")
    ) %>%
    mutate(
      annual_deaths_mean = (5/6 * annual_deaths_mean) + (1/6 * deaths_2021),
      deaths_2021 = NULL,
      historic_period = "2015-2019, 2021"
    )
  )
```

## Check weekly historic averages against NRS computed data

```{r}
weekly_deaths_overall %>%
  filter(sex=="all" & age=="all") %>%
  mutate(historic_period = historic_period_depending_on_year(year)) %>%
  left_join(historical_weekly_average_overall) %>%
  select(year, week_number, place_of_death, deaths_nrs_past_average_all_causes, deaths_mean, historic_period) %>%
  filter(round(deaths_nrs_past_average_all_causes) != round(deaths_mean)) %>%
  knitr::kable(caption = "Subset of weekly deaths where calculated mean deaths 2015-2019 differed between the COVID-19 related deaths file and the weekly deaths file (i.e. all other entries had the same figures!). Week 53 was computed as the average of deaths in week 53 in 2015 and weeks 52 in 2016-2019. For all the other datasets the mean deaths computed agree!")
```

# Merge weekly deaths with historic death rates

```{r}
## compute rolling average
merged_deaths_overall <-
  weekly_deaths_overall %>%
  filter(sex=="all" & age=="all") %>%
  mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
  inner_join(historical_weekly_average_overall %>% mutate(sex = "all", age = "all"),  # only merge for aggregated data b/c that's all we have!
  by = c("place_of_death", "week_number", "sex", "age", "historic_period"))

merged_deaths_hb <-
  weekly_deaths_hb %>%
  mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
  inner_join(historical_weekly_average_hb, by = c("week_number","ref_area","place_of_death", "historic_period"))

merged_deaths_la <-
  weekly_deaths_la %>%  # use data with weekly movingincluded
  mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
  inner_join(historical_weekly_average_la, by = c("week_number","ref_area","place_of_death", "historic_period"))

merged_deaths_sex_age <-
  weekly_deaths_sex_age %>%  # use data with weekly movingincluded
  mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
  inner_join(historical_weekly_average_sex_age, by = c("place_of_death", "week_number", "sex", "age", "historic_period"))

merged_deaths_sex <-
  weekly_deaths_sex %>%
  mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
  inner_join(historical_weekly_average_sex, by = c("place_of_death", "week_number", "sex", "historic_period"))

merged_deaths_age <-
  weekly_deaths_age %>%
  mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
  inner_join(historical_weekly_average_age, by = c("place_of_death", "week_number", "age", "historic_period"))

merged_deaths_cause_of_death <-
  weekly_deaths_cause_of_death %>%
  mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
  left_join(  # left_join so that the historical Covid-19 rates are blank!
    historical_average_cause_of_death %>% select(-week_number_run_over), 
    by = c("place_of_death", "week_number", "cause_of_death", "historic_period", "year", "date_w_c")
  )
```

## Weekly deaths including past weekly deaths

For this dataset, we simply `bind_rows` weekly deaths by year and place of death with past weekly deaths by year and location. Note that past data for years with 52 weeks includes an entry of 0 deaths for week 53, so we'll remove them, otherwise the date computation is thrown off!

```{r}
weekly_deaths_current_and_past <-
  bind_rows(
    weekly_deaths_overall %>%
    filter(sex=="all" & age=="all" & place_of_death!="All") %>%  # keep only aggregated values over sex & age
    select(year, week_number, week_number_run_over, place_of_death, matches("deaths"), -deaths_nrs_past_average_all_causes)
    ,
    weekly_deaths_2015_2019_overall %>%
      rename(deaths_all_causes = number_of_deaths) %>%
      filter(place_of_death!="All" & year!="average") %>%  # remove aggregated entries
      mutate(year = as.integer(year)) %>%
      filter(!(year %in% 2016:2019 & week_number==53))  # remove week 53 entries from years 2016-2019
  ) %>%
  mutate(date_w_c = compute_start_date_from_week_number(week_number = week_number, year_number = year))
```

# Annual deaths for 2020 & 2021

For visualising data using maps, we need to compute data for the entire year. Note: the non-covid and covid-related deaths in 2020 don't add up to all cause deaths due to the covid status only starting in week 12!

```{r}
compute_total_deaths_in_year <- function(data_tbl, ...) {
  grouping_vars <- enquos(...)
  data_tbl %>%
    group_by(year, place_of_death, !!!grouping_vars) %>%
    summarise(across(.cols = c(deaths_all_causes, deaths_covid_related, deaths_non_covid), .fns = ~sum(.x,na.rm=TRUE)), .groups = "drop") %>%  # compute sums of all types of death
    mutate(deaths_all_causes_since_covid = deaths_covid_related + deaths_non_covid)  # compute a smaller number of all-cause deaths since covid started (over 2020, covid- and non-covid-related deaths don't add up to all-cause deaths because pre-covid deaths are included also!)
}

compute_ratio_of_annual_deaths_to_historical <- function(data_tbl) {
  data_tbl %>%
    mutate(
      ratio_annual_deaths_to_historical_mean = deaths_all_causes / annual_deaths_mean,
      ratio_annual_deaths_to_historical_min = deaths_all_causes / annual_deaths_min,
      ratio_annual_deaths_to_historical_max = deaths_all_causes / annual_deaths_max
    )
}

## helper function to remove rows of data where the historic period is inappropriate
remove_invalid_historic_period_for_year <- function(data_tbl) {
  data_tbl %>%
    filter(!(historic_period == "2016-2019, 2021" & year < 2022))
}

total_deaths_overall <-
  merged_deaths_overall %>%
  compute_total_deaths_in_year() %>%
  left_join(historical_annual_average_overall, by = c("place_of_death")) %>%
  compute_ratio_of_annual_deaths_to_historical %>%
  remove_invalid_historic_period_for_year()

total_deaths_la <-
  merged_deaths_la %>%
  compute_total_deaths_in_year(ref_area) %>%
  left_join(historical_annual_average_la, by = c("place_of_death", "ref_area")) %>%
  compute_ratio_of_annual_deaths_to_historical %>%
  remove_invalid_historic_period_for_year()

total_deaths_hb <-
  merged_deaths_hb %>%
  compute_total_deaths_in_year(ref_area) %>%
  left_join(historical_annual_average_hb, by = c("place_of_death", "ref_area")) %>%
  compute_ratio_of_annual_deaths_to_historical %>%
  remove_invalid_historic_period_for_year()

total_deaths_sex <-
  merged_deaths_sex %>%
  compute_total_deaths_in_year(sex) %>%
  left_join(historical_annual_average_sex, by = c("place_of_death", "sex")) %>%
  compute_ratio_of_annual_deaths_to_historical %>%
  remove_invalid_historic_period_for_year()

total_deaths_age <-
  merged_deaths_age %>%
  compute_total_deaths_in_year(age) %>%
  left_join(historical_annual_average_age, by = c("place_of_death", "age")) %>%
  compute_ratio_of_annual_deaths_to_historical %>%
  remove_invalid_historic_period_for_year()
```

## Annual deaths by cause of death

```{r}
## bespoke computation because format is different... deaths split by cause!
total_deaths_cause_of_death <-
  merged_deaths_cause_of_death %>%
  group_by(year, place_of_death, cause_of_death, historic_period) %>%
  summarise(deaths = sum(number_of_deaths, na.rm = TRUE), .groups="drop") %>%
  left_join(historical_annual_average_cause_of_death, by = c("place_of_death", "cause_of_death", "historic_period")) %>%
  mutate(
    ratio_annual_deaths_to_historical_mean = if_else(  # only compute this for complete years, we can't compare 2022 to annual average!
      condition = year == 2022,
      true = NA_real_,
      false = deaths / annual_deaths_mean
    )
  )

current_and_historical_deaths_by_cause_and_location <-
  bind_rows(
    ## 2020-2022 separately
    total_deaths_cause_of_death %>%
      select(year, place_of_death, cause_of_death, deaths) %>%
      mutate(year = as.character(year)),
    ## historical data
    total_deaths_cause_of_death %>%  # these are for 2020-2021, but we use the historical mean
      filter(year == 2020) %>%
      select(year, place_of_death, cause_of_death, deaths = annual_deaths_mean) %>%
      mutate(year = "2015-2019")
  ) %>%
  filter(place_of_death != "All") %>%
  group_by(year, cause_of_death) %>%
  mutate(proportion = deaths / sum(deaths)) %>%
  ungroup %>%
  arrange(year, cause_of_death)
```

## Historical average weekly deaths, including for incomplete year 2022

```{r}
merged_deaths_hb %>%
  group_by(year, ref_area, place_of_death, historic_period) %>%
  summarise(
    deaths_all_causes = sum(deaths_all_causes, na.rm = TRUE),
    deaths_mean = sum(deaths_mean, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(proportion_home_deaths_increase = (deaths_all_causes/deaths_mean) - 1)

# TODO: fix below to include historic_period
# proportions_of_historical_deaths_by_place <-
#   historical_weekly_average_overall %>% 
#   group_by(year, place_of_death) %>% summarise(number_of_deaths = sum(number_of_deaths), .groups="drop") %>%
#   filter(place_of_death != "All" & !is.na(as.integer(year))) %>%
#   group_by(year) %>%
#   mutate(
#     total_deaths = sum(number_of_deaths)
#   ) %>%
#   ungroup %>%
#   mutate(proportion_of_total = number_of_deaths / total_deaths, year=as.integer(year))
# 
# proportions_of_current_deaths_by_place <-
#   total_deaths_overall %>%
#   select(year, place_of_death, number_of_deaths = deaths_all_causes) %>%
#   filter(place_of_death != "All") %>%
#   group_by(year) %>%
#   mutate(
#     total_deaths = sum(number_of_deaths)
#   ) %>%
#   ungroup %>%
#   mutate(proportion_of_total = number_of_deaths / total_deaths)
# 
# 
# merged_proportions_of_deaths_by_place <-
#   bind_rows(
#     proportions_of_historical_deaths_by_place,
#     proportions_of_current_deaths_by_place
#   )
```

# Constants about latest data available

These are used for saying in plots what data were included e.g. weeks 1-12 in 2022.

```{r}
## constants for use in graphs & annotations
weeks_available_2022 <- max(merged_deaths_age$week_number[merged_deaths_age$year==2022])
most_recent_date_available_2022 <- max(merged_deaths_age$date_w_c)

earliest_date <- min(merged_deaths_age$date_w_c)
```

# Ordering geographical data by population size

The latest population estimates by HB/LA are available [from NRS Scotland here](https://www.opendata.nhs.scot/dataset/population-estimates).

I'm using the estimates for 2020 (based on 2019 data) to compute the relative sizes of health boards!

```{r}
population_estimate_la_2019 <- read_csv(file = "./population_estimates/ca2019_pop_est_29062021.csv") %>%
  select(Year, CA, Sex, AllAges) %>%
  clean_names() %>%
  make_human_readable_ref_area(ref_area = ca) %>%
  filter(year==2020) %>%
  filter(ca != "Scotland") %>%
  filter(sex == "All") %>%
  rename(la = ca)

population_estimate_hb_2019 <- read_csv(file = "./population_estimates/hb2019_pop_est_29062021.csv") %>%
  select(Year, HB, Sex, AllAges) %>%
  clean_names() %>%
  make_human_readable_ref_area(ref_area = hb) %>%
  filter(year==2020) %>%
  filter(hb != "Scotland") %>%
  filter(sex == "All")
```

## Compute order

```{r}
order_la_by_population <-
  population_estimate_la_2019 %>%
  arrange(desc(all_ages)) %>%
  .$la

order_hb_by_population <-
  population_estimate_hb_2019 %>%
  arrange(desc(all_ages)) %>%
  .$hb
```

# Calculate case counts: weekly counts to match weekly death reporting

Note: the cumulative case counts have already been calculated, we just need to subset them to match the weeks. This is not trivial - the death counts cover an entire week starting from the "w/c date"; the cumulative cases for that week could be taken either at the start, the end of the week, or the start of the following week even!

However, when calculating weekly positive cases we can simply take the maximum of that week's cumulative count - that will be the cumulative count at the end of the week!

Note also that the `crude_rate_positive_per100k` represents the cumulative rate!

## Calculating population size of areas

Update 7 Mr 2022: at some point the file structure changed, and the crude positive rate per 100k was removed... however, the crude mortality rate remains, so we can infer the population from the `crude_rate_deaths` and `cumulative_deaths` figures, and compute the `crude_rate_positive` from those.

The trouble is that early on in the data, when `cumulative_deaths` and `crude_rate_deaths` are both 0, we can't infer the population size. If the population size is constant in the data, we can calculate it for each `ref_area` (HB or LA) and use that as a lookup table.

It looks like the same population estimate was used for data 2020-2022, and it is the **mid-year 2019 estimate**.

## Importing & calculating case trends

```{r}
weekly_covid_cases_hb <-
  cases_hb %>%
  left_join(
    population_estimate_hb_2019 %>%
    add_scotland_total_to_population_estimate %>%
    select(ref_area = hb, population_size = all_ages)
    ) %>%
  group_by(year, week_number, ref_area) %>%
  summarise(
    weekly_positive = sum(daily_positive, na.rm = TRUE),
    cumulative_positive = max(cumulative_positive, na.rm = TRUE),
    # crude_rate_positive_per100k = max(crude_rate_positive),  # this column is no longer in the data, so we compute it below...
    crude_rate_positive_per100k = max(cumulative_positive / population_size * 1e5, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    date_w_c = compute_start_date_from_week_number(week_number = week_number, year = year),
    week_number_run_over = if_else(year == 2021, 53L + week_number, week_number)
    )

## extract Scotland-wide data and remove from HB split
weekly_covid_cases_overall <- weekly_covid_cases_hb %>% filter(ref_area=="Scotland")
weekly_covid_cases_hb <- weekly_covid_cases_hb %>% filter(ref_area!="Scotland")

weekly_covid_cases_la <-
  cases_la %>%
  left_join(
    population_estimate_la_2019 %>%
    add_scotland_total_to_population_estimate %>%
    select(ref_area = la, population_size = all_ages)
    ) %>%
  group_by(year, week_number, ref_area) %>%
  summarise(
    weekly_positive = sum(daily_positive, na.rm = TRUE),
    cumulative_positive = max(cumulative_positive, na.rm = TRUE),
    # crude_rate_positive_per100k = max(crude_rate_positive),  # this column is no longer in the data, so we compute it below...
    crude_rate_positive_per100k = max(cumulative_positive / population_size * 1e5, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    date_w_c = compute_start_date_from_week_number(week_number = week_number, year = year),
    week_number_run_over = if_else(year == 2021, 53L + week_number, week_number)
    )
```

# Comparisons of annual rates with current rates

```{r}
table_current_deaths_compared_to_past <-
  bind_rows(
    ## all available data
    merged_deaths_overall %>%
      filter(age == "all" &
               sex == "all"
             ) %>% 
      group_by(year, place_of_death, week_number) %>%
      summarise(
        deaths_all_causes = sum(deaths_all_causes),
        # deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
      mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
      left_join(
        weekly_deaths_2015_2019_overall %>% filter(year != "average") %>% group_by(week_number, place_of_death) %>% summarise(
                                           average_weekly_deaths = sum(number_of_deaths) / 5,
                                           .groups = "drop"
                                         ),
        by = c("week_number", "place_of_death")
      ) %>% 
      group_by(place_of_death) %>%
      summarise(
        period = glue::glue("Data from week 1, 2020 to week {weeks_available_2022}, 2022"),
        historic_period = "2015-2019",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        mean_annual_deaths_2015_2019 = sum(average_weekly_deaths),
        .groups = "drop"
      )
    ,
    ## 2020 only
    merged_deaths_overall %>%
      filter(age == "all" &
               sex == "all" &
               year == 2020
             ) %>% 
      group_by(place_of_death, historic_period) %>%
      summarise(
        period = "2020",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
      mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      historical_annual_average_overall %>% 
        select(place_of_death, mean_annual_deaths_2015_2019 = annual_deaths_mean, historic_period),
      by = c("place_of_death","historic_period")
    )
    ,
    ## 2020 from week 12 onwards (including week 12) - pandemic part of 2020
    merged_deaths_overall %>%
      filter(
        age == "all" &
          sex == "all" &
          year == 2020 & week_number >= 12
      ) %>%
      group_by(place_of_death) %>%
      summarise(
        period = "2020, week 12 - week 53",
        historic_period = "2015-2019",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
    mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      weekly_deaths_2015_2019_overall %>% 
        filter(week_number >= 12 & year != "average") %>%
        group_by(place_of_death) %>%
        summarise(mean_annual_deaths_2015_2019 = sum(number_of_deaths) / 5),
      by = "place_of_death"
    )
    ,
    ## 2021 only
    merged_deaths_overall %>%
      filter(age == "all" &
               sex == "all" &
               year == 2021
             ) %>% 
      group_by(place_of_death) %>%
      summarise(
        period = "2021",
        historic_period = "2015-2019",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
      mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      historical_annual_average_overall %>% 
        select(place_of_death, mean_annual_deaths_2015_2019 = annual_deaths_mean, historic_period),
      by = c("place_of_death","historic_period")
    )
    ,
    ## 2022 data so far
    merged_deaths_overall %>%
      filter(
        age == "all" &
          sex == "all" &
          year == 2022
      ) %>%
      group_by(place_of_death, week_number) %>%
      summarise(
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
    mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      weekly_deaths_2015_2019_overall %>% filter(year != "average") %>% group_by(week_number, place_of_death) %>% summarise(
                                         average_weekly_deaths = sum(number_of_deaths) / 5,
                                         .groups = "drop"
                                       ),
      by = c("week_number", "place_of_death")
    ) %>% 
    group_by(place_of_death) %>%
    summarise(
      period = glue::glue("2022 up to week {weeks_available_2022}"),
      historic_period = "2015-2019",
      deaths_all_causes = sum(deaths_all_causes),
      deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
      deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
      mean_annual_deaths_2015_2019 = sum(average_weekly_deaths),
      .groups = "drop"
    )
  ) %>%
  mutate(
    ratio_deaths_all_causes_to_historical = deaths_all_causes / mean_annual_deaths_2015_2019,
    ratio_non_covid_deaths_to_historical = deaths_non_covid / mean_annual_deaths_2015_2019,
    proportion_covid_deaths = deaths_covid_related / deaths_all_causes
  )

write_csv(x = table_current_deaths_compared_to_past, file = file.path(dir_outputs, "annual_death_rates_comparison_with_historical_deaths.csv"))

table_current_deaths_compared_to_past %>% knitr::kable(caption = "Comparing annual home death rate during COVID pandemic to historical deaths at home")
```

# Moving averages: 4 weeks

```{r}
merged_deaths_overall <-
  compute_moving_average_for_current_data(
    current_data = merged_deaths_overall, 
    historical_data = 
      weekly_deaths_2015_2019_overall %>% mutate(year=as.integer(year)) %>% drop_na(year),  # remove the entries with 'average' for year
    num_weeks_to_average = 4)

merged_deaths_hb <-
  compute_moving_average_for_current_data(
    current_data = merged_deaths_hb, 
    historical_data = 
      weekly_deaths_2015_2019_hb, 
    num_weeks_to_average = 4)

merged_deaths_la <-
  compute_moving_average_for_current_data(
    current_data = merged_deaths_la, 
    historical_data = 
      weekly_deaths_2015_2019_la, 
    num_weeks_to_average = 4)

merged_deaths_sex <-
  compute_moving_average_for_current_data(
    current_data = merged_deaths_sex, 
    historical_data = 
      weekly_deaths_2015_2019_sex, 
    num_weeks_to_average = 4)

merged_deaths_age <-
  compute_moving_average_for_current_data(
    current_data = merged_deaths_age, 
    historical_data = 
      weekly_deaths_2015_2019_age, 
    num_weeks_to_average = 4)

merged_deaths_sex_age <-
  compute_moving_average_for_current_data(
    current_data = merged_deaths_sex_age, 
    historical_data = 
      weekly_deaths_2015_2019_sex_age, 
    num_weeks_to_average = 4)
```

# Annual deaths by place, 2015-2022

```{r}
merged_proportions_of_deaths_by_place <-
  bind_rows(
    ## 2015-2019
    weekly_deaths_2015_2019_overall %>% 
    group_by(year, place_of_death) %>% summarise(number_of_deaths = sum(number_of_deaths), .groups="drop") %>%
    filter(place_of_death != "All" & !is.na(as.integer(year))) %>%
    group_by(year) %>%
    mutate(
      total_deaths = sum(number_of_deaths)
    ) %>%
    ungroup %>%
    mutate(
      proportion_of_total = number_of_deaths / total_deaths, 
      year=as.integer(year)
      ),
    ## 2020 onwards
    total_deaths_overall %>%
    select(year, place_of_death, number_of_deaths = deaths_all_causes) %>%
    distinct %>%  # remove duplicate entries for 2022 due to having multiple possible historic periods
    filter(place_of_death != "All") %>%
    group_by(year) %>%
    mutate(
      total_deaths = sum(number_of_deaths)
    ) %>%
    ungroup %>%
    mutate(proportion_of_total = number_of_deaths / total_deaths)
  )
```

## By HB/LA

```{r}
proportions_of_deaths_by_place_by_geography <-
  bind_rows(
    ## historic deaths
    bind_rows(  # merge HB & LA past deaths, and tag them appropriately
      weekly_deaths_2015_2019_hb %>% mutate(geography = "hb"),
      weekly_deaths_2015_2019_la %>% mutate(geography = "la")
    ) %>%
    group_by(year, place_of_death, ref_area, geography) %>% summarise(number_of_deaths = sum(number_of_deaths, na.rm = TRUE), .groups="drop") %>%  # compute the total number of yearly deaths by place of death & geography
    filter(place_of_death != "All" & !is.na(as.integer(year))) %>%  # remove blank/irrelevant entries
    group_by(year, ref_area, geography) %>%
    mutate(
      total_deaths = sum(number_of_deaths)  # compute total annual deaths across place of death, but within each geographical area & year
    ) %>%
    ungroup %>%
    mutate(proportion_of_total = number_of_deaths / total_deaths, year=as.integer(year))  # compute proportion of deaths in any palce relative to total annual deaths in geographical area
    ,
    ## current deaths
    bind_rows(  # merge HB & LA past deaths, and tag them appropriately
      total_deaths_hb %>% mutate(geography = "hb"),
      total_deaths_la %>% mutate(geography = "la")
    ) %>%
    select(year, place_of_death, number_of_deaths = deaths_all_causes, geography, ref_area) %>%  # keep only relevant columns & rename deaths due to all causes
    group_by(year, place_of_death, ref_area, geography) %>% summarise(number_of_deaths = sum(number_of_deaths), .groups="drop") %>%  # compute the total number of yearly deaths by place of death & geography
    filter(place_of_death != "All" & !is.na(as.integer(year))) %>%  # remove blank/irrelevant entries
    group_by(year, ref_area, geography) %>%
    mutate(
      total_deaths = sum(number_of_deaths)  # compute total annual deaths across place of death, but within each geographical area & year
    ) %>%
    ungroup %>%
    mutate(proportion_of_total = number_of_deaths / total_deaths, year=as.integer(year))  # compute proportion of deaths in any palce relative to total annual deaths in geographical area
  )

## write to csv
proportions_of_deaths_by_place_by_geography %>%
  write_csv(file = paste0(dir_outputs,"/proportions_of_deaths_by_place_by_geography.csv"))
```

# Deaths by cause, ordered by number of deaths in 2020

```{r}
## get "canonical" order for causes of death in 2020
order_of_causes_of_death <- 
  total_deaths_cause_of_death %>% 
  filter(year == 2020 & place_of_death == "All") %>% 
  arrange(desc(deaths)) %>% 
  pull(cause_of_death)
```

# TODO: other thigns like excess death & historical rates by geography

```{r}
# TODO: copy some of the old calculations over
```

# Save most recent workspace for further analyses

```{r}
# save.image(file = "./workspace.RData", safe = TRUE)  # this is calling .GlobalEnv, which is empty
save(list = ls(all.names = TRUE), file = "./workspace.RData", envir = environment())
```

# Print session info

```{r}
sessionInfo()
```