### 2023-01-11 parsing_toc/example.Rmd
toy example for parsing .md files created alongside rendering rmarkdown files, generating a listing of 3rd level headings (###) and captions of tables using DT::datatable()

---
title: "Example of report with tables"
output: 
  html_document:
    keep_md: true
    number_sections: true
date: "2022-11-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(DT)
```

# Top level header

## Subheading L2 without much content

### What

### is this

## Subheading

### L3 subheading without a table

Just some lorem ipsum

### Table 1: SW characters shorter than 100

Some lorem ipsum that doesn't need to be in a listing of tables.

```{r}
starwars %>% filter(height<100) %>%
datatable(caption = "SW characters shorter than 100 units, presumably centimeters.")
```

### Table 2: SW characters taller than 200

Some more lorem ipsum that doesn't need to be in a listing of tables.

```{r}
starwars %>% filter(height>200) %>%
  datatable(caption = "SW characters taller than 200 units, presumably centimeters.")
```

## Subheading 2

### Table 2 contains `glue`

Some more lorem ipsum that doesn't need to be in a listing of tables.

```{r}
starwars %>% filter(height>200) %>%
  datatable(caption = glue::glue("SW characters taller than 200 units, presumably centimeters."))
```