### 2023-02-13 compile_medications_from_bnf.Rmd
added more testing code for BNF medicine compiling

---
title: "Compiling list of medications based on BNF codes for dementia/AD & palliative care drugs"
author: "Jan Savinc"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

For the analysis of prescription/dispensing data, lists of specific medicines are needed to make the analysis tractable and applicable to dementia/Alzheimer's disease and for the analysis of anticipatory medicines used in palliative care.

## Source of BNF codes

One source for a list of all BNF codes is the Information Services Portal of the NHS Business Services Authority, https://applications.nhsbsa.nhs.uk/infosystems/data/showDataSelector.do?reportId=126. This is available after putting in a captcha-like password, so I've downloaded the file manually.

File used: `01-01-2023 : 84`

## List of discrete medications

To count the number of different prescriptions or measure polypharmacy, we need to define discrete medications that can be prescribed/dispensed, whilst removing things like dressings, vaccines, etc.

[Guthrie et al., 2015 ("The rising tide of polypharmacy and drug-drug interactions: population database analysis 1995–2010")](https://doi.org/10.1186/s12916-015-0322-7) define this in their Additional file, table S1, which I'll copy-paste and export as .csv below.

# Load packages

```{r}
library(tidyverse)
library(readxl)  # for loading excel files
library(janitor)  # for cleaning data
library(knitr)  # for pretty tables in the report
library(docxtractr)  # for reading data from table in .docx file
library(fuzzyjoin)  # for regex merges
```

# Download file

## BNF codes

Note: this was done manually from the above website!

## List of medicines

Note: there appears to be an error with using `docxtractor::read_docx()` with a file downloaded to OneDrive; reading directly from the URL works fine, however! Therefore I've skipped downloading the file below.

```{r}
url_medicines <- "https://static-content.springer.com/esm/art%3A10.1186%2Fs12916-015-0322-7/MediaObjects/12916_2015_322_MOESM1_ESM.doc"

# path_medicines <- file.path("./bnf", basename(url_medicines))
# download.file(url = url_medicines, destfile = path_medicines)
```

# Load BNF data

```{r}
## the .csv file is inside a .zip file!
bnf <- read_csv(file = unzip( "./bnf/20230104_1672833147341_BNF_Code_Information.zip", "20230104_1672833147341_BNF_Code_Information.csv")) %>%
  clean_names()
```

## List of medicines

```{r}
# doc <- read_docx(path_medicines)
## error with reading from OneDrive folder?
doc <- read_docx(url_medicines)
extracted_table <- 
  docx_extract_tbl(doc, 1)
```

# Defining medicines

## Helper functions for searching BNF

```{r}
find_substance <- function(keyword) {
  bnf %>%
    filter(str_detect(bnf_chemical_substance, pattern = regex(keyword, ignore_case = TRUE))) %>%
    group_by(bnf_chemical_substance, bnf_paragraph) %>%
    summarise(
      n = n(),
      products = paste(bnf_presentation, collapse = ";")
    )
}
```

## Dementia / Alzheimer's disease

The medicines of interest are:

- donepezil
- galantamine
- rivastigmine
- memantine

~~I can manually search for the above keywords in the `chemical substance` field in the BNF and include all relevant medicines.~~

Note that the above drugs correspond to the entirety of the BNF section `4.11, Drugs for dementia`

```{r}
bnf_dementia <- 
  bnf %>%
  filter(bnf_section == "Drugs for dementia") %>%
  mutate(medication = word(bnf_chemical_substance, 1)) %>%
  select(medication, bnf_section, bnf_section_code, bnf_chemical_substance_code) %>%
  distinct
```

## Palliative care / anticipatory medicines

Marie Curie lists [the following anticipatory medicines](https://www.mariecurie.org.uk/professionals/palliative-care-knowledge-zone/symptom-control/anticipatory-medicines):

- Medicine for pain in palliative care – an appropriate opioid, for example, morphine, diamorphine, oxycodone or alfentanil.
- Medicine for breathlessness – midazolam or an opioid.
- Medicine for anxiety – midazolam.
- Medicine for delirium or agitation – haloperidol, levomepromazine, midazolam or phenobarbital.
- Medicine for nausea and vomiting – cyclizine, metoclopramide, haloperidol or levomepromazine.
- Medicine for noisy chest secretions – hyoscine hydrobromide or glycopyrronium.

Some of the above are specific chemicals, and some are classes of medicines (e.g. opioids). For opioids, some examples are listed, but I will include the entire class, i.e. the BNF section `4.7.2, Opioid analgesics`. For the other medicines, I will only include the specific substances listed.

Note that alfentanil is not listed as an opioid in the BNF but as a general anaesthetic!

```{r}
bnf_palliative <-
  ## pain: opioids + alfentanil
  list(
    bind_rows(
      bnf %>% filter(bnf_paragraph_code == "040702"),  # opioids
      bnf %>% filter(str_detect(bnf_chemical_substance, pattern = regex("alfentanil", ignore_case = TRUE)))
    ) %>% mutate(medicine = "Pain")
    ,
    ## breathlessness: midazolam or opioid
    bind_rows(
      bnf %>% filter(bnf_paragraph_code == "040702"),  # opioids
      bnf %>% filter(str_detect(bnf_chemical_substance, pattern = regex("midazolam", ignore_case = TRUE)))
    ) %>% mutate(medicine = "Breathlessness")
    ,
    ## anxiety
    bnf %>% filter(str_detect(bnf_chemical_substance, pattern = regex("midazolam", ignore_case = TRUE))) %>% 
      mutate(medicine = "Anxiety")
    ,
    ## delirium/agitation
    bnf %>% filter(str_detect(bnf_chemical_substance, pattern = regex("haloperidol|levomepromazine|midazolam|phenobarbital", ignore_case = TRUE))) %>% 
      mutate(medicine = "Delirium or agitation")
    ,
    ## nausea or vomiting
    bnf %>% filter(str_detect(bnf_chemical_substance, pattern = regex("cyclizine|metoclopramide|haloperidol|levomepromazine", ignore_case = TRUE))) %>% 
      mutate(medicine = "Nausea or vomiting")
    ,
    ## noisy chest secretions
    bnf %>% filter(str_detect(bnf_chemical_substance, pattern = regex("hyoscine hydrobromide|glycopyrronium", ignore_case = TRUE))) %>% 
      mutate(medicine = "Noisy chest secretions")
  ) %>%
  bind_rows() %>%
  select(medicine, bnf_chemical_substance, bnf_chemical_substance_code) %>%
  distinct %>%
  pivot_wider(names_from = medicine, values_from = medicine, names_prefix = "bin_", values_fill = 0L, values_fn = ~1L) %>%
  clean_names()
```

## Discrete medicines from Guthrie et al., 2015 paper

```{r}
# TODO: process this if needed?
medicines_list <- 
  extracted_table %>%
  clean_names() %>%
  mutate(
    chapter = as.integer(str_extract(drug_class_name, pattern = "(?<=Chapter )\\d+"))
  ) %>%
  fill(chapter, .direction = "down") %>%  # complete blanks between chapters with the last non-blank chapter
  filter(bnf_subsection!="")  # remove chapters from the drug class column

zero_pad <- function(x) str_pad(string = x, width = 2, pad = "0")

test_list <-
  medicines_list %>%
  separate(col = bnf_subsection, into = paste0("c",1:4), sep = "\\.", remove = FALSE) %>%
  mutate(across(c1:c3, zero_pad)) %>%  # zero-pad digits except for 4th character, c4
  unite(col = "code", c1:c4, sep = "", na.rm = TRUE, remove = TRUE)
```

## Test against BNF to see if I get the same substances

Note: these need to be done in sections by how long the code given is - does it pick up a chapter, a section, a subsection, or specific medicines only?

```{r}
test_merge <-
  # test_list %>% 
  # select(drug_class_name, bnf_subsection, code) %>%
  # mutate(code = paste0("^",code)) %>%
  # regex_right_join(bnf, by = c("code"="bnf_chemical_substance_code"))
  bnf %>%
  regex_right_join(
    test_list %>%
      select(drug_class_name, bnf_subsection, code) %>%
      mutate(code = paste0("^",code)),
    by = c(bnf_chemical_substance_code = "code")
  )

# View(test_merge)

## what's left after matching all of the bnf to the provided codes?
test_merge2 <-
  bnf %>%
  regex_anti_join(
    test_list %>%
      select(drug_class_name, bnf_subsection, code) %>%
      mutate(code = paste0("^",code)),
    by = c(bnf_chemical_substance_code = "code")
  )
test_merge2 %>% count(bnf_section, bnf_section_code)
```

# Conclusion

We are now left with two files listing medicines used in dementia/AD and palliative/anticipatory care; the palliative list is slightly more complex and it includes, for each substance, a number of binary variables indicating its use (e.g. pain, breathlessness, agitation, etc.)

# Save .csv files

```{r}
write_csv(bnf_dementia, file = "./bnf/bnf_codes_dementia_alzheimers.csv")

write_csv(bnf_palliative, file = "./bnf/bnf_codes_palliative_anticipatory.csv")
```