### 2024-05-08 covid_nrs_place_of_death_data_2024.Rmd
modified these to include finalised 2022 data as historical, and to download the latest 2024 dataset; I've deleted & commented out the bits for the disaggregated age & sex data (currently only the split by sex/age or PoD, but not both, is available)

---
title: "Place of death in NRS COVID-19 deaths: data processing for analysis, 2024 data"
author: "Jan Savinc"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages

```{r, warning=FALSE, message=FALSE}
## install packages from github if not yet available!
### remotes::install_github("datasciencescotland/opendatascot", force = TRUE)  # this is the old repo
# devtools::install_github("jsphdms/opendatascot")
# remotes::install_github("Health-SocialCare-Scotland/phsmethods", force = TRUE)

library(tidyverse)  # for tidy workflow
library(opendatascot)  # importing data from ScotGov open data website
library(phsmethods)  # methods for working with PHS data
library(readxl)  # for reading excel files
# library(SPARQL)  # taken care of by opendatascot
library(lubridate)  # dealing with dates
library(janitor)  # for cleaning column names
library(ISOweek)  # for computing date from ISO week number + year
library(sf)  # for mapping
library(ggrepel)  # for 'mapping'repelling' labels/texst in ggplot
library(patchwork)  # for assembling plots
library(extrafont)  # for working with fonts
library(openxlsx)  # for creating xlsx files
library(knitr)  # for displaying tables in rmd

## load functions
source("./functions.R")
```

# Copyright attribution

All data used for this study were obtained from the National Records Scotland (NRS) and are © Crown copyright, 2020; the details of the licence can be viewed on the [Open Government Licence website](http://www.nationalarchives.gov.uk/doc/open-government-licence/open-government-licence.htm)

# Dates represent event registration, not occurrence!

From NRS website:

> All routine vital events information we publish is based on the date of registration, not the date on which the event occurred. Inevitably, of course, there are delays between the occurrence of an event and its registration.

# Load info about latest data

A separate script checks for updated data and downloads it, also updating these files with the locations of the downloaded data:

```{r}
file_data_sources <- "./table_of_data_sources.csv"
file_historical_data_sources <- "./table_of_historical_data_sources.csv"
file_case_trends_data_sources <- "./table_of_case_trends_data_sources.csv"

table_of_data_sources <- read_csv(file = file_data_sources)
table_of_historical_data_sources <- read_csv(file = file_historical_data_sources)
table_of_case_trends_data_sources <- read_csv(file = file_case_trends_data_sources)

table_of_data_sources %>%
  knitr::kable(caption = "Latest data used in this report.")

files_weekly_deaths <- set_names(x = as.list(table_of_data_sources$path_latest), nm = table_of_data_sources$short_name)
files_historical_deaths <- set_names(x = as.list(table_of_historical_data_sources$path_latest), nm = table_of_historical_data_sources$short_name)
files_case_counts <- set_names(x = as.list(table_of_case_trends_data_sources$path_latest), nm = table_of_case_trends_data_sources$short_name)
```

## Set up output directory

Because data are updated on a weekly basis, we'll have a separate directory for every week the data are updated:

```{r}
latest_data_modified_date <- strftime(max(table_of_data_sources$last_modified), format="%F")
latest_year <- year(latest_data_modified_date)

dir_outputs <- file.path("./outputs", latest_year, latest_data_modified_date)
if (!dir.exists(dir_outputs)) dir.create(dir_outputs, recursive = TRUE)
```

# Import & wrangle data

There are several files that the data are imported from.

TODO: now that the ad-hoc files are no longer available, some of the data will need to be compiled from the tabs in the weekly data.

## Overall data

Note: this contains data split by sex and age, but only for ALL places of death; individual PoD data are available for ALL sex and ALL age only

```{r}
## a separate raw file makes it easier to debug, and also reduces the number of requests I make while debugging!
ods_raw <- ods_dataset("deaths-involving-coronavirus-covid-19", geography = "sc")

## Note: as of 12 Dec 2022, the format of the ref_period variable changed from
## e.g. "2020-W12" to "w/c 2019-12-31"
## this means the w/c value is already provided
## and also the week number is not provided

weekly_deaths_overall <-
  ods_raw %>%
  make_human_readable_ref_area %>%
  clean_names() %>%
  ## below series of rename operations are a stupid way of dealing with the possibility of different variable name conventions in the ods_dataset data: sometimes they are in snake_case, and sometimes in camelCase, seemingly unpredictably
  rename_at(.vars = vars(matches("location", ignore.case = TRUE)), ~"place_of_death") %>%
  rename_at(.vars = vars(matches("period", ignore.case = TRUE)), ~"ref_period") %>%
  rename_at(.vars = vars(matches("area", ignore.case = TRUE)), ~"ref_area") %>%
  rename_at(.vars = vars(matches("measure", ignore.case = TRUE)), ~"measure_type") %>%
  rename_at(.vars = vars(matches("cause", ignore.case = TRUE)), ~"cause_of_death") %>%
  rename_at(.vars = vars(matches("sex", ignore.case = TRUE)), ~"sex") %>%
  rename_at(.vars = vars(matches("age", ignore.case = TRUE)), ~"age") %>%
  rename_at(.vars = vars(matches("value", ignore.case = TRUE)), ~"number_of_deaths") %>%  # rename so it's consistent with other data
  filter(!ref_period %in% as.character(2020:2029)) %>% # remove totals for 2020, 2021, 2022, etc., we only want the weekly figures
  parse_ref_period() %>%
  mutate(
    number_of_deaths = as.numeric(number_of_deaths),  # convert to numeric!
  ) %>%
  select(-c(measure_type)) %>%  #  measure_type is count for all data
  recode_place_of_death() %>%
  recode_cause_of_death() %>%
  recalculate_number_of_deaths_after_recoding_cause_of_death %>%  # not needed here probably, but no harm in including it
  pivot_wider(names_from = cause_of_death, values_from = number_of_deaths) %>%  # this dataset is well-formatted and has no implicit missing values!
  mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related)  # compute non-covid deaths from the other two figures - will be NA where either is missing
```

## LA & HB

```{r}
weekly_deaths_la <- 
  read_excel(files_weekly_deaths$la, sheet = "Data", skip = 2) %>% 
  clean_names %>%
  rename(week_number = week_of_occurrence, ref_area = council, place_of_death = location_of_death, number_of_deaths = deaths) %>%
  filter(!is.na(number_of_deaths)) %>%  # remove blank rows & copyright entry
  parse_year_and_week_number_from_ad_hoc_data %>%
  recode_cause_of_death() %>%
  recalculate_number_of_deaths_after_recoding_cause_of_death %>%  # recompute the sum of deaths
  tidyr::complete(
      nesting(year, week_number, week_number_run_over, date_w_c),  # these are 'identifying' variables that shouldn't be combinatorily exploded
      ref_area, cause_of_death, place_of_death,  # these are the variables we want to have all combinations of, and missing combinations are set to 0 deaths
      fill = list(number_of_deaths = 0)
      ) %>%
  recode_place_of_death() %>%  # this needs to be done before recoding cause of death, otherwise we also include the "all" figure which isn't included in this dataset
  pivot_wider(names_from = cause_of_death, values_from = number_of_deaths) %>%   # make separate columns for causes of death figures
  mutate(deaths_all_causes = deaths_non_covid + deaths_covid_related) %>%  # compute all cause deaths from the other two figures reported
  set_covid_and_non_covid_deaths_to_zero_before_reporting_started()

weekly_deaths_hb <- 
  read_excel(files_weekly_deaths$hb, sheet = "Data", skip = 2) %>% 
  clean_names %>%
  rename(week_number = week_of_occurrence, ref_area = health_board, place_of_death = location_of_death, number_of_deaths = deaths) %>%
  filter(!is.na(number_of_deaths)) %>%  # remove blank rows & copyright entry
  parse_year_and_week_number_from_ad_hoc_data %>%
  recode_cause_of_death() %>%
  recalculate_number_of_deaths_after_recoding_cause_of_death %>%  # recompute the sum of deaths
  tidyr::complete(
      nesting(year, week_number, week_number_run_over, date_w_c),  # these are 'identifying' variables that shouldn't be combinatorily exploded
      ref_area, cause_of_death, place_of_death,  # these are the variables we want to have all combinations of, and missing combinations are set to 0 deaths
      fill = list(number_of_deaths = 0)
      ) %>%
  recode_place_of_death() %>%  # this needs to be done before recoding cause of death, otherwise we also include the "all" figure which isn't included in this dataset
  pivot_wider(names_from = cause_of_death, values_from = number_of_deaths) %>%  # make separate columns for causes of death figures
  mutate(deaths_all_causes = deaths_non_covid + deaths_covid_related) %>%  # compute all cause deaths from the other two figures reported
  set_covid_and_non_covid_deaths_to_zero_before_reporting_started()
```

## 'Main report' file

The format of the main report file has changed from what was used in 2021.

However, the main report file is still only being used for cause of death data, which simplifies parsing the file!

Causes of death are reported in `sheet 8` of the report, ~~and include figure from 2021 and 2022. The worksheet contains data split by location of death, which are contained within different tables arranged vertically, so their boundaries need to be written out; the tables have been pre-filled for all of 2021 and 2022, so some proportion of entries will be blank for 2022.~~

Note: the above no longer holds as of 20 July 2022 - only 2022 data re included in `sheet 8` now, which means the 2021 data need to be imported from an older file.

Note that excess deaths and 5-year average deaths are computed differently between 2020-2021 and 2022:

> For data registered in 2021, any comparison to a five year average is being made with the 2015-2019 average. For 2022, the comparison is being made to the average of 2016, 2017, 2018, 2019, 2021.

> Excess deaths are calculated by comparing the current year to the five year average from previous years. NRS, along with ONS and NISRA, agreed to leave 2020 out of five year average calculations so the five-year average which is used to compare 2022 figures against covers the years 2016, 2017, 2018, 2019 and 2021. More information about this can be found in a paper published on our website (see link on the right). Moveable public holidays, when registration offices are closed, affect the number of registrations made in the published weeks and in the corresponding weeks in previous years.

### Weekly causes of death

```{r}
# TODO: delete the below line, used for testing
files_weekly_deaths$main_report <- "../../../Downloads/weekly-deaths-24-week-17.xlsx"

sheet_cause_of_death_data <- "8"

## updated for 2024, this has all of 2023 plus incomplete 2024 data, with blanks up to week 52
## note that there are fewer columns now because excess deaths have been moved to a separate spreadsheet
# TODO: 2024 has 52 ISOweeks, but 2024-12-30 is W01 of 2025, or the 53rd week of 2024, so this may change
cell_ranges_for_causes_of_death_data <- list(
  "All" = "A6:J110",
  "Care home" = "A113:J217",
  "Home & other non-institution" = "A220:J324",
  "Hospital" = "A327:J431",
  "Other" = "A434:J538"
)

## updated for 2023 which contains both 2022 and 2023
# cell_ranges_for_causes_of_death_data <- list(
#   "All" = "A6:V110",
#   "Care home" = "A113:V217",
#   "Home & other non-institution" = "A220:V324",
#   "Hospital" = "A327:V431",
#   "Other" = "A434:V538"
# )

## this was for only 2022 data
## ...and is correct again for 2023
# cell_ranges_for_causes_of_death_data <- list(
#   "All" = "A6:V58",
#   "Care home" = "A61:V113",
#   "Home & other non-institution" = "A116:V168",
#   "Hospital" = "A171:V223",
#   "Other" = "A226:V278"
# )

weekly_deaths_cause_of_death_2023_onwards <-
  map_dfr(
    .x = cell_ranges_for_causes_of_death_data,
    .f = ~readxl::read_excel(files_weekly_deaths$main_report, sheet = sheet_cause_of_death_data, range = .x), 
    .id = "place_of_death"
  ) %>%
  janitor::clean_names() %>%
  janitor::remove_empty(which = "cols") %>%
  rename(year = registration_year, date_w_c = week_beginning) %>%
  select(-matches("_excess$")) %>%  # remove excess deaths, not currently used in analysis
  # select(-matches("_excess$|_average$")) %>%  # we can compute excess & averages later
  pivot_longer(cols = 5:ncol(.), names_to = "name", values_to = "number_of_deaths") %>%
  filter(!str_detect(name, "all_causes")) %>%  # we can calculate that from the others
  mutate(
    year = as.integer(year),
    measure_type = if_else(condition = str_detect(name, "_five_year_average"), true = "deaths_mean", false = "number_of_deaths"),
    cause_of_death = str_remove_all(name, "_five_year_average|_deaths"),
    cause_of_death = return_pretty_cause_of_death(cause_of_death),
    date_w_c = as.Date(date_w_c, origin = "1899-12-30"),  # origin to account for Excel's integer date format
    week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020)  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
  ) %>%
  select(-name) %>%
  pivot_wider(names_from = measure_type, values_from = number_of_deaths)
  # filter(!is.na(number_of_deaths))   # this will also remove the historic mean values, so I've opted not to remove blank entries!
```

### Data from previous years, 2020-2022

The reports available from NRS in 2022 include data from 2021-2022 up until week 25, and partly in weeks 26 and 27 where there is data from 2021 for specific places of death but not for all places of death. I also include data from 2020 for context. For this I'll load the last 'main report' file from 2021.

The weekly causes of death data for 2020 is included in the sheet `Table 3 (2020)`.

The weekly causes of death data for 2021 is included in the week 25 'main report' file, table `8`.

~~Note that due to how I originally processed these files, the weekly CoD data for 2020 *includes* the 5-year historic weekly average for 2015-2019, whereas I calculate that manually later for the 2021 data and later.~~ This is now changed so that the weekly CoD data is now

#### 2020

```{r}
# this uses the combined 2020-2021 file
# nrs_weekly_data_sheet_2020 <- 
#   excel_sheets(files_historical_deaths$`2020-2021`)[which(str_detect(tolower(excel_sheets(files_historical_deaths$`2020-2021`)), pattern = "table 3.*2020"))]

nrs_weekly_data_sheet_2020 <- 
  excel_sheets(files_historical_deaths$`2020`)[which(str_detect(tolower(excel_sheets(files_historical_deaths$`2020`)), pattern = "table 3.*2020"))]

## construct column names for the data
headers_in_nrs_data_2020 <-
  c(
    "place_of_death",
    "cause_of_death",
    paste0("week_", 1:53),
    "blank_remove",  # this and the rest are irrelevant entries that we'll remove
    "remove_1",
    "remove_2"
  )

## extract data... this is dense but I've tried to annotate it clearly!
weekly_deaths_cause_of_death_2020 <-
  readxl::read_excel(files_historical_deaths$`2020`, sheet = nrs_weekly_data_sheet_2020, col_names = headers_in_nrs_data_2020, range = cell_rows(6:125)) %>%  # the data are inrows 6:125
  select(-matches("remove")) %>%  # remove irrelevant summary columns at the end
  mutate(across(.cols = matches("week"), .fns = as.numeric)) %>%  # remove erroneous "count1, count2, ..." entries due to superscript in previous column
  mutate(dataset = if_else(condition = str_detect(string = cause_of_death, pattern = regex(pattern = "^registered|^difference", ignore_case = TRUE)), true = cause_of_death, false = NA_character_)) %>%  # extract the Registered deaths: historical, 2020/2021, and Difference sections
  tidyr::fill(place_of_death, .direction = "down") %>%  # complete the entries for place of death from the "last observed"
  tidyr::fill(dataset, .direction = "down") %>%  # same as above but for "dataset" column
  drop_na(week_1) %>%  # remove blank "header" rows with no data recorded
  pivot_longer(cols = matches("week"), names_to = "week_number", values_to = "number_of_deaths", values_drop_na = TRUE) %>%  # convert to long data format
  mutate(
    week_number = as.integer(parse_number(week_number)),  # extract week number
    place_of_death =  # find best match in the reference names used in other datasets
      order_of_place_of_death_levels[stringdist::amatch(x = str_sub(place_of_death,1,4), table = str_sub(order_of_place_of_death_levels,1,4), method = "lv", maxDist = 5)],  # the reference is in the variable order_of_place_of_death_levels
    year = 2020,
    dataset = str_remove_all(dataset , "Registered deaths\\: "),
    date_w_c = compute_start_date_from_week_number(week_number = week_number, year = year),
    week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020)  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
    ) %>%
  filter(dataset!="Difference") %>%
  pivot_wider(names_from = dataset, values_from = number_of_deaths) %>%
  rename(deaths_mean = `five year average`, number_of_deaths = `2020`)
```

#### 2021

```{r}
cell_ranges_for_causes_of_death_data_2021 <- list(
  "All" = "A6:V58",
  "Care home" = "A113:V165",
  "Home & other non-institution" = "A220:V272",
  "Hospital" = "A327:V379",
  "Other" = "A434:V486"
)

weekly_deaths_cause_of_death_2021 <-
  map_dfr(
    .x = cell_ranges_for_causes_of_death_data_2021,
    .f = ~readxl::read_excel(files_historical_deaths$`2021`, sheet = sheet_cause_of_death_data, range = .x), 
    .id = "place_of_death"
  ) %>%
  janitor::clean_names() %>%
  rename(year = registration_year, date_w_c = week_beginning) %>%
  select(-matches("_excess$")) %>%  # remove excess deaths, not currently used in analysis
  # select(-matches("_excess$|_average$")) %>%  # we can compute excess & averages later
  pivot_longer(cols = 5:ncol(.), names_to = "name", values_to = "number_of_deaths") %>%
  filter(!str_detect(name, "all_causes")) %>%  # we can calculate that from the others
  mutate(
    year = as.integer(year),
    measure_type = if_else(condition = str_detect(name, "_five_year_average"), true = "deaths_mean", false = "number_of_deaths"),
    cause_of_death = str_remove_all(name, "_five_year_average|_deaths"),
    cause_of_death = return_pretty_cause_of_death(cause_of_death),
    date_w_c = as.Date(date_w_c, origin = "1899-12-30"),  # origin to account for Excel's integer date format
    week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020)  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
  ) %>%
  select(-name) %>%
  pivot_wider(names_from = measure_type, values_from = number_of_deaths) %>%
  filter(!is.na(number_of_deaths))  # remove blank entries - these are for the current year but for dates in the future
```

#### 2022

```{r}
# TODO: deal with 2022 data here, now that it's 'historical', I load the last known 2023 file that included 2022 data, 2023-07-19_weekly-deaths-23.xlsx
cell_ranges_for_causes_of_death_data_2022 <- list(
  "All" = "A6:V58",
  "Care home" = "A61:V113",
  "Home & other non-institution" = "A116:V168",
  "Hospital" = "A171:V223",
  "Other" = "A226:V278"
)

weekly_deaths_cause_of_death_2022 <-
  map_dfr(
    .x = cell_ranges_for_causes_of_death_data_2022,
    .f = ~readxl::read_excel(files_historical_deaths$`2022`, sheet = sheet_cause_of_death_data, range = .x), 
    .id = "place_of_death"
  ) %>%
  janitor::clean_names() %>%
  rename(year = registration_year, date_w_c = week_beginning) %>%
  select(-matches("_excess$")) %>%  # remove excess deaths, not currently used in analysis
  # select(-matches("_excess$|_average$")) %>%  # we can compute excess & averages later
  pivot_longer(cols = 5:ncol(.), names_to = "name", values_to = "number_of_deaths") %>%
  filter(!str_detect(name, "all_causes")) %>%  # we can calculate that from the others
  mutate(
    year = as.integer(year),
    measure_type = if_else(condition = str_detect(name, "_five_year_average"), true = "deaths_mean", false = "number_of_deaths"),
    cause_of_death = str_remove_all(name, "_five_year_average|_deaths"),
    cause_of_death = return_pretty_cause_of_death(cause_of_death),
    date_w_c = as.Date(date_w_c, origin = "1899-12-30"),  # origin to account for Excel's integer date format
    week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020)  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
  ) %>%
  select(-name) %>%
  pivot_wider(names_from = measure_type, values_from = number_of_deaths) %>%
  filter(!is.na(number_of_deaths))  # remove blank entries - these are for the current year but for dates in the future
```

## Historical weekly deaths

The format of the .xlsx files is that sheet 1 is a table of contents, sheet 2 contains the data, and sheet 3 is a pivot table for subsetting data. These are downloaded from the ad-hoc section also.

```{r, warning=FALSE}
weekly_deaths_2015_2019_overall <- 
  read_csv(files_historical_deaths$overall, skip = 2) %>%
  rename(year = 1) %>%  # rename first column
  mutate(
    place_of_death = if_else(is.na(`1`), true = year, NA_character_)
  ) %>%
  fill(place_of_death, .direction="down") %>%  # LOCF on place_of_death
  drop_na(`1`) %>%
  pivot_longer(cols = all_of(as.character(1:53)), values_to = "number_of_deaths", names_to = "week_number") %>%
  mutate(week_number=as.integer(week_number)) %>%
  recode_place_of_death()

weekly_deaths_2015_2019_hb <- 
  readxl::read_excel(files_historical_deaths$hb, sheet = 2, skip = 2, col_names = TRUE) %>%  # first two rows are empty, row 3 has headers
  clean_names() %>%
  drop_na(number_of_deaths) %>%  # remove empty rows & one row where it's just the copyright notice
  rename(  # same format as current data
    week_number = matches("week"),
    place_of_death = matches("location"),
    ref_area = health_board
  ) %>%
  mutate(
    week_number = as.integer(week_number)
  ) %>%
  tidyr::complete(
    nesting(week_number,year),  # unique combinations that we don't want to combinatorily explode
    ref_area, place_of_death,  # fill in missing combinations of hb & pod with 0 deaths
    fill = list(number_of_deaths = 0)
  ) %>% 
  recode_place_of_death()  # recode place of death after completing missing combinations, otherwise the factor level "All" gets exploded also

weekly_deaths_2015_2019_la <-
  readxl::read_excel(files_historical_deaths$la, sheet = 2, skip = 2, col_names = TRUE) %>%  # first two rows are empty, row 3 has headers
  clean_names() %>%
  drop_na(number_of_deaths) %>%  # remove empty rows & one row where it's just the copyright notice
  rename(  # same format as current data
    week_number = matches("week"),
    place_of_death = matches("location"),
    ref_area = council_area
  ) %>%
  mutate(
    week_number = as.integer(week_number)
  ) %>%
  tidyr::complete(
    nesting(week_number,year),  # unique combinations that we don't want to combinatorily explode
    ref_area, place_of_death,  # fill in missing combinations of la & pod with 0 deaths
    fill = list(number_of_deaths = 0)
  ) %>% 
  recode_place_of_death()  # recode place of death after completing missing combinations, otherwise the factor level "All" gets exploded also

weekly_deaths_2015_2019_sex_age <-  # structure here is a bit more complicated - headers in row 3, but sex & age headers in row 4
  readxl::read_excel(files_historical_deaths$sex_age, sheet = 2, skip = 4, col_names = c("year","week_number","place_of_death","sex","age","number_of_deaths"), guess_max = 1e5) %>%
  drop_na(number_of_deaths) %>%    # remove empty rows & one row where it's just the copyright notice
  mutate(
    year = 2000L + as.integer(year),
    week_number = as.integer(week_number)
  ) %>%
  tidyr::complete(
    nesting(week_number,year),  # unique combinations that we don't want to combinatorily explode
    sex, age, place_of_death,  # fill in missing combinations of age, sex, pod, with 0 deaths
    fill = list(number_of_deaths = 0)
  ) %>% 
  recode_place_of_death()

## compute weekly data for sex and age separately
## sum over sex or age, respectively
weekly_deaths_2015_2019_sex <-
  weekly_deaths_2015_2019_sex_age %>%
  group_by_all %>%
  ungroup(age, number_of_deaths) %>%  # this is done by grouping over everything except the number of deaths & sex
  summarise(number_of_deaths = sum(number_of_deaths), .groups = "drop")

weekly_deaths_2015_2019_age <-
  weekly_deaths_2015_2019_sex_age %>%
  group_by_all %>%
  ungroup(sex, number_of_deaths) %>%  # this is done by grouping over everything except the number of deaths & sex
  summarise(number_of_deaths = sum(number_of_deaths), .groups = "drop")
```

## Map data

I'm using `sf` to read these and to include them invisualisations.

* Local Authority Districts (LAD) for 2020 (BUC-level resolution): https://geoportal.statistics.gov.uk/datasets/local-authority-districts-december-2020-uk-buc/
* Health Boards https://data.gov.uk/dataset/27d0fe5f-79bb-4116-aec9-a8e565ff756a/nhs-health-boards
  - Note: this is a fairly high resolution, don't need this much detail!
  - I found a discussion of this on [a github covid-related project](https://github.com/tomwhite/covid-19-uk-data/issues/18#issuecomment-622353841), and user **robchallen** produced a zip file containing [UK-wide health board boundaries](https://github.com/tomwhite/covid-19-uk-data/files/4563933/UK_covid_reporting_regions.zip)

```{r}
shapefile_hb_uk_wide <- st_read(dsn = "./map_data/hb_uk_wide/UK_covid_reporting_regions")
shapefile_la <- st_read(dsn = "./map_data/la")
```

## COVID case counts by LA/HB

```{r}
cases_hb <- read_csv(
  files_case_counts$hb, 
  col_types = 
    cols(
    .default = col_double(),
    `_id` = col_skip(),
    DailyDeaths = col_skip(),  # not needed, we get this elsewhere!
    Date = col_date(format = "%Y%m%d"),
    HB = col_character(),
    HBName = col_character(),
    HospitalAdmissionsQF = col_character(),
    ICUAdmissionsQF = col_character()
    ),
  ) %>% clean_names() %>%
  rename(ref_area = hb_name) %>%
  mutate(
    year = parse_integer(str_sub(ISOweek(date), start = 1, end = 4)),  # parse ISO year, otherwise the first few days of 2021 will belong to 2021 erroneously!
    week_number = parse_integer(str_sub(ISOweek(date), start = 7, end = 8)),
    ref_area = phsmethods::match_area(hb)  # convert to standard HB names
    )

cases_la <- read_csv(
  file = files_case_counts$la, 
  col_types = 
    cols(
    .default = col_double(),
    `_id` = col_skip(),
    DailyDeaths = col_skip(),  # not needed, we get this elsewhere!
    Date = col_date(format = "%Y%m%d"),
    CA = col_character(),
    CAName = col_character()
  )
) %>% clean_names() %>%
  rename(ref_area = ca_name) %>%
  mutate(
    year = parse_integer(str_sub(ISOweek(date), start = 1, end = 4)),  # parse ISO year, otherwise the first few days of 2021 will belong to 2021 erroneously!
    week_number = parse_integer(str_sub(ISOweek(date), start = 7, end = 8)),
    ref_area = phsmethods::match_area(ca)  # convert to standard LA names, jsut in case they weren't already!
    )
```

# Historic data for deaths

From the [methodology for computing excess deaths during the covid pandemic](https://www.nrscotland.gov.uk/files//statistics/covid19/covid-deaths-methodology-excess-deaths.pdf):

> For 2022, the excess deaths calculation would usually incorporate the mortality data for the most recent 5 year period (2017 – 2021). However, given the unusual nature of the 2020 data and the decision previously taken for 2021, it has been agreed with ONS that the excess deaths calculation will compare 2022 mortality data with the average of 2016, 2017, 2018, 2019 and 2021, and therefore continue to exclude 2020.
Similarly for 2023, the five year average will again leave out 2020 and will be built from the data from 2017, 2018, 2019, 2021 and 2022.

'Historical' here refers to the 5-year period preceding each year, useful for a historic comparison. The historic period depedning on year has been defined as:

| year|historic_period      |
|----:|:--------------------|
| 2020|2015-2019            |
| 2021|2015-2019            |
| 2022|2016-2019, 2021      |
| 2023|2017-2019, 2021-2022 |

There is a second sense of 'historical' here, which denotes cause of death figures for 2020 to 2022; these were only provided as 5-year averages for 2015-2019, whereas we have a per-cause breakdown from 2020 onwards - to combine these, a 5-year historical average needs to be recomputed using a combination of the 2015-2019 data and the 2020+ data (using weighted averages). This is done in [the later section recomputing historical averages for causes of death](#historical-annual-average-by-cause).

# Historical averages, minima & maxima

Due to the counting of weeks, most years have 52 weeks with some years (2015 & 2020) having 53. Therefore we need to decide how to deal with week 53 in 2020 since only 2015 has a corresponding weekly count. The solution used here is to calculate a weekly rate for week 53 from week 53 in 2015, and week 52 in 2016-2019 - this represents the "last week of the year", which includes December 31st.

Since 2021, the 5-year historic context has been amended to exclude 2020 (see [above section on historical cause of death data](#historical-causes-of-death)). Because of this, the weekly average needs to be recalculated depending on which historic period is of interest.

## Historic periods by year

This is a reference table for the year of data, its historic period, and a nested list of vectors of years of historical data.

```{r}
historic_period_by_year <-
  tibble(year = 2020:latest_year) %>%
  mutate(
    historic_period = historic_period_depending_on_year(year),
    years_in_historic_period = map(.x = as.integer(str_sub(historic_period, 1, 4)), .f = generate_5_years_historic_period_excl_2020)
    )
```

## Historic averages, overall

```{r, warning=FALSE}
## helper functions to filter historic data for different 
get_historic_and_current_data_from_year_range <- function(historic_data, current_data, year_range) {
  year_range <- year_range[year_range!=2020]  # remove 2020 from range
  historic_range <- year_range[year_range < 2020]
  current_range <- year_range[year_range >= 2020]
  bind_rows(
    historic_data %>% 
      filter(year %in% historic_range),
    current_data %>% 
      filter(year %in% current_range) %>% 
      rename(number_of_deaths = deaths_all_causes)  # also rename to historic data format
  )
}

historical_weekly_average_overall <-
  pmap_dfr(
    .l = historic_period_by_year,
    .f = function(year, historic_period, years_in_historic_period) {
      get_historic_and_current_data_from_year_range(
        historic_data = 
          weekly_deaths_2015_2019_overall %>%
          filter(year!="average") %>%
          mutate(year = as.integer(year))
          ,
        current_data = 
          weekly_deaths_overall %>%
          filter(age == "all" & sex == "all"),
        year_range = years_in_historic_period
      ) %>%
      group_by(year, week_number) %>%  # remove week 53 if it was 0 for all places  - this means there was no week 53!
      filter(!all(number_of_deaths==0 & week_number==53)) %>%
      ungroup %>%
      create_week_53_data_for_years_with_52_weeks %>%
      group_by(week_number,place_of_death) %>% 
      summarise(
        across(number_of_deaths, .fns = list(mean=mean,min=min,max=max)),
        .groups = "drop"
      ) %>% 
      rename_all(~str_replace_all(.x, pattern = "number_of_deaths", replacement = "deaths")) %>%
      mutate(historic_period = historic_period)
    }
  ) %>%
  distinct()

week_53_text <- "Due to the way ISO 8601 counts weeks, only 2015 in the 2015-2019 period had 53 weeks; averages for week 53 were computed from figures for week 53 in 2015 and week 52 for 2016-2019."
```

## Historic averages, HB, LA, sex, age

```{r}
## helper function to summarise with mean, min, and max
deaths_mean_min_max <- function(data_tbl) {
  data_tbl %>%
    summarise(
      deaths_mean = mean(number_of_deaths),
      deaths_min = min(number_of_deaths),
      deaths_max = max(number_of_deaths),
      .groups = "drop"
    )
}

historical_weekly_average_hb <-
  pmap_dfr(
    .l = historic_period_by_year,
    .f = function(year, historic_period, years_in_historic_period) {
      get_historic_and_current_data_from_year_range(
        historic_data = weekly_deaths_2015_2019_hb, 
        current_data = weekly_deaths_hb,
        year_range = years_in_historic_period
      ) %>%
      create_week_53_data_for_years_with_52_weeks() %>%
      group_by(
        week_number, place_of_death, ref_area
      ) %>%
      deaths_mean_min_max %>%
      mutate(historic_period = historic_period)
    }
  ) %>% 
  distinct  # the 2020 and 2021 data get identical historical periods & values, so I remove duplicates

historical_weekly_average_la <-
  pmap_dfr(
    .l = historic_period_by_year,
    .f = function(year, historic_period, years_in_historic_period) {
      get_historic_and_current_data_from_year_range(
        historic_data = weekly_deaths_2015_2019_la, 
        current_data = weekly_deaths_la,
        year_range = years_in_historic_period
      ) %>%
      create_week_53_data_for_years_with_52_weeks() %>%
      group_by(
        week_number, place_of_death, ref_area
      ) %>%
      deaths_mean_min_max %>%
      mutate(historic_period = historic_period)
    }
  ) %>% 
  distinct  # the 2020 and 2021 data get identical historical periods & values, so I remove duplicates
# 
# historical_weekly_average_sex_age <-
#   pmap_dfr(
#     .l = historic_period_by_year,
#     .f = function(year, historic_period, years_in_historic_period) {
#       get_historic_and_current_data_from_year_range(
#         historic_data = weekly_deaths_2015_2019_sex_age,
#         current_data = weekly_deaths_sex_age,
#         year_range = years_in_historic_period
#       ) %>%
#       create_week_53_data_for_years_with_52_weeks() %>%
#       group_by(
#         week_number, place_of_death, age, sex
#       ) %>%
#       deaths_mean_min_max %>%
#       mutate(historic_period = historic_period)
#     }
#   ) %>%
#   distinct  # the 2020 and 2021 data get identical historical periods & values, so I remove duplicates
# 
# historical_weekly_average_age <-
#   pmap_dfr(
#     .l = historic_period_by_year,
#     .f = function(year, historic_period, years_in_historic_period) {
#       get_historic_and_current_data_from_year_range(
#         historic_data = weekly_deaths_2015_2019_age, 
#         current_data = weekly_deaths_age,
#         year_range = years_in_historic_period
#       ) %>%
#       create_week_53_data_for_years_with_52_weeks() %>%
#       group_by(
#         week_number, place_of_death, age
#       ) %>%
#       deaths_mean_min_max %>%
#       mutate(historic_period = historic_period)
#     }
#   ) %>% 
#   distinct  # the 2020 and 2021 data get identical historical periods & values, so I remove duplicates
# 
# historical_weekly_average_sex <-
#   pmap_dfr(
#     .l = historic_period_by_year,
#     .f = function(year, historic_period, years_in_historic_period) {
#       get_historic_and_current_data_from_year_range(
#         historic_data = weekly_deaths_2015_2019_sex, 
#         current_data = weekly_deaths_sex,
#         year_range = years_in_historic_period
#       ) %>%
#       create_week_53_data_for_years_with_52_weeks() %>%
#       group_by(
#         week_number, place_of_death, sex
#       ) %>%
#       deaths_mean_min_max %>%
#       mutate(historic_period = historic_period)
#     }
#   ) %>% 
#   distinct  # the 2020 and 2021 data get identical historical periodfs & values, so I remove duplicates
```

## Historic averages, annual average

Here I'm computing the 5-year annual average for various groupings of variables, and across the various possible historic periods. Instead of taking the weekly averages for a historic period computed in the above section, I recompute this separately from the raw data for historic periods and current data.

```{r}
## helper function to compute annual average per historic period, per place of
## death, and per various grouping variables as needed
compute_historical_annual_average_deaths <- function(data_tbl, ...) {
  grouping_vars <- enquos(...)
  data_tbl %>%
    group_by(year, place_of_death, historic_period, !!!grouping_vars) %>%
    summarise(  # calculate annual totals first...
    number_of_deaths = sum(number_of_deaths), 
    .groups = "keep"
  ) %>%
  ungroup(year) %>%  # remove year as grouping to average over year now
  summarise(
    annual_deaths_mean = mean(number_of_deaths),
    annual_deaths_min = min(number_of_deaths),
    annual_deaths_max = max(number_of_deaths),
    .groups = "drop"
  )
}

## Note, there was an error in these before, where I erroneously omitted 2015
## figures from the annual averages computation for 2015-2019; this had minimal
## effect on the mean/min/max values because the algorithm used averaged over
## the available years (as opposed to e.g. dividing 4 years of data by 5)

# the below commented out section is no longer used, but I've left it to
# demonstrate the error in previous data - I think I started 'splitting' the
# 2015-2019 afrom the 2016-2021 historic periods but didn't end up implementing
# both, and instead left a confusing entry that compute the 2016-2019 period
# only.
# make_2016_to_2019_and_2021_data_with_historic_periods <- function(historical_data, current_data) {
#   bind_rows(
#     historical_data %>% 
#       filter(year %in% 2016:2019) %>%
#       mutate(historic_period = historic_period_depending_on_year(2020)),
#     current_data %>% 
#       filter(year == 2021) %>% 
#       rename(number_of_deaths = deaths_all_causes) %>%
#       mutate(historic_period = historic_period_depending_on_year(2022))
#   )
# }

historical_annual_average_overall <-
  pmap_dfr(
    .l = historic_period_by_year,
    .f = function(year, historic_period, years_in_historic_period) {
      get_historic_and_current_data_from_year_range(
        historic_data = 
          weekly_deaths_2015_2019_overall %>%  # a quirk of the weekly deaths 'overall' data
          filter(year != "average") %>%
          mutate(year = as.integer(year)), 
        current_data = 
          weekly_deaths_overall %>%
          filter(sex == "all" & age == "all"),
        year_range = years_in_historic_period
      ) %>%
        mutate(historic_period = historic_period)
    }
  ) %>%
  distinct %>%  # 2020 and 2021 use the same historic period, so I remove duplicate entries at this point
  compute_historical_annual_average_deaths()

historical_annual_average_hb <-
  pmap_dfr(
    .l = historic_period_by_year,
    .f = function(year, historic_period, years_in_historic_period) {
      get_historic_and_current_data_from_year_range(
        historic_data = weekly_deaths_2015_2019_hb, 
        current_data = weekly_deaths_hb, 
        year_range = years_in_historic_period
      ) %>%
        mutate(historic_period = historic_period)
    }
  ) %>%
  distinct %>%  # 2020 and 2021 use the same historic period, so I remove duplicate entries at this point
  compute_historical_annual_average_deaths(ref_area)

historical_annual_average_la <-
  pmap_dfr(
    .l = historic_period_by_year,
    .f = function(year, historic_period, years_in_historic_period) {
      get_historic_and_current_data_from_year_range(
        historic_data = weekly_deaths_2015_2019_la, 
        current_data = weekly_deaths_la, 
        year_range = years_in_historic_period
      ) %>%
        mutate(historic_period = historic_period)
    }
  ) %>%
  distinct %>%  # 2020 and 2021 use the same historic period, so I remove duplicate entries at this point
  compute_historical_annual_average_deaths(ref_area)

# historical_annual_average_sex_age <-
#   pmap_dfr(
#     .l = historic_period_by_year,
#     .f = function(year, historic_period, years_in_historic_period) {
#       get_historic_and_current_data_from_year_range(
#         historic_data = weekly_deaths_2015_2019_sex_age, 
#         current_data = weekly_deaths_sex_age, 
#         year_range = years_in_historic_period
#       ) %>%
#         mutate(historic_period = historic_period)
#     }
#   ) %>%
#   distinct %>%  # 2020 and 2021 use the same historic period, so I remove duplicate entries at this point
#   compute_historical_annual_average_deaths(sex, age)
# 
# historical_annual_average_sex <-
#   pmap_dfr(
#     .l = historic_period_by_year,
#     .f = function(year, historic_period, years_in_historic_period) {
#       get_historic_and_current_data_from_year_range(
#         historic_data = weekly_deaths_2015_2019_sex, 
#         current_data = weekly_deaths_sex, 
#         year_range = years_in_historic_period
#       ) %>%
#         mutate(historic_period = historic_period)
#     }
#   ) %>%
#   distinct %>%  # 2020 and 2021 use the same historic period, so I remove duplicate entries at this point
#   compute_historical_annual_average_deaths(sex)
# 
# historical_annual_average_age <-
#   pmap_dfr(
#     .l = historic_period_by_year,
#     .f = function(year, historic_period, years_in_historic_period) {
#       get_historic_and_current_data_from_year_range(
#         historic_data = weekly_deaths_2015_2019_age, 
#         current_data = weekly_deaths_age, 
#         year_range = years_in_historic_period
#       ) %>%
#         mutate(historic_period = historic_period)
#     }
#   ) %>%
#   distinct %>%  # 2020 and 2021 use the same historic period, so I remove duplicate entries at this point
#   compute_historical_annual_average_deaths(age)
```

## Check weekly historic averages against NRS computed data

```{r}
weekly_deaths_overall %>%
  filter(sex=="all" & age=="all") %>%
  mutate(historic_period = historic_period_depending_on_year(year)) %>%
  left_join(historical_weekly_average_overall) %>%
  select(year, week_number, place_of_death, deaths_nrs_past_average_all_causes, deaths_mean, historic_period) %>%
  filter(round(deaths_nrs_past_average_all_causes) != round(deaths_mean)) %>%
  knitr::kable(caption = "Subset of weekly deaths where calculated mean deaths 2015-2019 differed between the COVID-19 related deaths file and the weekly deaths file (i.e. all other entries had the same figures!). Week 53 was computed as the average of deaths in week 53 in 2015 and weeks 52 in 2016-2019. For all the other datasets the mean deaths computed agree!")
```

# Historic average by cause of death

Cause of death data comes on a per-year basis, with the current year's data including the preceding year; at some point mid-year, this changes, and only the current year's data are included, and the preceding year's data need to be loaded from the last release that included them. (e.g. mid-2022, causes of death for 2021 stopped being released alongside 2022 data, so I load them from the last 2022 release that included 2021 causes of death also). They therefore come as one dataset per year, with the current year as a separate dataset. The weekly figures for cause of death are included alongside the 5-year historic average for that week except for Covid, which has been set to NA for all data up and including 2023 data as of February 2023.

In the `2015-2019` period there were no Covid-19 deaths, so the historical average should be `NA`, which can possibly be interpreted as 0 for the purposes of comparing a year to the historical average.

For the `2016-2019, 2021` period the average number of covid deaths should be equivalent to the deaths in 2021 - we wouldn't want to make an average between 4 years of 0 deaths and one year with any deaths! Similarly if the period were `2017-2019, 2021-2022`, we would compute the average for 2021-2022, and so on. The year 2020 is considered exceptional and so is not used for computing the historical averages in 2021 and onwards.

```{r}
# TODO: include 2023 data once it's reported separately

# TODO: I still need to separately process the most recent year because the average values aren't provided

## here I separately merge the datasets for cause of death data from different years
## I treat covid deaths separately to recalculate a realistic estimate of average deaths in prior years
weekly_deaths_cause_of_death_noncovid <-
  bind_rows(
    weekly_deaths_cause_of_death_2020,
    weekly_deaths_cause_of_death_2021,
    weekly_deaths_cause_of_death_2022,
    weekly_deaths_cause_of_death_2023_onwards
  ) %>%
  filter(cause_of_death != "COVID-19")

weekly_deaths_cause_of_death_covid_before_processing <-
  bind_rows(
    weekly_deaths_cause_of_death_2020,
    weekly_deaths_cause_of_death_2021,
    weekly_deaths_cause_of_death_2022,
    weekly_deaths_cause_of_death_2023_onwards
  )

weekly_deaths_cause_of_death_covid <-
  weekly_deaths_cause_of_death_covid_before_processing %>%
  filter(cause_of_death == "COVID-19") %>%
  group_by(year) %>%
  group_map(
    .f = function(data_tbl, year_tbl) {
      current_year <- year_tbl$year  # get year from the function arguments
      year_range <- infer_year_range_for_calculating_covid_deaths(current_year)  # work out which years of data to use
      weekly_means <-  
        weekly_deaths_cause_of_death_covid_before_processing %>%  # bind again weekly deaths
        filter(year %in% year_range) %>%  # filter the covid data to the relevant years to compute from
        create_week_53_data_for_years_with_52_weeks() %>%  # create 53rd week if needed
        group_by(place_of_death, cause_of_death, week_number) %>%
        summarise(
          deaths_mean = mean(number_of_deaths, na.rm=TRUE),
          # year_range = paste0(year_range,collapse=","),  # for testing
          .groups = "drop"
        )
      data_tbl %>%
        mutate(year = current_year) %>%  # this gets removed in the grouping, so I re-add it here
        select(-deaths_mean) %>%  # remove the NA/0 mean deaths
        create_week_53_data_for_years_with_52_weeks() %>%  # create 53rd week if needed
        left_join(weekly_means, by = c("place_of_death", "cause_of_death", "week_number"))  # merge the mean deaths
    }
  ) %>%
  bind_rows()  # merge into single tibble from list

## re-merge non-covid and covid deaths again
# TODO: make CoD, PoD factors
weekly_deaths_cause_of_death <-
  bind_rows(
    weekly_deaths_cause_of_death_noncovid,
    weekly_deaths_cause_of_death_covid
  ) %>%
  mutate(historic_period = historic_period_depending_on_year(year))

## note that these might differ from actual historic figures a bit because of
## my hack to create a fake 53rd week in years with only 52 weeks for easier 
## comparison

# TODO: I get different numbers of 5-year annual average deaths between 2020 and 2021, how come?
historical_annual_average_cause_of_death <-
  weekly_deaths_cause_of_death %>%
  group_by(year, historic_period, place_of_death, cause_of_death) %>%
  summarise(
    annual_deaths_mean = sum(deaths_mean),
    .groups = "drop"
  ) %>% 
  filter(year!=2020) %>%  # 2020 and 2021 should have identical annual means so I remove one %>%
  select(-year)

## checking discrepancy:
## I get different numbers of 5-year annual average deaths between 2020 and 2021, how come?
weekly_deaths_cause_of_death %>% filter(year %in% 2020:2021) %>% select(place_of_death, cause_of_death, year, week_number, deaths_mean) %>%
  pivot_wider(names_from = year, values_from = deaths_mean) %>%
  filter(`2020`!=`2021`) %>%
  arrange(desc(week_number)) %>%
  summarise(across(matches("\\d{4}"), ~sum(.x, na.rm=TRUE)))
## looks like 2020 and 2021 don't actually use the same 5-year weekly average deaths by cause!!!
# TODO: act on this knowledge...
```

# Merge weekly deaths with historic death rates

```{r}
## compute rolling average
merged_deaths_overall <-
  weekly_deaths_overall %>%
  filter(sex=="all" & age=="all") %>%
  mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
  inner_join(historical_weekly_average_overall %>% mutate(sex = "all", age = "all"),  # only merge for aggregated data b/c that's all we have!
  by = c("place_of_death", "week_number", "sex", "age", "historic_period"))

merged_deaths_hb <-
  weekly_deaths_hb %>%
  mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
  inner_join(historical_weekly_average_hb, by = c("week_number","ref_area","place_of_death", "historic_period"))

merged_deaths_la <-
  weekly_deaths_la %>%  # use data with weekly movingincluded
  mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
  inner_join(historical_weekly_average_la, by = c("week_number","ref_area","place_of_death", "historic_period"))
# 
# merged_deaths_sex_age <-
#   weekly_deaths_sex_age %>%  # use data with weekly movingincluded
#   mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
#   inner_join(historical_weekly_average_sex_age, by = c("place_of_death", "week_number", "sex", "age", "historic_period"))
# 
# merged_deaths_sex <-
#   weekly_deaths_sex %>%
#   mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
#   inner_join(historical_weekly_average_sex, by = c("place_of_death", "week_number", "sex", "historic_period"))
# 
# merged_deaths_age <-
#   weekly_deaths_age %>%
#   mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
#   inner_join(historical_weekly_average_age, by = c("place_of_death", "week_number", "age", "historic_period"))

## Note: previously I computed the historical average separately fro mthe weekly cause of death data
## Now it's already computed in the same dataset, so I just assign it here, no joining needed
merged_deaths_cause_of_death <-
  weekly_deaths_cause_of_death
# merged_deaths_cause_of_death <-
#   weekly_deaths_cause_of_death %>%
#   mutate(historic_period = historic_period_depending_on_year(year)) %>%  # add historic period so we match with the correct historic period data
#   left_join(  # left_join so that the historical Covid-19 rates are blank!
#     historical_average_cause_of_death %>% select(-week_number_run_over), 
#     by = c("place_of_death", "week_number", "cause_of_death", "historic_period", "year", "date_w_c")
#   )
```

## Weekly deaths including past weekly deaths

For this dataset, we simply `bind_rows` weekly deaths by year and place of death with past weekly deaths by year and location. Note that past data for years with 52 weeks includes an entry of 0 deaths for week 53, so we'll remove them, otherwise the date computation is thrown off!

```{r}
weekly_deaths_current_and_past <-
  bind_rows(
    weekly_deaths_overall %>%
    filter(sex=="all" & age=="all" & place_of_death!="All") %>%  # keep only aggregated values over sex & age
    select(year, week_number, week_number_run_over, place_of_death, matches("deaths"), -deaths_nrs_past_average_all_causes)
    ,
    weekly_deaths_2015_2019_overall %>%
      rename(deaths_all_causes = number_of_deaths) %>%
      filter(place_of_death!="All" & year!="average") %>%  # remove aggregated entries
      mutate(year = as.integer(year)) %>%
      filter(!(year %in% 2016:2019 & week_number==53))  # remove week 53 entries from years 2016-2019
  ) %>%
  mutate(date_w_c = compute_start_date_from_week_number(week_number = week_number, year_number = year))
```

# Annual deaths for 2020 & 2021

For visualising data using maps, we need to compute data for the entire year. Note: the non-covid and covid-related deaths in 2020 don't add up to all cause deaths due to the covid status only starting in week 12!

```{r}
compute_total_deaths_in_year <- function(data_tbl, ...) {
  grouping_vars <- enquos(...)
  data_tbl %>%
    group_by(year, place_of_death, !!!grouping_vars) %>%
    summarise(across(.cols = c(deaths_all_causes, deaths_covid_related, deaths_non_covid), .fns = ~sum(.x,na.rm=TRUE)), .groups = "drop") %>%  # compute sums of all types of death
    mutate(deaths_all_causes_since_covid = deaths_covid_related + deaths_non_covid)  # compute a smaller number of all-cause deaths since covid started (over 2020, covid- and non-covid-related deaths don't add up to all-cause deaths because pre-covid deaths are included also!)
}

compute_ratio_of_annual_deaths_to_historical <- function(data_tbl) {
  data_tbl %>%
    mutate(
      ratio_annual_deaths_to_historical_mean = deaths_all_causes / annual_deaths_mean,
      ratio_annual_deaths_to_historical_min = deaths_all_causes / annual_deaths_min,
      ratio_annual_deaths_to_historical_max = deaths_all_causes / annual_deaths_max
    )
}

## helper function to remove rows of data where the historic period is inappropriate
remove_invalid_historic_period_for_year <- function(data_tbl) {
  data_tbl %>%
    filter(!(historic_period == "2016-2019, 2021" & year < 2022))
}

total_deaths_overall <-
  merged_deaths_overall %>%
  compute_total_deaths_in_year() %>%
  left_join(historical_annual_average_overall, by = c("place_of_death")) %>%
  compute_ratio_of_annual_deaths_to_historical %>%
  remove_invalid_historic_period_for_year()

total_deaths_la <-
  merged_deaths_la %>%
  compute_total_deaths_in_year(ref_area) %>%
  left_join(historical_annual_average_la, by = c("place_of_death", "ref_area")) %>%
  compute_ratio_of_annual_deaths_to_historical %>%
  remove_invalid_historic_period_for_year()

total_deaths_hb <-
  merged_deaths_hb %>%
  compute_total_deaths_in_year(ref_area) %>%
  left_join(historical_annual_average_hb, by = c("place_of_death", "ref_area")) %>%
  compute_ratio_of_annual_deaths_to_historical %>%
  remove_invalid_historic_period_for_year()

# total_deaths_sex <-
#   merged_deaths_sex %>%
#   compute_total_deaths_in_year(sex) %>%
#   left_join(historical_annual_average_sex, by = c("place_of_death", "sex")) %>%
#   compute_ratio_of_annual_deaths_to_historical %>%
#   remove_invalid_historic_period_for_year()
# 
# total_deaths_age <-
#   merged_deaths_age %>%
#   compute_total_deaths_in_year(age) %>%
#   left_join(historical_annual_average_age, by = c("place_of_death", "age")) %>%
#   compute_ratio_of_annual_deaths_to_historical %>%
#   remove_invalid_historic_period_for_year()
```

## Annual deaths by cause of death

```{r}
## bespoke computation because format is different... deaths split by cause!
total_deaths_cause_of_death <-
  merged_deaths_cause_of_death %>%
  group_by(year, place_of_death, cause_of_death, historic_period) %>%
  summarise(deaths = sum(number_of_deaths, na.rm = TRUE), .groups="drop") %>%
  left_join(historical_annual_average_cause_of_death, by = c("place_of_death", "cause_of_death", "historic_period")) %>%
  mutate(
    ratio_annual_deaths_to_historical_mean = if_else(  # only compute this for complete years, we can't compare 2022 to annual average!
      condition = year == 2022,
      true = NA_real_,
      false = deaths / annual_deaths_mean
    )
  )

current_and_historical_deaths_by_cause_and_location <-
  bind_rows(
    ## 2020-2022 separately
    total_deaths_cause_of_death %>%
      select(year, place_of_death, cause_of_death, deaths) %>%
      mutate(year = as.character(year)),
    ## historical data
    total_deaths_cause_of_death %>%  # these are for 2020-2021, but we use the historical mean
      filter(year == 2020) %>%
      select(year, place_of_death, cause_of_death, deaths = annual_deaths_mean) %>%
      mutate(year = "2015-2019")
  ) %>%
  filter(place_of_death != "All") %>%
  group_by(year, cause_of_death) %>%
  mutate(proportion = deaths / sum(deaths)) %>%
  ungroup %>%
  arrange(year, cause_of_death)
```

## Historical average weekly deaths, including for incomplete year 2022

```{r}
merged_deaths_hb %>%
  group_by(year, ref_area, place_of_death, historic_period) %>%
  summarise(
    deaths_all_causes = sum(deaths_all_causes, na.rm = TRUE),
    deaths_mean = sum(deaths_mean, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(proportion_home_deaths_increase = (deaths_all_causes/deaths_mean) - 1)

# TODO: fix below to include historic_period
# proportions_of_historical_deaths_by_place <-
#   historical_weekly_average_overall %>% 
#   group_by(year, place_of_death) %>% summarise(number_of_deaths = sum(number_of_deaths), .groups="drop") %>%
#   filter(place_of_death != "All" & !is.na(as.integer(year))) %>%
#   group_by(year) %>%
#   mutate(
#     total_deaths = sum(number_of_deaths)
#   ) %>%
#   ungroup %>%
#   mutate(proportion_of_total = number_of_deaths / total_deaths, year=as.integer(year))
# 
# proportions_of_current_deaths_by_place <-
#   total_deaths_overall %>%
#   select(year, place_of_death, number_of_deaths = deaths_all_causes) %>%
#   filter(place_of_death != "All") %>%
#   group_by(year) %>%
#   mutate(
#     total_deaths = sum(number_of_deaths)
#   ) %>%
#   ungroup %>%
#   mutate(proportion_of_total = number_of_deaths / total_deaths)
# 
# 
# merged_proportions_of_deaths_by_place <-
#   bind_rows(
#     proportions_of_historical_deaths_by_place,
#     proportions_of_current_deaths_by_place
#   )
```

# Constants about latest data available

These are used for saying in plots what data were included e.g. weeks 1-12 in 2023.

There is as of April 2023 a split in the availability of data:

- Data are being continually updated for the all age all sex all HB/LA dataset
- Data split by age, sex, HB or LA stopped being updated (technically they are still available but not without place of death data)

Therefore there are two separate most recent available data, I'll call them 'basic' and 'disaggregated'/

```{r}
## constants for use in graphs & annotations
weeks_available_2024 <- max(merged_deaths_overall$week_number[merged_deaths_overall$year==2024])
most_recent_date_available_2024 <- max(merged_deaths_overall$date_w_c)

earliest_date <- min(merged_deaths_overall$date_w_c)
current_year <- isoyear(most_recent_date_available_2024)

weeks_available_2024_disaggregated <- max(merged_deaths_cause_of_death$week_number[merged_deaths_cause_of_death$year==2024])
most_recent_date_available_2024_disaggregated <- max(merged_deaths_cause_of_death$date_w_c)
```

# Ordering geographical data by population size

The latest population estimates by HB/LA are available [from NRS Scotland here](https://www.opendata.nhs.scot/dataset/population-estimates).

I'm using the estimates for 2020 (based on 2019 data) to compute the relative sizes of health boards!

```{r}
population_estimate_la_2019 <- read_csv(file = "./population_estimates/ca2019_pop_est_29062021.csv") %>%
  select(Year, CA, Sex, AllAges) %>%
  clean_names() %>%
  make_human_readable_ref_area(ref_area = ca) %>%
  filter(year==2020) %>%
  filter(ca != "Scotland") %>%
  filter(sex == "All") %>%
  rename(la = ca)

population_estimate_hb_2019 <- read_csv(file = "./population_estimates/hb2019_pop_est_29062021.csv") %>%
  select(Year, HB, Sex, AllAges) %>%
  clean_names() %>%
  make_human_readable_ref_area(ref_area = hb) %>%
  filter(year==2020) %>%
  filter(hb != "Scotland") %>%
  filter(sex == "All")
```

## Compute order

```{r}
order_la_by_population <-
  population_estimate_la_2019 %>%
  arrange(desc(all_ages)) %>%
  .$la

order_hb_by_population <-
  population_estimate_hb_2019 %>%
  arrange(desc(all_ages)) %>%
  .$hb
```

# Calculate case counts: weekly counts to match weekly death reporting

Note: the cumulative case counts have already been calculated, we just need to subset them to match the weeks. This is not trivial - the death counts cover an entire week starting from the "w/c date"; the cumulative cases for that week could be taken either at the start, the end of the week, or the start of the following week even!

However, when calculating weekly positive cases we can simply take the maximum of that week's cumulative count - that will be the cumulative count at the end of the week!

Note also that the `crude_rate_positive_per100k` represents the cumulative rate!

## Calculating population size of areas

Update 7 Mr 2022: at some point the file structure changed, and the crude positive rate per 100k was removed... however, the crude mortality rate remains, so we can infer the population from the `crude_rate_deaths` and `cumulative_deaths` figures, and compute the `crude_rate_positive` from those.

The trouble is that early on in the data, when `cumulative_deaths` and `crude_rate_deaths` are both 0, we can't infer the population size. If the population size is constant in the data, we can calculate it for each `ref_area` (HB or LA) and use that as a lookup table.

It looks like the same population estimate was used for data 2020-2022, and it is the **mid-year 2019 estimate**.

## Importing & calculating case trends

```{r}
weekly_covid_cases_hb <-
  cases_hb %>%
  left_join(
    population_estimate_hb_2019 %>%
    add_scotland_total_to_population_estimate %>%
    select(ref_area = hb, population_size = all_ages)
    ) %>%
  group_by(year, week_number, ref_area) %>%
  summarise(
    weekly_positive = sum(daily_positive, na.rm = TRUE),
    cumulative_positive = max(cumulative_positive, na.rm = TRUE),
    # crude_rate_positive_per100k = max(crude_rate_positive),  # this column is no longer in the data, so we compute it below...
    crude_rate_positive_per100k = max(cumulative_positive / population_size * 1e5, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    date_w_c = compute_start_date_from_week_number(week_number = week_number, year = year),
    week_number_run_over = if_else(year == 2021, 53L + week_number, week_number)
    )

## extract Scotland-wide data and remove from HB split
weekly_covid_cases_overall <- weekly_covid_cases_hb %>% filter(ref_area=="Scotland")
weekly_covid_cases_hb <- weekly_covid_cases_hb %>% filter(ref_area!="Scotland")

weekly_covid_cases_la <-
  cases_la %>%
  left_join(
    population_estimate_la_2019 %>%
    add_scotland_total_to_population_estimate %>%
    select(ref_area = la, population_size = all_ages)
    ) %>%
  group_by(year, week_number, ref_area) %>%
  summarise(
    weekly_positive = sum(daily_positive, na.rm = TRUE),
    cumulative_positive = max(cumulative_positive, na.rm = TRUE),
    # crude_rate_positive_per100k = max(crude_rate_positive),  # this column is no longer in the data, so we compute it below...
    crude_rate_positive_per100k = max(cumulative_positive / population_size * 1e5, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    date_w_c = compute_start_date_from_week_number(week_number = week_number, year = year),
    week_number_run_over = if_else(year == 2021, 53L + week_number, week_number)
    )
```

# Comparisons of annual rates with current rates

```{r}
table_current_deaths_compared_to_past <-
  bind_rows(
    ## all available data
    merged_deaths_overall %>%
      filter(age == "all" &
               sex == "all"
             ) %>% 
      group_by(year, place_of_death, week_number) %>%
      summarise(
        deaths_all_causes = sum(deaths_all_causes),
        # deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
      mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
      left_join(
        weekly_deaths_2015_2019_overall %>% filter(year != "average") %>% group_by(week_number, place_of_death) %>% summarise(
                                           average_weekly_deaths = sum(number_of_deaths) / 5,
                                           .groups = "drop"
                                         ),
        by = c("week_number", "place_of_death")
      ) %>% 
      group_by(place_of_death) %>%
      summarise(
        period = glue::glue("Data from week 1, 2020 to week {weeks_available_2024}, 2022"),
        historic_period = "2015-2019",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        mean_annual_deaths_2015_2019 = sum(average_weekly_deaths),
        .groups = "drop"
      )
    ,
    ## 2020 only
    merged_deaths_overall %>%
      filter(age == "all" &
               sex == "all" &
               year == 2020
             ) %>% 
      group_by(place_of_death, historic_period) %>%
      summarise(
        period = "2020",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
      mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      historical_annual_average_overall %>% 
        select(place_of_death, mean_annual_deaths_2015_2019 = annual_deaths_mean, historic_period),
      by = c("place_of_death","historic_period")
    )
    ,
    ## 2020 from week 12 onwards (including week 12) - pandemic part of 2020
    merged_deaths_overall %>%
      filter(
        age == "all" &
          sex == "all" &
          year == 2020 & week_number >= 12
      ) %>%
      group_by(place_of_death) %>%
      summarise(
        period = "2020, week 12 - week 53",
        historic_period = "2015-2019",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
    mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      weekly_deaths_2015_2019_overall %>% 
        filter(week_number >= 12 & year != "average") %>%
        group_by(place_of_death) %>%
        summarise(mean_annual_deaths_2015_2019 = sum(number_of_deaths) / 5),
      by = "place_of_death"
    )
    ,
    ## 2021 only
    merged_deaths_overall %>%
      filter(age == "all" &
               sex == "all" &
               year == 2021
             ) %>% 
      group_by(place_of_death) %>%
      summarise(
        period = "2021",
        historic_period = "2015-2019",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
      mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      historical_annual_average_overall %>% 
        select(place_of_death, mean_annual_deaths_2015_2019 = annual_deaths_mean, historic_period),
      by = c("place_of_death","historic_period")
    )
    ,
    ## 2022
    merged_deaths_overall %>%
      filter(age == "all" &
               sex == "all" &
               year == 2022
             ) %>% 
      group_by(place_of_death) %>%
      summarise(
        period = "2022",
        historic_period = "2015-2019",
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
      mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      historical_annual_average_overall %>% 
        select(place_of_death, mean_annual_deaths_2015_2019 = annual_deaths_mean, historic_period),
      by = c("place_of_death","historic_period")
    ),
    ## 2024 data so far
    merged_deaths_overall %>%
      filter(
        age == "all" &
          sex == "all" &
          year == 2024
      ) %>%
      group_by(place_of_death, week_number) %>%
      summarise(
        deaths_all_causes = sum(deaths_all_causes),
        deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
        .groups = "drop"
      ) %>% 
    mutate(deaths_non_covid = deaths_all_causes - deaths_covid_related) %>%
    left_join(
      weekly_deaths_2015_2019_overall %>% filter(year != "average") %>% group_by(week_number, place_of_death) %>% summarise(
                                         average_weekly_deaths = sum(number_of_deaths) / 5,
                                         .groups = "drop"
                                       ),
      by = c("week_number", "place_of_death")
    ) %>% 
    group_by(place_of_death) %>%
    summarise(
      period = glue::glue("2024, up to week {weeks_available_2024}"),
      historic_period = "2015-2019",
      deaths_all_causes = sum(deaths_all_causes),
      deaths_non_covid = sum(deaths_non_covid, na.rm = TRUE),
      deaths_covid_related = sum(deaths_covid_related, na.rm = TRUE),
      mean_annual_deaths_2015_2019 = sum(average_weekly_deaths),
      .groups = "drop"
    )
  ) %>%
  mutate(
    ratio_deaths_all_causes_to_historical = deaths_all_causes / mean_annual_deaths_2015_2019,
    ratio_non_covid_deaths_to_historical = deaths_non_covid / mean_annual_deaths_2015_2019,
    proportion_covid_deaths = deaths_covid_related / deaths_all_causes
  )

write_csv(x = table_current_deaths_compared_to_past, file = file.path(dir_outputs, "annual_death_rates_comparison_with_historical_deaths.csv"))

table_current_deaths_compared_to_past %>% knitr::kable(caption = "Comparing annual home death rate during COVID pandemic to historical deaths at home")
```

# Moving averages: 4 weeks

```{r}
merged_deaths_overall <-
  compute_moving_average_for_current_data(
    current_data = merged_deaths_overall, 
    historical_data = 
      weekly_deaths_2015_2019_overall %>% mutate(year=as.integer(year)) %>% drop_na(year),  # remove the entries with 'average' for year
    num_weeks_to_average = 4)

merged_deaths_hb <-
  compute_moving_average_for_current_data(
    current_data = merged_deaths_hb, 
    historical_data = 
      weekly_deaths_2015_2019_hb, 
    num_weeks_to_average = 4)

merged_deaths_la <-
  compute_moving_average_for_current_data(
    current_data = merged_deaths_la, 
    historical_data = 
      weekly_deaths_2015_2019_la, 
    num_weeks_to_average = 4)
# 
# merged_deaths_sex <-
#   compute_moving_average_for_current_data(
#     current_data = merged_deaths_sex, 
#     historical_data = 
#       weekly_deaths_2015_2019_sex, 
#     num_weeks_to_average = 4)

# merged_deaths_age <-
#   compute_moving_average_for_current_data(
#     current_data = merged_deaths_age, 
#     historical_data = 
#       weekly_deaths_2015_2019_age, 
#     num_weeks_to_average = 4)

# merged_deaths_sex_age <-
#   compute_moving_average_for_current_data(
#     current_data = merged_deaths_sex_age, 
#     historical_data = 
#       weekly_deaths_2015_2019_sex_age, 
#     num_weeks_to_average = 4)
```

# Annual deaths by place, 2015-2022

```{r}
merged_proportions_of_deaths_by_place <-
  bind_rows(
    ## 2015-2019
    weekly_deaths_2015_2019_overall %>% 
    group_by(year, place_of_death) %>% summarise(number_of_deaths = sum(number_of_deaths), .groups="drop") %>%
    filter(place_of_death != "All" & !is.na(as.integer(year))) %>%
    group_by(year) %>%
    mutate(
      total_deaths = sum(number_of_deaths)
    ) %>%
    ungroup %>%
    mutate(
      proportion_of_total = number_of_deaths / total_deaths, 
      year=as.integer(year)
      ),
    ## 2020 onwards
    total_deaths_overall %>%
    select(year, place_of_death, number_of_deaths = deaths_all_causes) %>%
    distinct %>%  # remove duplicate entries for 2022 due to having multiple possible historic periods
    filter(place_of_death != "All") %>%
    group_by(year) %>%
    mutate(
      total_deaths = sum(number_of_deaths)
    ) %>%
    ungroup %>%
    mutate(proportion_of_total = number_of_deaths / total_deaths)
  ) %>%
  (function(x) {
    bind_rows(
      x,
      x %>%
        group_by(year) %>%
        summarise(
          place_of_death = "All",
          number_of_deaths = sum(number_of_deaths, na.rm = TRUE),
          total_deaths = unique(total_deaths),
          proportion_of_total = 1
          )
    )
  }) %>%
  mutate(place_of_death = factor(place_of_death, levels = order_of_place_of_death_levels[c(1,3,2,4,5)])) %>%
  arrange(year, place_of_death)

write_csv(x = merged_proportions_of_deaths_by_place, file = file.path(dir_outputs, "annual_deaths_by_place.csv"))
```

## By HB/LA

```{r}
proportions_of_deaths_by_place_by_geography <-
  bind_rows(
    ## historic deaths
    bind_rows(  # merge HB & LA past deaths, and tag them appropriately
      weekly_deaths_2015_2019_hb %>% mutate(geography = "hb"),
      weekly_deaths_2015_2019_la %>% mutate(geography = "la")
    ) %>%
    group_by(year, place_of_death, ref_area, geography) %>% summarise(number_of_deaths = sum(number_of_deaths, na.rm = TRUE), .groups="drop") %>%  # compute the total number of yearly deaths by place of death & geography
    filter(place_of_death != "All" & !is.na(as.integer(year))) %>%  # remove blank/irrelevant entries
    group_by(year, ref_area, geography) %>%
    mutate(
      total_deaths = sum(number_of_deaths)  # compute total annual deaths across place of death, but within each geographical area & year
    ) %>%
    ungroup %>%
    mutate(proportion_of_total = number_of_deaths / total_deaths, year=as.integer(year))  # compute proportion of deaths in any palce relative to total annual deaths in geographical area
    ,
    ## current deaths
    bind_rows(  # merge HB & LA past deaths, and tag them appropriately
      total_deaths_hb %>% mutate(geography = "hb"),
      total_deaths_la %>% mutate(geography = "la")
    ) %>%
    select(year, place_of_death, number_of_deaths = deaths_all_causes, geography, ref_area) %>%  # keep only relevant columns & rename deaths due to all causes
    group_by(year, place_of_death, ref_area, geography) %>% summarise(number_of_deaths = sum(number_of_deaths), .groups="drop") %>%  # compute the total number of yearly deaths by place of death & geography
    filter(place_of_death != "All" & !is.na(as.integer(year))) %>%  # remove blank/irrelevant entries
    group_by(year, ref_area, geography) %>%
    mutate(
      total_deaths = sum(number_of_deaths)  # compute total annual deaths across place of death, but within each geographical area & year
    ) %>%
    ungroup %>%
    mutate(proportion_of_total = number_of_deaths / total_deaths, year=as.integer(year))  # compute proportion of deaths in any palce relative to total annual deaths in geographical area
  )

## write to csv
proportions_of_deaths_by_place_by_geography %>%
  write_csv(file = paste0(dir_outputs,"/proportions_of_deaths_by_place_by_geography.csv"))
```

# Deaths by cause, ordered by number of deaths in 2020

```{r}
## get "canonical" order for causes of death in 2020
order_of_causes_of_death <- 
  total_deaths_cause_of_death %>% 
  filter(year == 2020 & place_of_death == "All") %>% 
  arrange(desc(deaths)) %>% 
  pull(cause_of_death)
```

# Causes of death over time, proportion inside and between places

Table of number of deaths by cause, place of death and year, showing proportion of deaths by cause within each place `prop_cod_within_pod` (e.g. Covid deaths were 2% of home deaths) and proportion of deaths by place within each cause `prop_pod_within_cod` (e.g. 70% of all Covid deaths happened in hospital).
Note that because historic data was given averaged over the 5 year period, the historic figures are given as a mean over the period rather than an annual total.

```{r}
table_proportions_of_cause_of_death <-
  bind_rows(
    ## historic periods
    historical_annual_average_cause_of_death %>%
      rename(year = historic_period, deaths = annual_deaths_mean),
    ## current data
    total_deaths_cause_of_death %>%
      select(-c(historic_period,annual_deaths_mean:ncol(.))) %>%
      mutate(year = as.character(year))
  ) %>%
  filter(place_of_death!="All" & cause_of_death!="All") %>%
  group_by(year,place_of_death) %>%
  mutate(prop_cod_within_pod = deaths / sum(deaths)) %>%
  group_by(year,cause_of_death) %>%
  mutate(prop_pod_within_cod = deaths / sum(deaths)) %>%
  ungroup %>%
  (function(data_tbl){
    bind_rows(
      data_tbl,
      ## all cause
      data_tbl %>%
        group_by(year, place_of_death) %>%
        summarise(
          cause_of_death = "All",
          deaths = sum(deaths),
          prop_cod_within_pod = 1,
          .groups = "drop"
        ) %>%
        group_by(year, cause_of_death) %>%
        mutate(prop_pod_within_cod = deaths/sum(deaths)) %>%
        ungroup,
      ## all place of death
      data_tbl %>%
        group_by(year, cause_of_death) %>%
        summarise(
          place_of_death = "All",
          deaths = sum(deaths),
          prop_pod_within_cod = 1,
          .groups = "drop"
        ) %>%
        group_by(year, place_of_death) %>%
        mutate(prop_cod_within_pod = deaths/sum(deaths)) %>%
        ungroup
    )
  }) %>%
  mutate(
    year = factor(year, levels = unique(year)),
    cause_of_death = factor(cause_of_death, levels = order_of_causes_of_death),
    place_of_death = factor(place_of_death, levels = order_of_place_of_death_levels[c(5,1,3,2,4)])
    ) %>%
  arrange(year, place_of_death, cause_of_death)

table_proportions_of_cause_of_death %>%
  write_csv(file = paste0(dir_outputs,"/proportions_of_deaths_by_place_&_cause.csv"))
```

# Save most recent workspace for further analyses

```{r}
# save.image(file = "./workspace.RData", safe = TRUE)  # this is calling .GlobalEnv, which is empty
save(list = ls(all.names = TRUE), file = "./workspace.RData", envir = environment())
```

# Print session info

```{r}
sessionInfo()
```