### 2024-07-25 just_in_case_bnf_medications.qmd
finished incorporating comments from Sarah Mills into the JIC medications list

---
title: "Just-in-case medications list"
author: "Sarah Mills & Jan Savinc"
format: html
---

# Introduction {#sec-introduction}

Here I'll compile the BNF codes corresponding to medicines Sarah Mills compiled as being "Just in case" for use in palliative care/end of life care.

- Strong opioids: Alfentanil, Buprenorphine, Diamorphine hydrochloride, Dihydrocodeine tartrate, Fentanyl, Hydromorphone hydrochloride, Meptazinol, Methadone hydrochloride, Morphine sulfate, Oxycodone hydrochloride, Tramadol hydrochloride
 - Anti-muscarinics: Aclidinium bromide, Alverine citrate, Bupropion hydrochloride, Minims atropine sulfate, Atropine sulfate, Minims cyclopentolate hydrochloride, Cyclopentolate hydrochloride, Dicycloverine hydrochloride, Erdosteine, Fesoterodine fumarate, Flavoxate hydrochloride, Glycopyrrolate bromide, Hyoscine, Procyclidine hydrochloride, Propantheline bromide, Propiverine hydrochloride
 - Anti-emetics: Betahistine hydrochloride, Cyclizine, Domperidone, Granisetron hydrochloride, Levomepromazine hydrochloride, Metoclopramide hydrochloride, Ondansetron hydrochloride, Prochlorperazine maleate
 - Anxiolytics: Buspirone hydrochloride, Chlordiazepoxide hydrochloride, Clobazam, Clomethiazole, Clonazepam, Diazepam, Loprazolam, Lorazepam, Lormetazepam, Midazolam, Nitrazepam, Temazepam, Trazodone hydrochloride, Zopiclone, Zolpidem tartrate

The strategy for these is to look at each medication, look at its formulations, and decide which to include.

Versions:

- v1 is my first attempt at compiling the codes, without a clinician checking over the codes
- v2 includes amendments suggested by Sarah Mills, I've labelled these with 'SM' in the text

# BNF source

One source for a list of all BNF codes is the [Information Services Portal of the NHS Business Services Authority] (https://applications.nhsbsa.nhs.uk/infosystems/data/showDataSelector.do?reportId=126). This is available after putting in a captcha-like password, so I've downloaded the file manually.

File used: `01-01-2024 : 86`

# Setup

```{r}
library(tidyverse)
library(janitor)
```

# Compile codes for JIC medicines

## Read BNF codes

```{r}
bnf_zip_file <- "./bnf/20240701_1719843096151_BNF_Code_Information.zip"  # manually downloaded to this location
csv_file <- as.character(unzip(bnf_zip_file, list = TRUE)[1])  # get the filename from the zip
temp_dir <- tempdir()  # create temporary directory for unzipping to
unzip(zipfile = bnf_zip_file, exdir = temp_dir)  # extract csv file from zip

bnf <- read_csv(file.path(temp_dir, csv_file)) %>% clean_names()  # read csv file
```

## Encode the list of medicines

```{r}
strong_opioids <- "Alfentanil, Buprenorphine, Diamorphine hydrochloride, Dihydrocodeine tartrate, Fentanyl, Hydromorphone hydrochloride, Meptazinol, Methadone hydrochloride, Morphine sulfate, Oxycodone hydrochloride, Tramadol hydrochloride"

anti_muscarinics <- "Aclidinium bromide, Alverine citrate, Bupropion hydrochloride, Minims atropine sulfate, Atropine sulfate, Minims cyclopentolate hydrochloride, Cyclopentolate hydrochloride, Dicycloverine hydrochloride, Erdosteine, Fesoterodine fumarate, Flavoxate hydrochloride, Glycopyrrolate bromide, Hyoscine, Procyclidine hydrochloride, Propantheline bromide, Propiverine hydrochloride"

anti_emetics <- "Betahistine hydrochloride, Cyclizine, Domperidone, Granisetron hydrochloride, Levomepromazine hydrochloride, Metoclopramide hydrochloride, Ondansetron hydrochloride, Prochlorperazine maleate"

anxiolytics <- "Buspirone hydrochloride, Chlordiazepoxide hydrochloride, Clobazam, Clomethiazole, Clonazepam, Diazepam, Loprazolam, Lorazepam, Lormetazepam, Midazolam, Nitrazepam, Temazepam, Trazodone hydrochloride, Zopiclone, Zolpidem tartrate"

jic <- map2(
  .x = c(strong_opioids, anti_muscarinics, anti_emetics, anxiolytics),
  .y = c("strong opioids", "anti-muscarinics", "anti-emetics", "anxiolytics"),
  .f = function(meds, name) {
    tibble(
      group = name,
      meds = 
        str_split(string = meds, pattern = ",") %>%
        unlist() %>%
        str_trim() %>%
        tolower()
    )
  }
) %>% list_rbind()
```

## Check for non-matches in BNF

A small number of medications don't have exact matches in the BNF when searching by substance:

```{r}
nonmatches <- 
  jic %>%
  anti_join(
    bnf %>% mutate(bnf_chemical_substance = tolower(bnf_chemical_substance)), 
    by = join_by(meds == bnf_chemical_substance)
  )

nonmatches %>% knitr::kable(caption = "Medications that don't match the BNF entries exactly")
```

Note that even exact matches need to be carefully checked as not all formulations are relevant (specifically combination medicines, for example).

## Adding medicines to the list

```{r}
## initialise the blank list of included medicines
jic_medicines <- tibble()

## helper function to subset the bnf list, finding a keyword in the product name
## or chemical substance
search_bnf <- function(keyword) {
  bnf %>%
    filter(
      str_detect(paste0(bnf_product, bnf_chemical_substance), regex(keyword, ignore_case = TRUE))
    )
}

## helper function to add medicines as I go along
add_entries <- function(meds_to_add, bnf_to_add) {
  bind_rows(
    jic_medicines,
    bind_cols(jic %>% filter(meds==meds_to_add), bnf_to_add)
    ) %>%
    distinct  # to precvent accidental duplication along the way
}
```

## strong opioids

### alfentanil

- Alfentanil – I’ve included all formulations, including an “Alfentanil 0.1% nasal spray”
- SM: agree

```{r}
# search_bnf(keyword="alfentanil") %>% View()

jic_medicines <- add_entries(meds = "alfentanil", bnf_to_add = search_bnf(keyword="alfentanil"))
```

### buprenorphine

- •	Burprenorphine – Keep only the analgesics section (0407), not the opioid dependence section (0410) - the latter uses buprenorphine hydrochloride and a buprenorphine hcl/naloxone combination. 
- SM: agree

```{r}
# search_bnf(keyword="buprenorphine") %>% View()

jic_medicines <- add_entries(
  meds = "buprenorphine", 
  bnf_to_add = search_bnf(keyword="buprenorphine") %>%
    filter(bnf_section_code == "0407")
  )
```

### diamorphine hydrochloride

 - I’ve included both the cough suppressants and analgesics sections
 - SM: agree

```{r}
# search_bnf(keyword="diamorphine hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "diamorphine hydrochloride", 
  bnf_to_add = search_bnf(keyword="diamorphine hydrochloride")
  )
```

### dihydrocodeine tartrate

```{r}
# search_bnf(keyword="dihydrocodeine tartrate") %>% View()

jic_medicines <- add_entries(
  meds = "dihydrocodeine tartrate", 
  bnf_to_add = search_bnf(keyword="dihydrocodeine tartrate")
  )
```

### fentanyl

- Keep only the analgesics subsection (0407), not the general anaesthesia and local anaesthesia sections.
- SM: Agree (patches, which are in the analgesics subsection, should be included, and infusion bags/fentanyl citrate, which is in the anaesthesia sections, should be excluded)

```{r}
# search_bnf(keyword="fentanyl") %>% View()

jic_medicines <- add_entries(
  meds = "fentanyl", 
  bnf_to_add = search_bnf(keyword="fentanyl") %>%
    filter(bnf_section_code == "0407")
  )
```

### hydromorphone hydrochloride

```{r}
# search_bnf(keyword="hydromorphone hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "hydromorphone hydrochloride", 
  bnf_to_add = search_bnf(keyword="hydromorphone hydrochloride")
  )
```

### meptazinol

```{r}
# search_bnf(keyword="meptazinol") %>% View()

jic_medicines <- add_entries(
  meds = "meptazinol", 
  bnf_to_add = search_bnf(keyword="meptazinol")
  )
```

### methadone hydrochloride

- Kept the cough suppressants section (0309) and analgesics (0407), not drug dependence (0410).
- SM: agree

```{r}
# search_bnf(keyword="methadone hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "methadone hydrochloride", 
  bnf_to_add = search_bnf(keyword="methadone hydrochloride") %>%
    filter(bnf_section_code %in% c("0309","0407"))
  )
```

### morphine sulfate

- I’ve kept all formulations, including several gels/intrasite gels? 
- SM: agree, gels often used for fungating tumours

```{r}
# search_bnf(keyword="morphine sulfate") %>% View()

jic_medicines <- add_entries(
  meds = "morphine sulfate", 
  bnf_to_add = search_bnf(keyword="morphine sulfate")
  )
```

### oxycodone hydrochloride

- I’ve kept all formulations, except for the combination “Oxycodone hydrochloride/naloxone hydrochloride” 
- SM: agree

```{r}
# search_bnf(keyword="oxycodone hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "oxycodone hydrochloride", 
  bnf_to_add = search_bnf(keyword="oxycodone hydrochloride")
  )
```

### tramadol hydrochloride

```{r}
# search_bnf(keyword="tramadol hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "tramadol hydrochloride", 
  bnf_to_add = search_bnf(keyword="tramadol hydrochloride")
  )
```

## anti-muscarinics

### aclidinium bromide

- Keep only paragraph code 030102 for antimuscarinic bronchodilators, not compound bronchodilator preparations 030104 (Aclidinium bromide/formoterol)
- SM: agree

```{r}
# search_bnf(keyword="aclidinium bromide") %>% View()

jic_medicines <- add_entries(
  meds = "aclidinium bromide", 
  bnf_to_add = search_bnf(keyword="aclidinium bromide") %>%
    filter(bnf_paragraph_code == "030102")
  )
```

### alverine citrate

- SM: not familiar with combination Alverine citrate/simethicone, or Alverine citrate compound preparations (Sterculia 62% / Alverine 0.5% granules and Spasmonal Fibre granules)
- therefore, removed those combinations, only kept preparations with only alverine citrate

```{r}
# search_bnf(keyword="alverine citrate") %>% View()

jic_medicines <- add_entries(
  meds = "alverine citrate", 
  bnf_to_add = search_bnf(keyword="alverine citrate") %>% filter(bnf_chemical_substance == "Alverine citrate")  # only Alverine citrate
  )
```

### bupropion hydrochloride

```{r}
# search_bnf(keyword="bupropion hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "bupropion hydrochloride", 
  bnf_to_add = search_bnf(keyword="bupropion hydrochloride")
  )
```

### minims atropine sulfate

```{r}
# search_bnf(keyword="minims atropine sulfate") %>% View()

jic_medicines <- add_entries(
  meds = "minims atropine sulfate", 
  bnf_to_add = search_bnf(keyword="minims atropine sulfate")
  )
```

### atropine sulfate

- Query whether only eye drops, or oral/parenteral is also relevant here? Also "Co-phenotrope (Diphenox hydrochloride/atropine sulfate)" for acute diarrhea?
- SM: I’d restrict to eye drops (it’s an off licence use but recognosed for PEOLC use)

```{r}
# search_bnf(keyword="atropine sulfate") %>% View()

jic_medicines <- add_entries(
  meds = "atropine sulfate", 
  bnf_to_add = search_bnf(keyword="atropine sulfate") %>%
    # minims already included by an above entry
    filter(!str_detect(bnf_product, regex("minims", ignore_case=TRUE))) %>%
    filter(bnf_chemical_substance_code == "1105000B0")  # eye drops only
  )
```

### minims cyclopentolate hydrochloride

```{r}
# search_bnf(keyword="minims cyclopentolate hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "minims cyclopentolate hydrochloride", 
  bnf_to_add = search_bnf(keyword="minims cyclopentolate hydrochloride")
  )
```

### cyclopentolate hydrochloride

```{r}
# search_bnf(keyword="cyclopentolate hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "cyclopentolate hydrochloride", 
  bnf_to_add = search_bnf(keyword="cyclopentolate hydrochloride") %>%
    # minims already included by an above entry
    filter(!str_detect(bnf_product, regex("minims", ignore_case=TRUE)))
  )
```

### dicycloverine hydrochloride

- I’ve included all formulations, including “Dicycloverine hydrochloride compound preparations (Kolanticon gel)”
- SM:agree

```{r}
# search_bnf(keyword="dicycloverine hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "dicycloverine hydrochloride", 
  bnf_to_add = search_bnf(keyword="dicycloverine hydrochloride")
  )
```

### erdosteine

```{r}
# search_bnf(keyword="erdosteine") %>% View()

jic_medicines <- add_entries(
  meds = "erdosteine", 
  bnf_to_add = search_bnf(keyword="erdosteine")
  )
```

### fesoterodine fumarate

```{r}
# search_bnf(keyword="fesoterodine fumarate") %>% View()

jic_medicines <- add_entries(
  meds = "fesoterodine fumarate", 
  bnf_to_add = search_bnf(keyword="fesoterodine fumarate")
  )
```

### flavoxate hydrochloride

```{r}
# search_bnf(keyword="flavoxate hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "flavoxate hydrochloride", 
  bnf_to_add = search_bnf(keyword="flavoxate hydrochloride")
  )
```

### glycopyrrolate bromide

Note: "glycopyrrolate" doesn't appear in the BNF, but "glycopyrronium" does.

Query: this appears under chapters: gastro-intestinal, skin, respiratory, as well as anaesthesia - which ones are relevant?

- glycopyrrolate bromide (glycopyrronium bromide in BNF) – this spans multiple chapters – gastro-intestinal (oral preparations), respiratory (inhalation), skin (topical, powder), and general anaesthesia (parenteral) – which of these are relevant? (the anaesthesia section has a combination of Neostigmine metilsulfite/glycopyrronium bromide also)

- SM: in my experience only the oral and sub cutaneous form is used in palliative care.  Definitely wouldn’t include the inhalation and topical/powder versions as they are used for non-palliative considitons.
- SM: don't include the Neostigmine metilsulfite/glycopyrronium bromide combination

```{r}
# search_bnf(keyword="glycopyrrolate bromide") %>% View()
# search_bnf(keyword="glycopyrrolate") %>% View()
# search_bnf(keyword="glycopyrr.* bromide") %>% View()

jic_medicines <- add_entries(
  meds = "glycopyrrolate bromide", 
  bnf_to_add = search_bnf(keyword="glycopyrr.* bromide") %>%
    filter(bnf_chemical_substance_code %in% c("0102000L0", "1501030G0"))  # oral & parenteral, respectively
  )
```

### hyoscine

- Query:this occurs in 2 forms, butylbromide (antispasmodic) and hydrobromide (nausea and vertigo) – are both relevant? And in all formulations (oral, injectable, transdermal patch)
- SM: all formulations and forms are used for PEOLC

```{r}
# search_bnf(keyword="hyoscine") %>% View()

jic_medicines <- add_entries(
  meds = "hyoscine", 
  bnf_to_add = search_bnf(keyword="hyoscine")
  )
```

### procyclidine hydrochloride

```{r}
# search_bnf(keyword="procyclidine hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "procyclidine hydrochloride", 
  bnf_to_add = search_bnf(keyword="procyclidine hydrochloride")
  )
```

### propantheline bromide

```{r}
# search_bnf(keyword="propantheline bromide") %>% View()

jic_medicines <- add_entries(
  meds = "propantheline bromide", 
  bnf_to_add = search_bnf(keyword="propantheline bromide")
  )
```

### propiverine hydrochloride

```{r}
# search_bnf(keyword="propiverine hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "propiverine hydrochloride", 
  bnf_to_add = search_bnf(keyword="propiverine hydrochloride")
  )
```

## anti-emetics

### betahistine hydrochloride

```{r}
# search_bnf(keyword="betahistine hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "betahistine hydrochloride", 
  bnf_to_add = search_bnf(keyword="betahistine hydrochloride")
  )
```

### cyclizine

- Keeping cyclizine hydrochloride and lactate forms, but not the combination "Morphine tartrate and cyclizine tartrate" in the Analgesics section (0407)
- SM: I’d include the combination version

```{r}
# search_bnf(keyword="cyclizine") %>% View()

jic_medicines <- add_entries(
  meds = "cyclizine", 
  bnf_to_add = search_bnf(keyword="cyclizine")  # don't filter out the combination with morphine tartrate
  )
```

### domperidone

```{r}
# search_bnf(keyword="domperidone") %>% View()

jic_medicines <- add_entries(
  meds = "domperidone", 
  bnf_to_add = search_bnf(keyword="domperidone")
  )
```

### granisetron hydrochloride

```{r}
# search_bnf(keyword="granisetron hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "granisetron hydrochloride", 
  bnf_to_add = search_bnf(keyword="granisetron hydrochloride")
  )
```

### levomepromazine hydrochloride

```{r}
# search_bnf(keyword="levomepromazine hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "levomepromazine hydrochloride", 
  bnf_to_add = search_bnf(keyword="levomepromazine hydrochloride")
  )
```

### metoclopramide hydrochloride

```{r}
# search_bnf(keyword="metoclopramide hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "metoclopramide hydrochloride", 
  bnf_to_add = search_bnf(keyword="metoclopramide hydrochloride")
  )
```

### ondansetron hydrochloride

```{r}
# search_bnf(keyword="ondansetron hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "ondansetron hydrochloride", 
  bnf_to_add = search_bnf(keyword="ondansetron hydrochloride")
  )
```

### prochlorperazine maleate

```{r}
# search_bnf(keyword="prochlorperazine maleate") %>% View()

jic_medicines <- add_entries(
  meds = "prochlorperazine maleate", 
  bnf_to_add = search_bnf(keyword="prochlorperazine maleate")
  )
```

## anxiolytics

### buspirone hydrochloride

```{r}
# search_bnf(keyword="buspirone hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "buspirone hydrochloride", 
  bnf_to_add = search_bnf(keyword="buspirone hydrochloride")
  )
```

### chlordiazepoxide hydrochloride

```{r}
# search_bnf(keyword="chlordiazepoxide hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "chlordiazepoxide hydrochloride", 
  bnf_to_add = search_bnf(keyword="chlordiazepoxide hydrochloride")
  )
```

### clobazam

```{r}
# search_bnf(keyword="clobazam") %>% View()

jic_medicines <- add_entries(
  meds = "clobazam", 
  bnf_to_add = search_bnf(keyword="clobazam")
  )
```

### clomethiazole

```{r}
# search_bnf(keyword="clomethiazole") %>% View()

jic_medicines <- add_entries(
  meds = "clomethiazole", 
  bnf_to_add = search_bnf(keyword="clomethiazole")
  )
```

### clonazepam

- Keeping only systemic clonazepam (BNF subparagraph “control of epilepsy”) and not parenteral (Clonazepam 1mg/1ml inj ampoules and diluent; subparagraph “drugs used in status epilepticus”)
- SM: sometimes parenteral can be used in terminal agitation or in brain tumours with seizures, so would include

```{r}
# search_bnf(keyword="clonazepam") %>% View()

jic_medicines <- add_entries(
  meds = "clonazepam", 
  bnf_to_add = search_bnf(keyword="clonazepam")  # don't remove parenteral
  )
```

### diazepam

- Keeping all formulations, oral solution/suspension, rectal solution tube, suppositories, injection ampoules
- SM: agree

```{r}
# search_bnf(keyword="diazepam") %>% View()

jic_medicines <- add_entries(
  meds = "diazepam", 
  bnf_to_add = search_bnf(keyword="diazepam")
  )
```

### loprazolam

```{r}
# search_bnf(keyword="loprazolam") %>% View()

jic_medicines <- add_entries(
  meds = "loprazolam", 
  bnf_to_add = search_bnf(keyword="loprazolam")
  )
```

### lorazepam

- Keeping all formulations, tablets, oral solution/suspension, injection ampoules
- SM: agree

```{r}
# search_bnf(keyword="lorazepam") %>% View()

jic_medicines <- add_entries(
  meds = "lorazepam", 
  bnf_to_add = search_bnf(keyword="lorazepam")
  )
```

### lormetazepam

```{r}
# search_bnf(keyword="lormetazepam") %>% View()

jic_medicines <- add_entries(
  meds = "lormetazepam", 
  bnf_to_add = search_bnf(keyword="lormetazepam")
  )
```

### midazolam

- keeping only BNF section 0408 “antiepileptic drugs” and not 1501 ”General anaesthesia”; of 0408, keeping both midazolam maleate (tablets, oral liquid, oral syringe) and midazolam hydrochloride (oral syringe, oral solution); (note: 1501 has injection ampoules and infusion vials, as well as oral solution)
- SM: agree

```{r}
# search_bnf(keyword="midazolam") %>% View()

jic_medicines <- add_entries(
  meds = "midazolam", 
  bnf_to_add = search_bnf(keyword="midazolam") %>%
    filter(bnf_section_code == "0408")
  )
```

### nitrazepam

- This picks up flunitrazepam (Rohypnol) also, but I didn't include that.
- SM: agree

```{r}
# search_bnf(keyword="nitrazepam") %>% View()

jic_medicines <- add_entries(
  meds = "nitrazepam", 
  bnf_to_add = search_bnf(keyword="nitrazepam") %>%
    filter(bnf_chemical_substance_code == "0401010R0")  # keep only nitrazepam, not flunitrazepam
  )
```

### temazepam

```{r}
# search_bnf(keyword="temazepam") %>% View()

jic_medicines <- add_entries(
  meds = "temazepam", 
  bnf_to_add = search_bnf(keyword="temazepam")
  )
```

### trazodone hydrochloride

```{r}
# search_bnf(keyword="trazodone hydrochloride") %>% View()

jic_medicines <- add_entries(
  meds = "trazodone hydrochloride", 
  bnf_to_add = search_bnf(keyword="trazodone hydrochloride")
  )
```

### zopiclone

```{r}
# search_bnf(keyword="zopiclone") %>% View()

jic_medicines <- add_entries(
  meds = "zopiclone", 
  bnf_to_add = search_bnf(keyword="zopiclone")
  )
```

### zolpidem tartrate

```{r}
# search_bnf(keyword="zolpidem tartrate") %>% View()

jic_medicines <- add_entries(
  meds = "zolpidem tartrate", 
  bnf_to_add = search_bnf(keyword="zolpidem tartrate")
  )
```

## Check I've included all of them

```{r}
stopifnot(
  nrow(
    ## this should be an empty table - findings meds that were listed in jic but not in jic_medicines
    jic %>% anti_join(jic_medicines, by = "meds")
  )==0
)
```

# Save resulting file

See @sec-introduction for version listing.

```{r}
# write_csv(jic_medicines, file = "./bnf/bnf_codes_just_in_case_v1.csv")  # v1 before SM checked
write_csv(jic_medicines, file = "./bnf/bnf_codes_just_in_case_v2.csv")  # v2 after SM checked
```