### 2025-10-27 place_of_death_report_2025.qmd
quick pared-down version of the full script for producing graphs of the latest figures for deaths at home in 2025

---
title: "Place of death trend in Scotland, 2025"
author: "Jan Savinc"
format: html
editor_options: 
  chunk_output_type: console
---

# Introduction

As of 2025-03-03, the NRS Covid-specific updates on weekly death registrations are no longer published. The new publication at https://www.nrscotland.gov.uk/publications/deaths-registered-weekly-in-scotland/ is updated weekly, and Table 7 contains the breakdown by place of death ("location").

# Preliminaries

## Load packages

```{r}
library(tidyverse)
library(readxl)  # for reading excel files
library(curl)  # for downloading files
library(janitor)  # for cleaning datasets quickly
library(rvest)  # for scraping NRS website
library(ISOweek)  # for dealing with ISO 8601 weeks
```

## Load functions

```{r}
source("./functions.R")
```

## Find latest file

```{r}
url_nrs_weekly_deaths <- "https://www.nrscotland.gov.uk/publications/deaths-registered-weekly-in-scotland/"

all_nrs_links <-
  read_html(url_nrs_weekly_deaths) %>%
  html_nodes("a") %>%
  html_attr("href") %>%  # find all links on the page
  tibble(links = .)

nrs_filename_pattern <- "weekly-deaths-.*\\.xlsx"

weekly_deaths_url_relative <-
  all_nrs_links %>%
  filter(str_detect(links, pattern = nrs_filename_pattern)) %>%
  filter(!str_detect(links, pattern = "monthly")) %>%  # remove monthly data
  filter(!str_detect(links, pattern = "excess")) %>%  # remove excess deaths data
  filter(!str_detect(links, pattern = "influenz")) %>%  # remove influenza data
  pull(links)

stopifnot(length(weekly_deaths_url_relative)==1)  # stop if more than one URL found

## get absolute URL
weekly_deaths_url_absolute <- url_absolute(weekly_deaths_url_relative, base = url_nrs_weekly_deaths)
```

## Download file

```{r}
date_string <- strftime(today(), format = "%F")
path_today <- file.path("./downloaded_data", date_string)
path_to_downloaded_file <- file.path(path_today, basename(weekly_deaths_url_absolute))

if (!dir.exists(path_today)) dir.create(path_today)

# download.file(url = weekly_deaths_url_absolute, destfile = path_to_downloaded_file, quiet = TRUE)
todays_file <- curl_fetch_disk(weekly_deaths_url_absolute, path = path_to_downloaded_file)
```

## Read file

```{r}
deaths_by_pod_raw <- 
  read_excel(path = path_to_downloaded_file, sheet = "7", skip = 4) %>%
  clean_names()
```

## Set up output directory

Because data are updated on a weekly basis, we'll have a separate directory for every week the data are updated:

```{r}
latest_data_modified_date <- strftime(todays_file$modified, format="%F")
latest_year <- year(latest_data_modified_date)

dir_outputs <- file.path("./outputs", latest_year, latest_data_modified_date)
if (!dir.exists(dir_outputs)) dir.create(dir_outputs, recursive = TRUE)
```

# Wrangle data

## Load previous data

```{r}
deaths_2015_to_2019 <- 
  read_csv("./historical_data/2020-04-30_weekly-deaths-by-location-2015-2019.csv", skip = 2) %>%
  rename(year = 1) %>%  # rename first column
  mutate(
    place_of_death = if_else(is.na(`1`), true = year, NA_character_)
  ) %>%
  fill(place_of_death, .direction="down") %>%  # LOCF on place_of_death
  drop_na(`1`) %>%
  pivot_longer(cols = all_of(as.character(1:53)), values_to = "n_deaths", names_to = "week_number") %>%
  mutate(week_number=as.integer(week_number)) %>%
  recode_place_of_death()

## construct column names for the 2020 data
headers_in_nrs_data_2020 <-
  c(
    "place_of_death",
    "cause_of_death",
    paste0("week_", 1:53),
    "blank_remove",  # this and the rest are irrelevant entries that we'll remove
    "remove_1",
    "remove_2"
  )

deaths_2020 <- read_excel(
  path = "./historical_data/2022-01-06_covid-deaths-21-data-week-52.xlsx",
  sheet = "Table 3 (2020)",
  col_names = headers_in_nrs_data_2020,
  range = cell_rows(6:125)
) %>%  # the data are inrows 6:125
  select(-matches("remove")) %>%  # remove irrelevant summary columns at the end
  mutate(across(.cols = matches("week"), .fns = as.numeric)) %>%  # remove erroneous "count1, count2, ..." entries due to superscript in previous column
  mutate(dataset = if_else(condition = str_detect(string = cause_of_death, pattern = regex(pattern = "^registered|^difference", ignore_case = TRUE)), true = cause_of_death, false = NA_character_)) %>%  # extract the Registered deaths: historical, 2020/2021, and Difference sections
  tidyr::fill(place_of_death, .direction = "down") %>%  # complete the entries for place of death from the "last observed"
  tidyr::fill(dataset, .direction = "down") %>%  # same as above but for "dataset" column
  drop_na(week_1) %>%  # remove blank "header" rows with no data recorded
  pivot_longer(cols = matches("week"), names_to = "week_number", values_to = "n_deaths", values_drop_na = TRUE) %>%  # convert to long data format
  mutate(
    week_number = as.integer(parse_number(week_number)),  # extract week number
    place_of_death =  # find best match in the reference names used in other datasets
      order_of_place_of_death_levels[stringdist::amatch(x = str_sub(place_of_death,1,4), table = str_sub(order_of_place_of_death_levels,1,4), method = "lv", maxDist = 5)],  # the reference is in the variable order_of_place_of_death_levels
    year = 2020,
    dataset = str_remove_all(dataset , "Registered deaths\\: "),
    date_w_c = compute_start_date_from_week_number(week_number = week_number, year = year),
    week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020)  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
    ) %>%
  filter(dataset!="Difference") %>%
  pivot_wider(names_from = dataset, values_from = n_deaths) %>%
  rename(deaths_mean = `five year average`, n_deaths = `2020`) %>%
  recode_place_of_death()

cell_ranges_for_causes_of_death_data_2021 <- list(
  "All" = "A6:V58",
  "Care home" = "A113:V165",
  "Home & other non-institution" = "A220:V272",
  "Hospital" = "A327:V379",
  "Other" = "A434:V486"
)

deaths_2021 <-
  map_dfr(
    .x = cell_ranges_for_causes_of_death_data_2021,
    .f = ~readxl::read_excel("./historical_data/2022-06-29_covid-deaths-22-data-week-25.xlsx", sheet = "8", range = .x), 
    .id = "place_of_death"
  ) %>%
  janitor::clean_names() %>%
  rename(year = registration_year, date_w_c = week_beginning) %>%
  select(-matches("_excess$")) %>%  # remove excess deaths, not currently used in analysis
  # select(-matches("_excess$|_average$")) %>%  # we can compute excess & averages later
  pivot_longer(cols = 5:ncol(.), names_to = "name", values_to = "n_deaths") %>%
  # filter(!str_detect(name, "all_causes")) %>%  # we can calculate that from the others
  mutate(
    year = as.integer(year),
    measure_type = if_else(condition = str_detect(name, "_five_year_average"), true = "deaths_mean", false = "n_deaths"),
    cause_of_death = str_remove_all(name, "_five_year_average|_deaths"),
    cause_of_death = return_pretty_cause_of_death(cause_of_death),
    date_w_c = as.Date(date_w_c, origin = "1899-12-30"),  # origin to account for Excel's integer date format
    week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020)  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
  ) %>%
  select(-name) %>%
  pivot_wider(names_from = measure_type, values_from = n_deaths) %>%
  filter(!is.na(n_deaths)) %>%  # remove blank entries - these are for the current year but for dates in the future
  recode_place_of_death() %>%
  mutate(cause_of_death = if_else(cause_of_death=="All causes", "All", cause_of_death))

cell_ranges_for_causes_of_death_data_2022 <- list(
  "All" = "A6:V58",
  "Care home" = "A61:V113",
  "Home & other non-institution" = "A116:V168",
  "Hospital" = "A171:V223",
  "Other" = "A226:V278"
)

deaths_2022 <-
  map_dfr(
    .x = cell_ranges_for_causes_of_death_data_2022,
    .f = ~readxl::read_excel("./historical_data/2023-08-02_covid-deaths-22-data-final.xlsx", sheet = "8", range = .x), 
    .id = "place_of_death"
  ) %>%
  janitor::clean_names() %>%
  rename(year = registration_year, date_w_c = week_beginning) %>%
  select(-matches("_excess$")) %>%  # remove excess deaths, not currently used in analysis
  # select(-matches("_excess$|_average$")) %>%  # we can compute excess & averages later
  pivot_longer(cols = 5:ncol(.), names_to = "name", values_to = "n_deaths") %>%
  # filter(!str_detect(name, "all_causes")) %>%  # we can calculate that from the others
  mutate(
    year = as.integer(year),
    measure_type = if_else(condition = str_detect(name, "_five_year_average"), true = "deaths_mean", false = "n_deaths"),
    cause_of_death = str_remove_all(name, "_five_year_average|_deaths"),
    cause_of_death = return_pretty_cause_of_death(cause_of_death),
    date_w_c = as.Date(date_w_c, origin = "1899-12-30"),  # origin to account for Excel's integer date format
    week_number_run_over = compute_week_run_over(week_number, year, start_year = 2020)  # invent a "run-over" week number - weeks 54+ are in 2021, weeks 106+ are in 2022
  ) %>%
  select(-name) %>%
  pivot_wider(names_from = measure_type, values_from = n_deaths) %>%
  filter(!is.na(n_deaths)) %>%  # remove blank entries - these are for the current year but for dates in the future
  recode_place_of_death() %>%
  mutate(cause_of_death = if_else(cause_of_death=="All causes", "All", cause_of_death))
```

## Convert new data to tidy (long) format

```{r}
deaths_by_pod_clean <-
  deaths_by_pod_raw %>%
  pivot_longer(cols = all_locations_all_causes:other_institution_covid_19_mentioned, names_to = c("place_of_death", "type_of_cause"), names_pattern = "(all_locations|care_home|hospital|home_non_institution|other_institution)\\_(all_causes|covid_19_mentioned)", values_to = "n_deaths") %>%
  recode_place_of_death() %>%
  mutate(
    year = as.integer(registration_year),
    cause_of_death = case_when(
      type_of_cause == "all_causes" ~ "All",
      type_of_cause == "covid_19_mentioned" ~ "Covid-19 mentioned",
      TRUE ~ NA_character_
    ),
    date_w_c = as_date(week_beginning)
  ) %>% 
  select(-c(type_of_cause, week_beginning, registration_year))
```

## Merge new data and old

```{r}
deaths_current <-
  bind_rows(
    deaths_2020 %>%
      filter(cause_of_death == "All"),
    deaths_2021 %>%
      filter(cause_of_death == "All"),
    deaths_2022 %>%
      filter(cause_of_death == "All"),
    deaths_by_pod_clean %>%
      filter(cause_of_death == "All")
  )

deaths_range_2015_2019 <-
  deaths_2015_to_2019 %>%
  filter(year != "average") %>%
  filter(!(year!=2015 & week_number==53)) %>%  # remove entries for w53 in years where there wasn't one, #ISOweekproblems
  group_by(place_of_death, week_number) %>%
  summarise(
    cause_of_death = "All",
    year_range = "2015-2019",
    mean_deaths = mean(n_deaths),
    min_deaths = min(n_deaths),
    max_deaths = max(n_deaths),
    .groups = "drop"
    )
```

## Annual totals so far

```{r}
deaths_annual <-
  bind_rows(
    deaths_2015_to_2019 %>% filter(year!="average" & place_of_death!="average") %>% mutate(year=as.integer(year), cause_of_death="All"),
    deaths_current %>% filter(cause_of_death=="All")
  ) %>%
  count(year, place_of_death, cause_of_death, wt = n_deaths, name = "n_deaths") %>%
  group_by(year, place_of_death != "All") %>%
  mutate(  # add percentage
    prop_deaths = n_deaths / sum(n_deaths)
  ) %>%
  ungroup %>%
  mutate(year = if_else(  # rename current year so it's clear when the data end
    year == max(deaths_current$year),
    paste0(year, " up to W", deaths_current %>% filter(year==max(deaths_current$year)) %>% pull(week_number) %>% max),
    as.character(year)
  ))
```

# Analyses

## Home deaths plot

```{r}
(fig_weekly_home_deaths_2015_2019_range <-
  deaths_current %>%
  filter(place_of_death == "Home & other non-institution") %>%
  left_join(deaths_range_2015_2019, join_by(place_of_death, cause_of_death, week_number)) %>%
  ggplot(aes(x = as_date(date_w_c))) +
  add_vertical_lines_for_start_of_year(years = 2020:2025) +
  geom_text(
    data = tibble(year = 2020:2025) %>%
      mutate(date_w_c = map_vec(.x = year, ~compute_start_date_from_week_number(week_number = 1, year_number = .x))),
    aes(x = date_w_c, y = 30, label = paste0("'",str_sub(year,3,4))), hjust = -0.5, angle = -30
  ) +
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "12 weeks") +
  scale_y_continuous(breaks = scales::breaks_pretty(n = 6), limits = c(0,NA)) +
  geom_line(aes(y = n_deaths)) +
  geom_line(aes(y = mean_deaths), linetype = "dotted", colour = "grey20") +
  geom_ribbon(aes(ymin = min_deaths, ymax = max_deaths), fill = "#4477aa", alpha = 0.5) +
  # geom_line(aes(y = deaths_mean), linetype = "dashed", colour="grey20") +
  # facet_grid(~place_of_death) +
  # scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
  #                     labels = c("All deaths", 
  #                                "Non Covid-19 deaths"),
  #                     guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
  #                     ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = NULL,
       y = "Deaths (N)",
       # subtitle = date_range_text,
       # caption = paste0(c(death_registrations_text,historic_period_2022_text,source_nrs_text), collapse="\n")
       ))

fig_weekly_home_deaths_2015_2019_range %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_weekly_home_deaths_2015_2019_range", latest_data_modified_date), collapse = "_")), extensions = c(".pdf",".png", ".svg"), plot = ., width = 28, height = 14, dpi = 300, units = "cm")
```

## Annual deaths

```{r}
# uncomment to check that the assertion about Other being ~0.5% is true
# deaths_annual %>%
#   filter(place_of_death=="Other")

(fig_proportion_deaths_by_place_annual <- 
  deaths_annual %>%
  filter(place_of_death!="All") %>%
  filter(!str_starts(year, "2025")) %>%
  filter(place_of_death != "Other") %>%
  ggplot(aes(x=year, y=prop_deaths, colour=place_of_death, shape=place_of_death)) +
  geom_point(size = rel(3)) +
  geom_line(aes(group = place_of_death)) +
  geom_text(
    data = . %>% filter(!year %in% as.character(2015:2018)) %>%
      mutate(y = if_else(place_of_death=="Care home", prop_deaths - 0.03, prop_deaths + 0.03)),
    aes(
      x = year,
      y = y,
      label = scales::label_percent(accuracy = 0.1)(prop_deaths)
    ), 
    show.legend = FALSE
  ) +
  theme_minimal() +
  theme(legend.title = element_blank(), legend.position = "top") +
  scale_colour_manual(values = viridisLite::viridis(n=4, option = "D")[1:3]) +
  # scale_colour_manual(values = RColorBrewer::brewer.pal(3, "Dark2")) +
  # scale_colour_manual(values = scales::pal_dichromat(name = "Categorical.12")(3)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1), limits = c(0, NA)) +
  labs(x = NULL, y = "Deaths (%)", caption = '"Other institutions" represent ~0.5% of annual deaths and are not shown.'))

fig_proportion_deaths_by_place_annual %>%
save_output_file(filename = file.path(dir_outputs, paste0(c("fig_proportion_deaths_by_place_annual", latest_data_modified_date), collapse = "_")), extensions = c(".png"), plot = ., width = 20, height = 10, dpi = 300, units = "cm")
```